### 2023-09-07 draft.Rmd
adding stuff

---
output:
  pdf_document:
    keep_tex: true
    citation_package: natbib
    number_sections: yes
fontsize: 11pt
# geometry: "left=1in,right=1in,top=1in,bottom=1in"
urlcolor: blue
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
bibliography: bibliography.bib
abstract: |
  asdf
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      # eval = FALSE,
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.pos = 'H', 
                      fig.align = 'center', 
                      fig.height = 3, 
                      fig.width = 5, 
                      fig.dpi = 300)

options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')

import::from(magrittr, `%>%`, `%<>%`)
library(ggplot2)
source('~/dev/pabm-grdpg/functions.R')
source('~/dev/manifold-block-models/functions.R')
```

```{=tex}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\PABM}{\mathrm{PABM}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}
\newcommand{\ER}{\text{Erd\"{o}s-R\'{e}nyi}}
\newcommand{\SBM}{\mathrm{SBM}}
\newcommand{\DCBM}{\mathrm{DCBM}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\MBM}{\mathrm{MBM}}
```

# Introduction

# Methods

# Simulation Study

## Classification

In this simulation experiment, the latent vectors were sampled along a Bezier curve defined by $g(t) = \begin{bmatrix} t^2 & 2 t (1-t) \end{bmatrix}^\top$. 
The timepoints $t_i$ were sampled as iid Beta random variables with two sets of parameters, $(\alpha_1, \beta_1)$ and $(\alpha_2, \beta_2)$. 
The setup is as follows:

1. Draw response variables $y_1, ..., y_n \iid \Multinomial(1/2, 1/2)$.
2. For each $i = 1, ..., n$, draw $t_i \mid y_i \indep \Betadist(\alpha_{y_i}, \beta_{y_i})$, such that
    * $\alpha_1 = 1$
    * $\beta_1 = 2$
    * $\alpha_2 = 2$
    * $\beta_2 = 1$
3. Construct each latent vector as $x_i = g(t_i)$ and compile them in data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
4. Sample graph and its adjacency matrix as $A \sim \RDPG(X)$.

For each graph, we constructed the ASE, which was used to estimate the parameters $(\hat{\alpha}, \hat{\beta})$ for the graph, using the maximum likelihood method. 
The estimated parameters were then used as predictors $y_1, ..., y_n$, setting aside half for training and half for testing. 
We investigated graphs of size $|V| = 32, 64, 128, 256$. 
The number of graphs for each experiment was set to $n = 64$.
For each (number of graphs, size of graph) pair, we performed 32 replicates. 
Figure \ref{fig:class-sim-ase} shows the ASE of one graph.  
Figure \ref{fig:classification-sim} shows the boxplots of the classification error rates.

```{r class-sim-ase, cache = TRUE}
K <- 2
a.vec <- c(1, 2)
b.vec <- c(2, 1)
n <- 256

z <- 1

t. <- rbeta(n, a.vec[z], b.vec[z])
x1 <- t. ^ 2
x2 <- 2 * t. * (1 - t.)
X <- cbind(x1, x2)
P <- X %*% t(X)
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)

curve.est <- estimate.bezier.curve.2(Xhat, 
                                     degree = 2, 
                                     intercept = FALSE, 
                                     initialization = 'isomap', 
                                     min.t = 0, max.t = 1)
ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2])) + 
  geom_point(aes(x = curve.est$X[, 1], y = curve.est$X[, 2]), 
             colour = 'red',
             size = .2) + 
  coord_fixed() + 
  labs(x = expression(x[1]), y = expression(x[2]))
```

```{r classification-sim}
sim.df <- readr::read_csv('~/Documents/postdoc-first-project/classification-sim.csv')

n.vec <- unique(sim.df$n)

sim.df %>% 
  ggplot() + 
  geom_boxplot(aes(x = n, y = 1 - acc, group = n)) + 
  scale_x_log10(breaks = sort(n.vec)) +
  # scale_y_log10() +
  theme_bw() + 
  labs(x = 'number of vertices', y = 'error rate')
```

## Regression

# Applications

HCP data

```{r hcp-eda, cache = TRUE}
data.dir <- '~/Downloads/repeated_10_scale_125/'
graph.list <- lapply(dir(data.dir), function(x) {
  graph <- igraph::read.graph(file.path(data.dir, x), format = 'graphml')
  A <- graph %>% 
    igraph::as_adjacency_matrix() %>% 
    as.matrix()
  z <- graph %>% 
    igraph::get.vertex.attribute() %>% 
    {.$dn_hemisphere} %>% 
    as.factor()
  return(list(A = A, z = z, id = substr(x, 1, 6)))
})

i <- 1
graph <- graph.list[[i]]
A <- graph$A
z <- graph$z
Xhat <- embedding(A, 2, 0, scale = TRUE)
clustering <- manifold.clustering(Xhat, 
                                  degree = 1,
                                  initialization = as.numeric(z),
                                  intercept = FALSE)
plot.estimated.curves(Xhat, clustering)
```

```{r hcp-gender, cache = TRUE}
hcp.df <- readr::read_csv('~/Documents/postdoc-first-project/hcp-eda.csv')
hcp.df %>% 
  na.omit() %>% 
  GGally::ggpairs(columns = 1:4, aes(colour = gender))
```
```{r hcp-age, cache = TRUE}
hcp.df <- readr::read_csv('~/Documents/postdoc-first-project/hcp-eda.csv')
hcp.df %>% 
  na.omit() %>% 
  dplyr::filter(age != '36+') %>% 
  GGally::ggpairs(columns = 1:4, aes(colour = age))
```

# Discussion

# Bibliography

### 2023-09-07 draft.tex
adding stuff

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{setspace}
\usepackage{float}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\setcitestyle{numbers,square,comma}
\usepackage{verbatim}
\usepackage{amsthm}
\usepackage{comment}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{plainnat}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={blue},
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\PABM}{\mathrm{PABM}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}
\newcommand{\ER}{\text{Erd\"{o}s-R\'{e}nyi}}
\newcommand{\SBM}{\mathrm{SBM}}
\newcommand{\DCBM}{\mathrm{DCBM}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\MBM}{\mathrm{MBM}}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{simulation-study}{%
\section{Simulation Study}\label{simulation-study}}

\hypertarget{classification}{%
\subsection{Classification}\label{classification}}

In this simulation experiment, the latent vectors were sampled along a
Bezier curve defined by
\(g(t) = \begin{bmatrix} t^2 & 2 t (1-t) \end{bmatrix}^\top\). The
timepoints \(t_i\) were sampled as iid Beta random variables with two
sets of parameters, \((\alpha_1, \beta_1)\) and \((\alpha_2, \beta_2)\).
The setup is as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Draw response variables \(y_1, ..., y_n \iid \Multinomial(1/2, 1/2)\).
\item
  For each \(i = 1, ..., n\), draw
  \(t_i \mid y_i \indep \Betadist(\alpha_{y_i}, \beta_{y_i})\), such
  that

  \begin{itemize}
  \tightlist
  \item
    \(\alpha_1 = 1\)
  \item
    \(\beta_1 = 2\)
  \item
    \(\alpha_2 = 2\)
  \item
    \(\beta_2 = 1\)
  \end{itemize}
\item
  Construct each latent vector as \(x_i = g(t_i)\) and compile them in
  data matrix
  \(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\).
\item
  Sample graph and its adjacency matrix as \(A \sim \RDPG(X)\).
\end{enumerate}

For each graph, we constructed the ASE, which was used to estimate the
parameters \((\hat{\alpha}, \hat{\beta})\) for the graph, using the
maximum likelihood method. The estimated parameters were then used as
predictors \(y_1, ..., y_n\), setting aside half for training and half
for testing. We investigated graphs of size \(|V| = 32, 64, 128, 256\).
The number of graphs for each experiment was set to \(n = 64\). For each
(number of graphs, size of graph) pair, we performed 32 replicates.
Figure \ref{fig:class-sim-ase} shows the ASE of one graph.\\
Figure \ref{fig:classification-sim} shows the boxplots of the
classification error rates.

\begin{center}\includegraphics{draft_files/figure-latex/class-sim-ase-1} \end{center}

\begin{center}\includegraphics{draft_files/figure-latex/classification-sim-1} \end{center}

\hypertarget{regression}{%
\subsection{Regression}\label{regression}}

\hypertarget{applications}{%
\section{Applications}\label{applications}}

HCP data

\begin{center}\includegraphics{draft_files/figure-latex/hcp-eda-1} \end{center}

\begin{center}\includegraphics{draft_files/figure-latex/hcp-gender-1} \end{center}

\begin{center}\includegraphics{draft_files/figure-latex/hcp-age-1} \end{center}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\renewcommand\refname{Bibliography}
  \bibliography{bibliography.bib}

\end{document}

### 2023-09-07 hcp-eda.R

@@ -1,0 +1,70 @@
+import::from(magrittr, `%>%`, `%<>%`)
+import::from(foreach, `%do%`, `%dopar%`)
+source('~/dev/pabm-grdpg/functions.R')
+source('~/dev/manifold-block-models/functions.R')
+library(mclust)
+library(ggplot2)
+
+doMC::registerDoMC(parallel::detectCores())
+
+data.dir <- '~/Downloads/repeated_10_scale_125/'
+graph.list <- lapply(dir(data.dir), function(x) {
+  graph <- igraph::read.graph(file.path(data.dir, x), format = 'graphml')
+  A <- graph %>% 
+    igraph::as_adjacency_matrix() %>% 
+    as.matrix()
+  z <- graph %>% 
+    igraph::get.vertex.attribute() %>% 
+    {.$dn_hemisphere} %>% 
+    as.factor()
+  return(list(A = A, z = z, id = substr(x, 1, 6)))
+})
+
+datamat.df <- foreach::foreach(i = seq_along(graph.list), .combine = dplyr::bind_rows, .errorhandling = 'remove') %dopar% {
+  # print(paste(i, 'out of', length(graph.list)))
+  graph <- graph.list[[i]]
+  A <- graph$A
+  z <- graph$z
+  Xhat <- embedding(A, 2, 0, scale = TRUE)
+  clustering <- manifold.clustering(Xhat, 
+                                    degree = 1,
+                                    initialization = as.numeric(z),
+                                    intercept = FALSE)
+  params.l <- EnvStats::ebeta(clustering$t[[1]])$parameters
+  params.r <- EnvStats::ebeta(clustering$t[[2]])$parameters
+  angle.l <- atan2(clustering$p[[1]][2, 2],
+                   clustering$p[[1]][2, 1])
+  angle.r <- atan2(clustering$p[[2]][2, 2],
+                   clustering$p[[2]][2, 1])
+  angle.diff <- abs(angle.r - angle.l)
+  if (angle.diff > pi) {
+    angle.diff <- 2 * pi - angle.diff
+  }
+  dplyr::tibble(id = graph$id,
+                a.l = params.l[1],
+                b.l = params.l[2],
+                a.r = params.r[1],
+                b.r = params.r[2],
+                angle.diff = angle.diff)
+}
+
+load('~/Downloads/HCP/HCP-Y/DTI_fMRI/Data.RData')
+
[…]

### 2023-09-07 simulations/angle-regression-sim.R

@@ -1,0 +1,79 @@
+import::from(magrittr, `%>%`)
+import::from(foreach, `%do%`, `%dopar%`)
+source('~/dev/pabm-grdpg/functions.R')
+source('~/dev/manifold-block-models/functions.R')
+library(ggplot2)
+
+doMC::registerDoMC(parallel::detectCores())
+
+nrep <- 32
+
+a.vec <- c(1, 1)
+b.vec <- c(1, 1)
+theta.min <- pi / 6
+theta.max <- pi / 3
+beta1 <- 1
+beta0 <- 1
+
+N <- 2 ^ 6
+n.vec <- 2 ^ c(10, 9, 8, 7)
+
+out.df <- foreach::foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+  print(paste('n =', n))
+  foreach::foreach(rep = seq(nrep), .combine = dplyr::bind_rows) %do% {
+    print(paste('rep', rep, 'out of', nrep))
+    datamat.df <- foreach::foreach(i = seq(N), .combine = dplyr::bind_rows, .errorhandling = 'remove') %dopar% {
+      z <- sample(seq(2), n, replace = TRUE)
+      t. <- rbeta(n, a.vec[z], b.vec[z])
+      theta <- runif(1, theta.min, theta.max)
+      y <- beta0 + beta1 * theta
+      x1 <- ifelse(z == 1, t., cos(theta) * t.)
+      x2 <- ifelse(z == 1, 0, sin(theta) * t.)
+      X <- cbind(x1, x2)
+      P <- X %*% t(X)
+      A <- draw.graph(P)
+      Xhat <- embedding(A, 2, 0)
+      clustering <- manifold.clustering(Xhat, 
+                                        degree = 1,
+                                        initialization = z,
+                                        intercept = FALSE)
+      angle.1 <- atan2(clustering$p[[1]][2, 2],
+                       clustering$p[[1]][2, 1])
+      angle.2 <- atan2(clustering$p[[2]][2, 2],
+                       clustering$p[[2]][2, 1])
+      theta.hat <- abs(angle.2 - angle.1)
+      if (theta.hat > pi) {
+        theta.hat <- 2 * pi - theta.hat
+      }
+      dplyr::tibble(y = y,
+                    theta = theta,
+                    theta.hat = theta.hat)
+    }
+    
+    train.ind <- sample(seq(N), N / 2)
[…]

### 2023-09-07 simulations/beta-regression-sim.R

@@ -1,0 +1,83 @@
+import::from(magrittr, `%>%`)
+import::from(foreach, `%do%`, `%dopar%`)
+source('~/dev/pabm-grdpg/functions.R')
+source('~/dev/manifold-block-models/functions.R')
+library(ggplot2)
+
+doMC::registerDoMC(parallel::detectCores())
+
+nrep <- 32
+
+K <- 2
+beta0 <- 1
+beta1 <- 2
+beta2 <- -1
+
+N <- 2 ^ 6
+n.vec <- 2 ^ c(9, 8, 7, 6)
+
+sim.dir <- '~/dev/multilayer-rdpg/simulations'
+
+out.df <- foreach::foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+  print(paste('n =', n))
+  foreach::foreach(rep = seq(nrep), .combine = dplyr::bind_rows) %do% {
+    print(paste('rep', rep, 'out of', nrep))
+    rep.filename <- paste0('beta-reg-n-', n, '-rep-', rep, '.csv')
+    if (rep.filename %in% dir(sim.dir)) {
+      one.rep.df <- readr::read_csv(file.path(sim.dir, rep.filename),
+                                    progress = FALSE,
+                                    show_col_types = FALSE)
+    } else {
+      datamat.df <- foreach::foreach(i = seq(N), .combine = dplyr::bind_rows, .errorhandling = 'remove') %dopar% {
+        a <- runif(1, 0, 2)
+        b <- runif(1, 0, 2)
+        y <- beta0 + beta1 * a + beta2 * b
+        t. <- rbeta(n, a, b)
+        x1 <- t. ^ 2
+        x2 <- 2 * t. * (1 - t.)
+        X <- cbind(x1, x2)
+        P <- X %*% t(X)
+        A <- draw.graph(P)
+        Xhat <- embedding(A, 2, 0)
+        curve.est <- estimate.bezier.curve.2(Xhat, 
+                                             degree = 2, 
+                                             intercept = FALSE, 
+                                             initialization = 'isomap', 
+                                             min.t = 0, max.t = 1)
+        t.hat <- curve.est$t
+        param.est <- EnvStats::ebeta(t.hat)
+        dplyr::tibble(y = y, 
+                      a = a, 
+                      b = b,
+                      a.hat = param.est$parameters[1], 
+                      b.hat = param.est$parameters[2])
+      }
[…]

### 2023-09-08 simulations/beta-regression-sim.R
more simulations

@@ -1,5 +1,5 @@
 import::from(magrittr, `%>%`)
-import::from(foreach, `%do%`, `%dopar%`)
+import::from(foreach, foreach, `%do%`, `%dopar%`)
 source('~/dev/pabm-grdpg/functions.R')
 source('~/dev/manifold-block-models/functions.R')
 library(ggplot2)
@@ -14,13 +14,22 @@ beta1 <- 2
 beta2 <- -1
 
 N <- 2 ^ 6
-n.vec <- 2 ^ c(9, 8, 7, 6)
+n.vec <- 2 ^ c(10, 9, 8, 7, 6)
 
 sim.dir <- '~/dev/multilayer-rdpg/simulations'
 
-out.df <- foreach::foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+parallelize.outer <- TRUE
+if (parallelize.outer) {
+  `%DO%` <- `%dopar%`
+  parallelize.curvefit <- FALSE
+} else {
+  `%DO%` <- `%do%`
+  parallelize.curvefit <- TRUE
+}
+
+out.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
   print(paste('n =', n))
-  foreach::foreach(rep = seq(nrep), .combine = dplyr::bind_rows) %do% {
+  foreach(rep = seq(nrep), .combine = dplyr::bind_rows) %do% {
     print(paste('rep', rep, 'out of', nrep))
     rep.filename <- paste0('beta-reg-n-', n, '-rep-', rep, '.csv')
     if (rep.filename %in% dir(sim.dir)) {
@@ -28,7 +37,14 @@ out.df <- foreach::foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
                                     progress = FALSE,
                                     show_col_types = FALSE)
     } else {
-      datamat.df <- foreach::foreach(i = seq(N), .combine = dplyr::bind_rows, .errorhandling = 'remove') %dopar% {
+      datamat.df <- foreach(i = seq(N), .combine = dplyr::bind_rows, .errorhandling = 'remove') %DO% {
+        print(paste(i, '/', N))
+        iter.filename <- paste0('beta-reg-n-', n, '-rep-', rep, '-iter-', i, '.csv')
+        if (iter.filename %in% dir(sim.dir)) {
+          one.iter.df <- readr::read_csv(file.path(sim.dir, iter.filename),
+                                         progress = FALSE,
[…]

### 2023-09-09 draft.Rmd
updating simulations

@@
 \newcommand{\MBM}{\mathrm{MBM}}
+\newcommand{\Poisson}{\mathrm{Poisson}}
 ```
@@
 
+1. Draw angles $\theta_1, ..., \theta_N \iid \Uniform(\pi / 6, \pi / 3)$.
+2. For each $k = 1, ..., N$, 
+    i. Draw $t_1, ..., t_n \iid \Uniform(0, 1)$;
+    ii. Draw $z_1, ..., z_n \iid \Multinomial(1/2, 1/2)$;
+    iii.
+For each $i = 1, ..., n$, set $x_i = \begin{cases} \begin{bmatrix} t_i & 0 \end{bmatrix}^\top & z_i = 1 \\ \begin{bmatrix} t_i \cos \theta_k & t_i \sin \theta_k \end{bmatrix}^\top & z_i = 2 \end{cases}$;
+    iv. Collect $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ and draw $A \sim \RDPG(X)$;
+    v. Set the response $y_k = \beta_0 + \beta_1 \theta_k$.
+
+In this simulation, we set $\beta_0 = \beta_1 = 1$. 
+The number of graphs was set to $N = 64$, and the number of vertices per graph was set to $n = 128, 256, 512, 1024$. 
+For each $n$, we simulated 32 replicates.
+
+```{r angle-reg-example, cache = TRUE}
+n <- 512
+a.vec <- c(1, 1)
+b.vec <- c(1, 1)
+theta.min <- pi / 6
+theta.max <- pi / 3
+beta1 <- 1
+beta0 <- 1
+
+z <- sample(seq(2), n, replace = TRUE)
+t. <- rbeta(n, a.vec[z], b.vec[z])
+theta <- runif(1, theta.min, theta.max)
+y <- beta0 + beta1 * theta
+x1 <- ifelse(z == 1, t., cos(theta) * t.)
+x2 <- ifelse(z == 1, 0, sin(theta) * t.)
+X <- cbind(x1, x2)
+P <- X %*% t(X)
+A <- draw.graph(P)
+Xhat <- embedding(A, 2, 0)
+clustering <- manifold.clustering(Xhat, 
+                                  degree = 1,
+                                  initialization = z,
+                                  intercept = FALSE)
+plot.estimated.curves(Xhat, clustering) + 
+  scale_colour_brewer(palette = 'Set1')
+```
+
+```{r angle-reg-results, cache = TRUE}
+out.df <- readr::read_csv('~/dev/multilayer-rdpg/simulations/angle-regression-sim.csv')
+n.vec <- c(128, 256, 512, 1024)
+N <- 64
+
+ggplot(out.df) + 
+  geom_boxplot(aes(x = n, y = mse * N, group = n)) + 
+  scale_x_log10(breaks = sort(n.vec)) +
+  scale_y_log10() +
+  labs(x = 'number of vertices', y = 'SSE') + 
+  theme_bw()
+```
+
 # Applications
 
-HCP data
+```{r load-hcp}
+load('~/Downloads/HCP/HCP-Y/DTI_fMRI/Data.RData')
+ids <- dimnames(dti.countV2)[[3]]
+```
 
-```{r hcp-eda, cache = TRUE}
-data.dir <- '~/Downloads/repeated_10_scale_125/'
-graph.list <- lapply(dir(data.dir), function(x) {
-  graph <- igraph::read.graph(file.path(data.dir, x), format = 'graphml')
-  A <- graph %>% 
-    igraph::as_adjacency_matrix() %>% 
-    as.matrix()
-  z <- graph %>% 
-    igraph::get.vertex.attribute() %>% 
-    {.$dn_hemisphere} %>% 
+In the first example, we analyzed fiber count data between brain regions from the Human Connectome Project (HCP). 
+When analyzing these data as graphs, we denote the regions as vertices and the fiber counts between pairs of regions as weighted edges. 
+A plausible statistical model for these data is to assume that the edge weights between pairs of vertices is Poisson distributed, i.e., the adjacency matrix is sampled as $A_{ij} \indep \Poisson(\Theta_{ij})$, where $\Theta \in \mathbb{R}_+^{n \times n}$ is a symmetric matrix of Poisson parameters.
+
+In this dataset, there are $N = 136$ graphs (corresponding to individual subjects), each with $n = 68$ vertices (corresponding to brain regions). 
+Analyzing these graphs as RDPGs reveals that the DCBM is a good candidate for these data (fig \ref{fig:hcp-ase}).
+
+```{r hcp-ase, cache = TRUE, fig.cap = 'ASE of one graph from the HCP dataset.
+The lines are fitted via $K$-curves clustering using degree = 1.
+The outputted clusters correspond exactly to the left (red) and right (blue) hemispheres.', fig.width = 10, fig.height = 10}
+id <- ids[1]
+A <- dti.countV2[, , id] %>% 
+  unique() %>% 
+  t() %>% 
+  unique()
+z <- colnames(A) %>% 
+  stringi::stri_sub(-1, -1) %>% 
   as.factor()
-  return(list(A = A, z = z, id = substr(x, 1, 6)))
-})
 
-i <- 1
-graph <- graph.list[[i]]
-A <- graph$A
-z <- graph$z
 Xhat <- embedding(A, 2, 0, scale = TRUE)
@@
                                   initialization = as.numeric(z),
-                                  intercept = FALSE)
-plot.estimated.curves(Xhat, clustering)
+                                  normalize = TRUE,
+                                  intercept = FALSE,
+                                  parallel = FALSE)
+plot.estimated.curves(Xhat, clustering) + 
+  scale_colour_brewer(palette = 'Set1') + 
+  ggrepel::geom_label_repel(aes(x = Xhat[, 1],
+                                y = Xhat[, 2],
+                                label = colnames(A)))
 ```
+
+For each graph, the
 
@@
   na.omit() %>% 
-  GGally::ggpairs(columns = 1:4, aes(colour = gender))
+  GGally::ggpairs(columns = 2:8, aes(colour = gender)) + 
+  scale_colour_brewer(palette = 'Set1')
 ```
@@
   dplyr::filter(age != '36+') %>% 
-  GGally::ggpairs(columns = 1:4, aes(colour = age))
+  GGally::ggpairs(columns = 2:8, aes(colour = age)) + 
+  scale_colour_brewer(palette = 'Set1')
 ```

### 2023-09-09 draft.tex
updating simulations

@@
 \newcommand{\MBM}{\mathrm{MBM}}
+\newcommand{\Poisson}{\mathrm{Poisson}}
 
@@
 
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Draw angles
+  \(\theta_1, ..., \theta_N \iid \Uniform(\pi / 6, \pi / 3)\).
+\item
+  For each \(k = 1, ..., N\),
+
+  \begin{enumerate}
+  \def\labelenumii{\roman{enumii}.}
+  \tightlist
+  \item
+    Draw \(t_1, ..., t_n \iid \Uniform(0, 1)\);
+  \item
+    Draw \(z_1, ..., z_n \iid \Multinomial(1/2, 1/2)\);
+  \item
+    For each \(i = 1, ..., n\), set
+    \(x_i = \begin{cases} \begin{bmatrix} t_i & 0 \end{bmatrix}^\top & z_i = 1 \\ \begin{bmatrix} t_i \cos \theta_k & t_i \sin \theta_k \end{bmatrix}^\top & z_i = 2 \end{cases}\);
+  \item
+    Collect
+    \(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\) and
+    draw \(A \sim \RDPG(X)\);
+  \item
+    Set the response \(y_k = \beta_0 + \beta_1 \theta_k\).
+  \end{enumerate}
+\end{enumerate}
+
+In this simulation, we set \(\beta_0 = \beta_1 = 1\). The number of
+graphs was set to \(N = 64\), and the number of vertices per graph was
+set to \(n = 128, 256, 512, 1024\). For each \(n\), we simulated 32
+replicates.
+
+\begin{center}\includegraphics{draft_files/figure-latex/angle-reg-example-1} \end{center}
+
+\begin{center}\includegraphics{draft_files/figure-latex/angle-reg-results-1} \end{center}
+
 \hypertarget{applications}{%
@@
 
-HCP data
+In the first example, we analyzed fiber count data between brain regions
+from the Human Connectome Project (HCP). When analyzing these data as
+graphs, we denote the regions as vertices and the fiber counts between
+pairs of regions as weighted edges. A plausible statistical model for
+these data is to assume that the edge weights between pairs of vertices
+is Poisson distributed, i.e., the adjacency matrix is sampled as
+\(A_{ij} \indep \Poisson(\Theta_{ij})\), where
+\(\Theta \in \mathbb{R}_+^{n \times n}\) is a symmetric matrix of
+Poisson parameters.
 
-\begin{center}\includegraphics{draft_files/figure-latex/hcp-eda-1} \end{center}
+In this dataset, there are \(N = 136\) graphs (corresponding to
+individual subjects), each with \(n = 68\) vertices (corresponding to
+brain regions). Analyzing these graphs as RDPGs reveals that the DCBM is
+a good candidate for these data (fig \ref{fig:hcp-ase}).
+
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/hcp-ase-1}
+
+}
+
+\caption{ASE of one graph from the HCP dataset.
+The lines are fitted via $K$-curves clustering using degree = 1.
+The outputted clusters correspond exactly to the left (red) and right (blue) hemispheres.}\label{fig:hcp-ase}
+\end{figure}
+
+For each graph, the
 

### 2023-09-09 hcp-eda.R

@@ -1,11 +1,21 @@
 import::from(magrittr, `%>%`, `%<>%`)
-import::from(foreach, `%do%`, `%dopar%`)
+import::from(foreach, foreach, `%do%`, `%dopar%`)
 source('~/dev/pabm-grdpg/functions.R')
 source('~/dev/manifold-block-models/functions.R')
 library(mclust)
 library(ggplot2)
 
 doMC::registerDoMC(parallel::detectCores())
+
+betadist.mom <- function(x) {
+  xbar <- mean(x)
+  s2 <- var(x)
+  norm.stat <- xbar * (1 - xbar) / s2 - 1
+  alpha <- xbar * norm.stat
+  beta <- (1 - xbar) * norm.stat
+  return(list(alpha = alpha, 
+              beta = beta))
+}
 
 data.dir <- '~/Downloads/repeated_10_scale_125/'
 graph.list <- lapply(dir(data.dir), function(x) {
@@ -20,18 +30,34 @@ graph.list <- lapply(dir(data.dir), function(x) {
   return(list(A = A, z = z, id = substr(x, 1, 6)))
 })
 
-datamat.df <- foreach::foreach(i = seq_along(graph.list), .combine = dplyr::bind_rows, .errorhandling = 'remove') %dopar% {
+load('~/Downloads/HCP/HCP-Y/DTI_fMRI/Data.RData')
+
+graph.list <- lapply(graph.list, function(graph) {
+  if (graph$id %in% as.character(dat.demo$Subject)) {
+    return(graph)
+  } else {
+    return(NULL)
+  }
+})
+graph.list <- graph.list[!sapply(graph.list, is.null)]
+
+datamat.df <- foreach(i = seq_along(graph.list), .combine = dplyr::bind_rows, .errorhandling = 'remove') %dopar% {
   # print(paste(i, 'out of', length(graph.list)))
   graph <- graph.list[[i]]
   A <- graph$A
   z <- graph$z
+  
   Xhat <- embedding(A, 2, 0, scale = TRUE)
   clustering <- manifold.clustering(Xhat, 
                                     degree = 1,
                                     initialization = as.numeric(z),
-                                    intercept = FALSE)
+                                    normalize = TRUE,
+                                    intercept = FALSE,
+                                    parallel = FALSE)
+  
   params.l <- EnvStats::ebeta(clustering$t[[1]])$parameters
   params.r <- EnvStats::ebeta(clustering$t[[2]])$parameters
+  
[…]

### 2023-09-09 simulations/beta-regression-sim.R

@@ -59,9 +59,12 @@ out.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
                                              degree = 2, 
                                              intercept = FALSE, 
                                              initialization = 'isomap', 
-                                             min.t = 0, max.t = 1,
+                                             # min.t = 0, max.t = 1,
+                                             normalize = TRUE,
                                              parallel = parallelize.curvefit)
+        plot(Xhat); points(curve.est$X, col = 2)
         t.hat <- curve.est$t
+        plot(t.hat, t.)
         param.est <- EnvStats::ebeta(t.hat)
         one.iter.df <- dplyr::tibble(y = y, 
                                      a = a, 

### 2023-11-14 draft.Rmd
still working on it

--- 2023-09-07 draft.Rmd
@@
 ---
-output:
-  pdf_document:
-    keep_tex: true
-    citation_package: natbib
-    number_sections: yes
+title: Classification and Regression on Random Dot Product Graphs
+
+# to produce blinded version set to 1
+blinded: 0
+
+authors: 
+- name: 
+  affiliation: Department of YYY, University of XXX
+
+keywords:
+- latent structure models
+- random dot product graph 
+- neuroimaging 
+- brain connectivity networks
+
+output: rticles::asa_article
+# output: rticles::arxiv_article
+# output:
+#   pdf_document:
+#     keep_tex: true
+#     citation_package: natbib
+#     number_sections: yes
+
 fontsize: 11pt
+
 # geometry: "left=1in,right=1in,top=1in,bottom=1in"
-urlcolor: blue
+
 header-includes:
@@
 - \usepackage{comment}
+- \usepackage{amsfonts}
+- \usepackage{pdfpages}
+- \usepackage{xcolor}
+- \usepackage{sectsty}
+
 bibliography: bibliography.bib
+biblio-style: apalike
+
 abstract: |
-  asdf
+  The random dot product graph (RDPG) has become a powerful modeling tool in uncovering latent structures within graphs.
+In particular, it has been shown that the RDPG describes a wide range of popular random graph models with rigid latent structures.
+More recently, joint modeling of mutliple random graphs that share common properties or structures across graphs have been introduced, such as the multilayer RDPG, multiple RPDG, multilayer stochastic block model, etc.
+In this work, we use these joint random graph models in the context of statistical learning, such as classification and regression, by introducing the multiple latent structure model, in which the graphs share a common latent structure with different parameters that correspond to different response variables.
+Then we propose various estimation techniques involving manifold learning to estimate these parameters and in turn predict the responses, with theorems guaranteeing convergence of the predictions.
+Simulations and applications on functional MRI data verify the performance of these methods.
 ---
@@
 import::from(magrittr, `%>%`, `%<>%`)
+import::from(foreach, foreach, `%do%`, `%dopar%`)
 library(ggplot2)
+library(ggraph)
 source('~/dev/pabm-grdpg/functions.R')
@@
 \newcommand{\Multinomial}{\mathrm{Multinomial}}
+\newcommand{\Categorical}{\mathrm{Categorical}}
 \newcommand{\dd}{\mathrm{d}}
@@
 \newcommand{\MBM}{\mathrm{MBM}}
+\newcommand{\LSM}{\mathrm{LSM}}
+\newcommand{\MLSM}{\mathrm{MLSM}}
+\newcommand{\Poisson}{\mathrm{Poisson}}
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
+\newtheorem{corollary}{Corollary}
+\newtheorem{proposition}{Proposition}
+\theoremstyle{remark}
+\newtheorem{remark}{Remark}
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}
+\newtheorem{example}{Example}
 ```
@@
 
+Graph and network data are now as ubiquitous as traditional feature data in the fields of sociology (e.g., social networks), neuroimaging (e.g., brain connectivity networks), and deep learning (e.g., graph neural networks).
+As a result, new statistical and machine learning methods have recently been developed to analyze network data. 
+One approach is to treat the network as a random graph that comes from some probability model. 
+In particular, if we sum up the network in an adjacency matrix $A \in \mathbb{R}^{n \times n}$, then one probability model might be to draw each element $A_{ij}$, which represents the existence of an edge or the edge weight from vertex $i$ to $j$, independently from some distribution, perhaps with a unique parameter for the pair $(i,j)$, e.g., $A_{ij} \indep F_{\theta_{ij}}$.
+The classical example of this is the $\ER$ model \citep{Gilbert:1959}, in which every edge is drawn from the same distribution, typically a Bernoulli distribution, using the same parameter, i.e., $A_{ij} \iid \Bernoulli (p)$.
+The inhomogeneous Bernoulli graph extends this by allowing each edge, represented by $A_{ij}$, to have its own parameter, $P_{ij}$, e.g., in the Bernoulli case, $A_{ij} \indep \Bernoulli(P_{ij})$. 
+Typically, the parameters are collected into an edge probability matrix (in the case of unweighted graphs) or edge parameter matrix (in the case of weighted graphs), denoted as $P \in \mathbb{R}^{n \times n}$.
+The network analysis problem in this setting is to estimate $P$ given $A$.
+
+If we let the parameter matrix $P$ be unconstrained, the inference problem is overparameterized. 
+On the other hand, the classical $\ER$ model is often too restrictive to describe real, observed networks. 
+Much work has been done to develop models that are constrained enough for robust statistical inference while being generalizable to describe a wide range of networks. 
+One such family of random graph models is the random dot product graph (RDPG), first proposed by \citet{10.1007/978-3-540-77004-6_11}, which is a type of latent space graph in which each vertex of the graph has a corresponding latent vector in a low-dimensional Euclidean space $\mathbb{R}^d$, and the edge parameter between each pair of vertices is determined by the dot product of the corresponding vectors.
+In this model, the constraint is the low rank of the parameter matrix $P$, assuming that the latent dimension $d$ is less than the number of vertices $n$. 
+Further constraints can be imposed on the RDPG in the form of distributional assumptions on the latent vectors or restricting the latent vectors to lie on subspaces or manifolds in the latent space \citep{athreya2020estimation}.
+
+It has been shown \citep{rubindelanchy2017statistical, Koo_2022} that the RDPG (as well as the generalized random dot product graph \citep{rubindelanchy2017statistical}) can describe a wide range of popular random graph models, such as the $\ER$ model, the stochastic block model (SBM) \citep{doi:10.1080/0022250X.1971.9989788}, degree corrected block model (DCBM) \citep{Karrer_2011}, and popularity adjusted block model (PABM) \citep{307cbeb9b1be48299388437423d94bf1}, and describing these models as RDPGs (or its generalized version) is useful for parameter estimation.
+The RDPG captures a wide range of phenomena that can be described as networks, and in fact any network that can be thought of as being sampled from a parameter matrix $P$ can be described as a (generalized) RDPG, especially if $P$ is low rank.
+The RPDG representation of these networks
+
 # Methods
+
+## Definitions and Models
+
+We begin by defining the RDPG and the LSM.
+
+\begin{definition}[Random dot product graph (RDPG) \citep{10.1007/978-3-540-77004-6_11}]
+\label{def:rdpg}
+Let $\mathcal{X}$ be a subset of $\mathbb{R}^d$ for some latent space dimension $d \geq 1$ such that for any $x_1, x_2 \in \mathcal{X}$, $x_1^\top x_2 \in [0, 1]$. 
+Let $F_\theta$ be a distribution with support $\mathcal{X}$ and parameters $\theta$, and sample $x_1, ..., x_n \iid F_\theta$. 
+A graph $G$ with adjacency matrix $A$ is a random dot product graph with latent vectors $X = \bigl[x_1 \; \cdots \; x_n\bigr]^\top$ drawn from distribution $F_\theta$ if $A \sim X X^\top$.
+
+We use the notation $A \sim \RDPG(F_\theta)$ to denote a random adjacency matrix $A$ drawn from latent vectors distributed as $F_\theta$. 
+\end{definition}
+
+\begin{remark}
+\label{remark:nonunique}
+The latent vectors of an RDPG are not unique. 
+Suppose that $P = X X^\top$ is the edge parameter matrix of an RDPG with latent positions $X$. 
+Then any orthogonal transformation $W$ on $X$ results in the same edge parameter matrix. 
+More precisely, let $\tilde{X} = X W$. 
+Then it is clear that $\tilde{X} \tilde{X}^\top = X W W^\top X^\top = X X^\top = P$ results in the same edge parameter matrix. 
+Thus, there are infinitely many latent vector configurations that can result in the same $P$, but for any two latent vector configurations, there exists an orthogonal mapping that connects the two. 
+\end{remark}
+
+\begin{definition}[Latent structure model (LSM) \citep{athreya2020estimation}]
+\label{def:lsm}
+Let $\mathcal{C} \subset \mathcal{X} \subset \mathbb{R}^d$ be a smooth, nonintersecting one-dimensional manifold on the domain of a RDPG as defined in definition \ref{def:rdpg}, parameterized by function $p(t) : [0, 1] \to \mathcal{C}$.
+Then if $t_1, ..., t_n \iid F_\theta$ for some distribution $F_\theta$ with support $[0, 1]$ and parameter $\theta$, each $x_i = p(t_i)$, and $ A \sim X X^\top$ for $X = \bigl[x_1 \; \cdots \; x_n\bigr]^\top$, $A$ is the adjacency matrix of a latent structure model on curve $\mathcal{C}$ with parameterization $p$ and underlying distribution $F_\theta$.
+
+We use the notation $A \sim \LSM(\mathcal{C}, F_\theta)$ or $A \sim \LSM(p, F_\theta)$ to denote an adjacency matrix $A$ drawn as an LSM on curve $\mathcal{C}$ or its parameterization $p$ with underlying distribution $F_\theta$.
+\end{definition}
+
+\begin{remark}
+Although \citet{athreya2020estimation} defined the LSM by a single one-dimensional manifold $\mathcal{C}$ in the latent space, in this paper, we will allow for the existence of multiple one-dimensional manifolds, $\mathcal{C}_1, ..., \mathcal{C}_K$, (i.e., mixture of manifolds distribution).
+This type of latent space mixture distribution is observed in networks with community structure (\citep{10.5555/3122009.3242083}). 
+For estimation, if the membership of each latent vector to the manifolds is known, then each manifold can be learned separately using the vectors that belong to that manifold. 
+If the memberships are not known, then we use an iterative algorithm to both cluster the latent vectors to a known number of manifolds and learn the manifolds using the cluster assignments.
+
+In the case of a mixture of $K$ curves, we use the notation $A \sim \LSM(\{\mathcal{C}_k\}_K, F_\theta, \alpha)$ or $A \sim \LSM(\{p_k\}_K, F_\theta, \alpha)$, where $\alpha = (\alpha_1, ..., \alpha_K)$, $\sum_{k=1}^K \alpha_k = 1$ is the mixture parameter.
+For simplicity, we only consider the case where the underlying distribution is the same for each curve. 
+\end{remark}
+
+A plausible inference task in the RDPG is to estimate the original latent vectors. 
+The adjacency spectral embedding \citep{doi:10.1080/01621459.2012.699795} is a consistent estimator of the latent vectors, up to some unknown orthogonal transformation.
+
+\begin{remark}[Sparsity parameter]
+\end{remark}
+
+\begin{definition}[Adjacency spectral embedding (ASE) \citep{doi:10.1080/01621459.2012.699795}]
+\label{def:ase}
+\end{definition}
+
+\begin{theorem}[Consistency of the ASE \citep{lyzinski2014}]
+\label{theorem:ase-consistency}
+\end{theorem}
+
+Theorem \ref{theorem:ase-consistency} implies that the embedding vectors of the ASE converge to the original latent vectors, up to some unidentifiable orthogonal transformation, and the maximum deviation from the original latent vectors after the orthogonal transformation is bounded by a value that decays to $0$.
+This implies that in the LSM, the ASE can lead to consistent estimation of $F_\theta$, the underlying distribution, and in fact, \citet{athreya2020estimation} showed exactly this.
+
+\begin{definition}[Multiple latent structure model (MLSM)]
+\label{def:mlsm}
+Let $\mathcal{C}^{(1)}, ..., \mathcal{C}^{(L)} \subset \mathbb{R}^d$ be a sequence of curves defining the latent positions of a sequence of $L$ LSMs as in definition \ref{def:lsm}. 
+Each $\mathcal{C}^{(\ell)}$ is parameterized by function $p^{(\ell)}(t) : [0, 1] \to \mathcal{C}^{(\ell)}$. 
+Let $p^{(1)}, ..., p^{(L)}$ be a sequence of functions with domain $[0, 1]$ that parameterize the curves $\mathcal{C}^{(1)}, ..., \mathcal{C}^{(L)}$, let $F$ be a parametric distribution with support $[0, 1]$, and let $\theta_1, ..., \theta_L$ be a sequence of parameters for distribution $F$.
+For each $p^{(\ell)}$, sample $t_1^{(\ell)}, ..., t_{n_\ell}^{(\ell)} \iid F_{\theta^{(\ell)}}$ for some distribution $F$ parameterized by $\theta_\ell$, and let $x_i^{(\ell)} = p_\ell(t_i^{(\ell)})$ for each $i = 1, ..., n_\ell$, again as in definition \ref{def:lsm}.
+Sample a sequence of $L$ adjacency matrices, each as $A^{(\ell)} \indep \LSM(p_\ell, F_{\theta_\ell})$. 
+Then the sequence $\{A^{(\ell)}\}_L$ is are the adjacency matrices of a multiple latent structure model with curves $\{C^{(\ell)}\}_L$ parameterized by $\{p^{(\ell)}\}$ and underlying distribution $F$ with parameters $\{\theta_\ell\}_L$.
+
+We use the notation $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$ to denote a sequence of adjacency matrices drawn from an MLSM with parameterizations $\{p^{(\ell)}\}$ and parameters $\{\theta_\ell\}$ on underlying distrubtion $F$.
+\end{definition}
+
+\begin{remark}
+\label{remark:mlsm-mixture}
+Again as in the case of a single LSM, we also allow for each $A^{(\ell)}$ to be sampled from a latent structure composed of $K$ curves with mixture parameter $\alpha^{(\ell)}$. 
+In this case, we use the notation $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{\mathcal{C}_k^{(\ell)}\}_{K, L}, F, \{\theta_\ell\}_L, \{\alpha^{(\ell)}\}_L)$ or $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p_k^{(\ell)}\}_{K, L}, F, \{\theta_\ell\}_L, \{\alpha^{(\ell)}\}_L)$.
+\end{remark}
+
+### Connections to Related Models
+
+In the MLSM defined in definition \ref{def:mlsm}, the only commonality that we assume from graph to graph is that they are all LSMs with the same underlying distribution family. 
+If we impose further restrictions, namely that the latent structures are linear, we obtain the multilayer random dot product graph \citep{jones2021multilayer}.
+
+\begin{example}[Comparison to the multilayer DCBM \citep{agterberg2022joint}, multi-RDPG \citep{nielsen2018multiple}, and MREG \citep{8889404}]
+In the $K$-community DCBM, the probability of an edge between a pair of vertices is given by
+
+$$A_{ij} \indep \Bernoulli(\omega_i \omega_j B_{z_i, z_j}),$$
+
+where $z_i \in \{1, ..., K\}$ is the community label for vertex $i$, $B_{k, \ell}$ is the block connectivity between communities $k$ and $\ell$, and $\omega_i$ is the degree correction parameter for vertex $i$.
+In order to preserve identifiability and uniqueness, a common constraint on these parameters is to set $\sum_{i : z_i = k} \omega_i^2 = 1$ \citep{Karrer_2011}. 
+As in
+
+The edge parameter matrix of a $K$-community DCBM with $n$ vertices can be decomposed as follows:
+
+$$P = \Omega B \Omega^\top,$$
+
+where $B$ is a $K \times K$ matrix of block connectivities and $\Omega$ is an $n \times K$ matrix such that $\Omega_{ik} = \omega_i$ if vertex $i$ is in community $k$ and $0$ otherwise. 
+Then it is clear that if $B$ is positive semidefinite matrix of rank $K$, $P$ can also be viewed as a $K$-dimensional RDPG, since the rows of $\Omega$ are normalized and can be seen as eigenvectors, and $B$ is full rank and can be decomposed into a diagonal matrix and a rotation matrix, i.e., $P = (\Omega V) \Lambda (\Omega V)^\top$.
+Furthermore, this matrix decomposition implies that the latent vectors lie on one of $K$ line segments that intersect at the origin \citep{rubindelanchy2017statistical}. 
+Thus, the DCBM is a special case of the LSM in which there are $K$ latent ``linear curves'' (i.e., lines) in $\mathbb{R}^K$.
+
+The multilayer DCBM as described by \citet{agterberg2022joint} extends this to a sequence of DCBMs with the same community structure, allowing the block connectivities and degree correction parameters to change for each layer but keeping the same community structure throughout, i.e., each element of $P^{(\ell)} = \Omega^{(\ell)} B^{(\ell)} (\Omega^{(\ell)})^\top$ can vary with $\ell$ but $\Omega^{(\ell)} = 0 \iff \Omega^{(\ell')} = 0$ and similarly, $\Omega^{(\ell)} > 0 \iff \Omega^{(\ell')} > 0$, for every pair $(\ell, \ell')$.
+On the other hand, if the further restriction of $\Omega^{(\ell)} = \Omega^{(\ell')}$ for all $(\ell, \ell')$, i.e., $P^{(\ell)} = \Omega B^{(\ell)} \Omega^\top$, we obtain a special case of the multiple RDPG (multi-RDPG) with the identity link function, as described by \citet{nielsen2018multiple}, or equivalently, a special case of the multiple random eigen graphs model (MREG), as described by \citet{8889404}.
+However, if the community labels are not identical from graph to graph, the sequence of DCBMs cannot be described as a multilayer DCBM, multi-RDPG, or MREG, but it can still be described as an MLSM. 
+\end{example}
+
+\begin{example}[Nonlinear MLSM and comparison to MREG and multi-RDPG]
+\end{example}
+
+### Classification and Regression on the MLSM
+
+To use the MLSM for classification or regression problems, we assign response variables $y_1, ..., y_L$ to each graph. 
+Then if each response $y_\ell$ depends on $p^{(\ell)}(t)$ (the parameterization of the $\ell^{th}$ curve) or $\theta_\ell$ (the parameter of the $\ell^{th}$ distribution), or some combination of the two, there is a plausible setup for a predictive modeling task for predicting $y_\ell$ after observing $A^{(\ell)}$.
+Four such scenarios are given:
+
+\begin{definition}[MLSM for classification 1]
+\label{def:mlsm-c-1}
+Let $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$, $y_1, ..., y_L \in \{1, ..., M\}$ be labels for each of the $L$ graphs of the MLSM, and each $\theta_\ell$ take on one of $M$ values $\{\phi_1, ..., \phi_M\}$ corresponding to its response, $y_\ell$, i.e., $\theta_\ell = \phi_{y_\ell}$.
+
+We observe the adjacency matrices $A^{(1)}, ..., A^{(L)}$ and the first $r$ labels $y_1, ..., y_r$, and the task is to predict the unknown labels $y_{r+1}, ..., y_L$.
+
+This can be described by the following generative model:
+
+\begin{enumerate}
+  \item Draw labels $y_1, ..., y_L \iid \Categorical(\{1, ..., M\}, \{\pi_1, ..., \pi_M\})$.
+  \item Set each $\theta_\ell = \phi_{y_\ell}$.
+  \item Draw adjacency matrices as $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$.
+\end{enumerate}
+\end{definition}
+
+\begin{definition}[MLSM for classification 2]
+\label{def:mlsm-c-2}
+Let $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{\mathcal{C}^{(\ell)}\}, F, \{\theta_\ell\}_L)$, $y_1, ..., y_L \in \{1, ..., M\}$ be labels for each of the $L$ graphs of the MLSM, and each $\mathcal{C}^{(\ell)}$ take on one of $M$ functional forms $\{p^{(1)}(t), ..., p^{(M)}(t)\}$ corresponding to its response variable, $y_\ell$, i.e., $\mathcal{C}^{(\ell)}$ is parameterized by $p^{(y_\ell)}(t)$.
+
+We observe the adjacency matrices $A^{(1)}, ..., A^{(L)}$ and the first $r$ labels $y_1, ..., y_r$, and the task is to predict the unknown labels $y_{r+1}, ..., y_L$.
+
+This can be described by the following generative model:
+
+\begin{enumerate}
+  \item Draw labels $y_1, ..., y_L \iid \Categorical(\{1, ..., M\}, \{\pi_1, ..., \pi_M\})$.
+  \item Set each $p^{(\ell)}(t) = p^{(y_\ell)}(t)$.
+  \item Draw adjacency matrices as $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$.
+\end{enumerate}
+
+As in remark \ref{remark:mlsm-mixture}, we also allow for each adjacency matrix $A^{(\ell)}$ to be drawn from a latent structure consisting of multiple curves $\{p^{(\ell)}_k\}_K$. 
+\end{definition}
+
+\begin{definition}[MLSM for regression 1]
+\label{def:mlsm-r-1}
+Let $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$, and suppose that for each $\ell = 1, ..., L$, the $\ell^{th}$ graph is coupled with a response variable, $y_\ell$, as $y_\ell \indep \mathcal{N}(\theta_\ell^\top \beta, \sigma^2)$.
+
+We observe the adjacency matrices $A^{(1)}, ..., A^{(L)}$ and the first $r$ response variables $y_1, ..., y_r$. 
+In this setting, there are two plausible inference tasks. 
+The first is to estimate the coefficient vector $\beta$. 
+The second is to predict the unobserved response variables $y_{r+1}, ..., y_L$. 
+\end{definition}
+
+\begin{definition}[MLSM for regression 2]
+\label{def:mlsm-r-2}
+Let $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$ and each $p^{(\ell)}(t) = p(t; \gamma_\ell)$, where $\gamma_\ell$ is the vector of parameters for function $p$. 
+Suppose that for each $\ell = 1, ..., L$, the $\ell^{th}$ graph is coupled with a response variable, $y_\ell$, as $y_\ell \indep \mathcal{N}(\gamma_\ell^\top \beta, \sigma^2)$.
+
+We observe the adjacency matrices $A^{(1)}, ..., A^{(L)}$ and the first $r$ response variables $y_1, ..., y_r$. 
+In this setting, there are two plausible inference tasks. 
+The first is to estimate the coefficient vector $\beta$. 
+The second is to predict the unobserved response variables $y_{r+1}, ..., y_L$.
+
+As in remark \ref{remark:mlsm-mixture}, we also allow for each adjacency matrix $A^{(\ell)}$ to be drawn from a latent structure consisting of multiple curves $\{p^{(\ell)}_k\}_K$. 
+\end{definition}
+
+In all of these settings, the ASE of $A^{(\ell)}$ provides some insight into both $p^{(\ell)}$ and $\theta_\ell$. 
+\citet{athreya2020estimation} showed that with some additional information, the ASE of $A^{(\ell)}$ can lead to a consistent estimator for $\theta_\ell$.
+
+## Main Results
 
@@
 
-In this simulation experiment, the latent vectors were sampled along a Bezier curve defined by $g(t) = \begin{bmatrix} t^2 & 2 t (1-t) \end{bmatrix}^\top$. 
+In this simulation experiment, the latent vectors were sampled along a Bezier curve defined by $g(t) = \bigl[ t^2 \; 2 t (1-t) \bigr]^\top$. 
 The timepoints $t_i$ were sampled as iid Beta random variables with two sets of parameters, $(\alpha_1, \beta_1)$ and $(\alpha_2, \beta_2)$. 
@@
     * $\beta_2 = 1$
-3. Construct each latent vector as $x_i = g(t_i)$ and compile them in data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
+3. Construct each latent vector as $x_i = g(t_i)$ and compile them in data matrix $X = \bigl[ x_1 \; \cdots \; x_n \bigr]^\top$.
 4. Sample graph and its adjacency matrix as $A \sim \RDPG(X)$.
@@
 We investigated graphs of size $|V| = 32, 64, 128, 256$. 
-The number of graphs for each experiment was set to $n = 64$.
+The number of graphs for each experiment was set to $L = 64$.
 For each (number of graphs, size of graph) pair, we performed 32 replicates. 
@@
 
-```{r class-sim-ase, cache = TRUE}
+```{r class-sim-ase, cache = TRUE, fig.cap = 'One simulated graph (left) and its ASE (right). The red curve is the fitted quadratic Bezier curve on the ASE.'}
 K <- 2
-a.vec <- c(1, 2)
-b.vec <- c(2, 1)
 n <- 256
 
-z <- 1
+z <- 2
 
-t. <- rbeta(n, a.vec[z], b.vec[z])
+t. <- rbeta(n, 1, 1)
 x1 <- t. ^ 2
@@
 
+A.graph <- tidygraph::as_tbl_graph(unname(A), directed = FALSE)
+graph.plot <- ggraph(A.graph, layout = 'auto') + 
+  geom_edge_link(alpha = n ^ -.6) + 
+  geom_node_point(show.legend = FALSE) + 
+  labs(colour = NULL) + 
+  theme_void() + 
+  scale_colour_brewer(palette = 'Set1')
+
 curve.est <- estimate.bezier.curve.2(Xhat, 
@@
                                      initialization = 'isomap', 
-                                     min.t = 0, max.t = 1)
-ggplot() + 
+                                     normalize = TRUE)
+ase.plot <- ggplot() + 
   geom_point(aes(x = Xhat[, 1], y = Xhat[, 2])) + 
-  geom_point(aes(x = curve.est$X[, 1], y = curve.est$X[, 2]), 
+  geom_line(aes(x = curve.est$X[, 1], y = curve.est$X[, 2]), 
              colour = 'red',
@@
   coord_fixed() + 
-  labs(x = expression(x[1]), y = expression(x[2]))
+  labs(x = expression(x[1]), y = expression(x[2])) + 
+  theme_bw()
+
+gridExtra::grid.arrange(graph.plot, ase.plot, nrow = 1)
 ```
 
-```{r classification-sim}
+```{r classification-sim, cache = TRUE, fig.cap = 'Median classification error rate and its IQR vs. number of vertices in each graph.', fig.height = 2, fig.width = 4}
 sim.df <- readr::read_csv('~/Documents/postdoc-first-project/classification-sim.csv')
@@
 sim.df %>% 
+  dplyr::group_by(n) %>% 
+  dplyr::summarise(acc.med = median(acc),
+                   acc.high = quantile(acc, .75),
+                   acc.low = quantile(acc, .25)) %>% 
   ggplot() + 
-  geom_boxplot(aes(x = n, y = 1 - acc, group = n)) + 
+  geom_line(aes(x = n, y = 1 - acc.med)) + 
+  geom_point(aes(x = n, y = 1 - acc.med)) + 
+  geom_errorbar(aes(x = n, ymin = 1 - acc.high, ymax = 1 - acc.low),
+                width = .1) + 
   scale_x_log10(breaks = sort(n.vec)) +
@@
 
+1. Draw angles $\theta_1, ..., \theta_N \iid \Uniform(\pi / 6, \pi / 3)$.
+2. For each $k = 1, ..., N$, 
+    i. Draw $t_1, ..., t_n \iid \Uniform(0, 1)$;
+    ii. Draw $z_1, ..., z_n \iid \Multinomial(1/2, 1/2)$;
+    iii. For each $i = 1, ..., n$, set $x_i = \begin{cases} \bigl[ t_i \; 0 \bigr]^\top \; z_i = 1 \\ \bigl[ t_i \cos \theta_\ell \; t_i \sin \theta_\ell \bigr]^\top \; z_i = 2 \end{cases}$;
+    iv. Collect $X = \bigl[ x_1 \; \cdots \; x_n \bigr]^\top$ and draw $A \sim \RDPG(X)$;
+    v. Set the response $y_\ell = \beta_0 + \beta_1 \theta_\ell$
+
+In this simulation, we set $\beta_0 = \beta_1 = 1$. 
+The number of graphs was set to $L = 64$, and the number of vertices per graph was set to $n = 128, 256, 512, 1024$. 
+For each $n$, we simulated 32 replicates.
+
+```{r angle-reg-example, cache = TRUE}
+n <- 512
+a.vec <- c(1, 1)
+b.vec <- c(1, 1)
+theta.min <- pi / 6
+theta.max <- pi / 3
+beta1 <- 1
+beta0 <- 1
+
+z <- sample(seq(2), n, replace = TRUE)
+t. <- rbeta(n, a.vec[z], b.vec[z])
+theta <- runif(1, theta.min, theta.max)
+y <- beta0 + beta1 * theta
+x1 <- ifelse(z == 1, t., cos(theta) * t.)
+x2 <- ifelse(z == 1, 0, sin(theta) * t.)
+X <- cbind(x1, x2)
+P <- X %*% t(X)
+A <- draw.graph(P)
+Xhat <- embedding(A, 2, 0)
+clustering <- manifold.clustering(Xhat, 
+                                  degree = 1,
+                                  initialization = z,
+                                  intercept = FALSE)
+ase.plot <- plot.estimated.curves(Xhat, clustering) +
+  scale_colour_brewer(palette = 'Set1')
+
+A.graph <- tidygraph::as_tbl_graph(unname(A), directed = FALSE)
+graph.plot <- ggraph(A.graph, layout = 'auto') + 
+  geom_edge_link(alpha = n ^ -.6) + 
+  geom_node_point(colour = 'black',
+                  fill = 'white',
+                  show.legend = FALSE) + 
+  labs(colour = NULL) + 
+  theme_void() + 
+  scale_colour_brewer(palette = 'Set1')
+
+gridExtra::grid.arrange(graph.plot, ase.plot, nrow = 1)
+```
+
+```{r angle-reg-results, cache = TRUE, fig.height = 2, fig.width = 4}
+sim.df <- readr::read_csv('~/dev/multilayer-rdpg/simulations/angle-regression-sim.csv')
+n.vec <- c(128, 256, 512, 1024)
+N <- 64
+
+sim.df %>% 
+  dplyr::group_by(n) %>% 
+  dplyr::summarise(mse.med = median(mse),
+                   mse.high = quantile(mse, .75),
+                   mse.low = quantile(mse, .25)) %>% 
+  ggplot() + 
+  geom_line(aes(x = n, y = mse.med)) + 
+  geom_point(aes(x = n, y = mse.med)) + 
+  geom_errorbar(aes(x = n, ymin = mse.high, ymax = mse.low),
+                width = .1) + 
+  scale_x_log10(breaks = sort(n.vec)) +
+  # scale_y_log10() +
+  theme_bw() + 
+  labs(x = 'number of vertices', y = 'MSE')
+```
+
 # Applications
 
-HCP data
+## Human Connectome Project Aging Study
 
-```{r hcp-eda, cache = TRUE}
-data.dir <- '~/Downloads/repeated_10_scale_125/'
-graph.list <- lapply(dir(data.dir), function(x) {
-  graph <- igraph::read.graph(file.path(data.dir, x), format = 'graphml')
-  A <- graph %>% 
-    igraph::as_adjacency_matrix() %>% 
+```{r load-hcp}
+load('~/Downloads/Data/HCP/HCP-A/DTI/Data.RData')
+z <- dti.info$hemisphere %>% 
+  as.factor()
+rois <- dti.info$aparc_aseg_region.names.short
+
+ase.stats.df <- readr::read_csv('~/dev/multilayer-rdpg/hcp-a-eda.csv')
+N <- nrow(ase.stats.df)
+n <- nrow(dat.NumFibers[[1]])
+
+ids <- ase.stats.df$id
+```
+
+In the first example, we analyzed fiber count data between brain regions from the Human Connectome Project (HCP). 
+When analyzing these data as graphs, we denote the regions as vertices and the fiber counts between pairs of regions as weighted edges. 
+A plausible statistical model for these data is to assume that the edge weights between pairs of vertices is Poisson distributed, i.e., the adjacency matrix is sampled as $A_{ij} \indep \Poisson(\Theta_{ij})$, where $\Theta \in \mathbb{R}_+^{n \times n}$ is a symmetric matrix of Poisson parameters.
+
+In this dataset, there are $N = `r N`$ graphs (corresponding to individual subjects), each with $n = `r n`$ vertices (corresponding to brain regions). 
+Analyzing these graphs as RDPGs reveals that the DCBM is a good candidate for these data (figure \ref{fig:hcp-ase}).
+
+```{r hcp-ase, cache = TRUE, fig.cap = 'One graph from the HCP dataset (left) and its ASE (right).
+In the ASE, the lines are fitted via $K$-curves clustering using degree = 1.
+The outputted clusters correspond exactly to the left (red) and right (blue) hemispheres.', fig.width = 15, fig.height = 5, out.width = '100%'}
+i <- 10
+A <- dat.NumFibers[[i]] %>% 
   as.matrix()
-  z <- graph %>% 
-    igraph::get.vertex.attribute() %>% 
-    {.$dn_hemisphere} %>% 
-    as.factor()
-  return(list(A = A, z = z, id = substr(x, 1, 6)))
-})
+A <- (A + t(A)) / 2
+diag(A) <- 0
 
-i <- 1
-graph <- graph.list[[i]]
-A <- graph$A
-z <- graph$z
 Xhat <- embedding(A, 2, 0, scale = TRUE)
@@
                                   initialization = as.numeric(z),
-                                  intercept = FALSE)
-plot.estimated.curves(Xhat, clustering)
+                                  normalize = TRUE,
+                                  intercept = FALSE,
+                                  parallel = FALSE)
+
+A.graph <- tidygraph::as_tbl_graph(unname(A), directed = FALSE)
+graph.plot <- ggraph(A.graph, layout = 'auto') + 
+  geom_edge_link(alpha = n ^ -.6) + 
+  geom_node_point(aes(colour = factor(z)),
+                  show.legend = FALSE) + 
+  labs(colour = NULL) + 
+  theme_void() + 
+  scale_colour_brewer(palette = 'Set1')
+
+ase.plot <- plot.estimated.curves(Xhat, clustering) +
+  # ggrepel::geom_label_repel(aes(x = Xhat[, 1],
+  #                               y = Xhat[, 2],
+  #                               label = rois)) +
+  scale_colour_brewer(palette = 'Set1')
+gridExtra::grid.arrange(graph.plot, ase.plot, nrow = 1)
 ```
 
-```{r hcp-gender, cache = TRUE}
-hcp.df <- readr::read_csv('~/Documents/postdoc-first-project/hcp-eda.csv')
-hcp.df %>% 
-  na.omit() %>% 
-  GGally::ggpairs(columns = 1:4, aes(colour = gender))
+```{r age-lm}
+set.seed(12345)
+
+train.ind <- sample(c(TRUE, FALSE), nrow(ase.stats.df), replace = TRUE)
+train.df <- ase.stats.df[train.ind, ]
+test.df <- ase.stats.df[!train.ind, ]
+
+angle.lm <- lm(age ~ angle.diff, data = train.df)
+y.hat.angle <- predict(angle.lm, newdata = test.df)
+
+degree.lm <- lm(age ~ deg.within + deg.between, data = train.df)
+y.hat.deg <- predict(degree.lm, newdata = test.df)
+
+rmse.angle <- mean((test.df$age - y.hat.angle) ^ 2) ^ .5
+rmse.deg <- mean((test.df$age - y.hat.deg) ^ 2) ^ .5
 ```
-```{r hcp-age, cache = TRUE}
-hcp.df <- readr::read_csv('~/Documents/postdoc-first-project/hcp-eda.csv')
-hcp.df %>% 
-  na.omit() %>% 
-  dplyr::filter(age != '36+') %>% 
-  GGally::ggpairs(columns = 1:4, aes(colour = age))
+
+```{r hcp-age-regression, fig.cap = 'Scatterplot between the subject age (in months) vs. the fitted angle between hemispheres of the brain in the latent space.'}
+ggplot(ase.stats.df) + 
+  geom_point(aes(x = angle.diff, 
+                 y = age)) + 
+  labs(x = 'angle between hemispheres in latent space (radians)',
+       y = 'subject age (months)') + 
+  theme_bw() + 
+  stat_smooth(aes(x = angle.diff, y = age), method = 'lm')
+```
+
+```{r hcp-age-regression-degree, fig.cap = 'Scatterplot between the subject age (in months) vs.
+the average degree within (left) and between (right) hemispheres.', fig.width = 10, fig.height = 3, out.width = '100%'}
+w.plot <- ggplot(ase.stats.df) + 
+  geom_point(aes(x = deg.within, 
+                 y = age)) + 
+  labs(x = 'average degree within hemispheres',
+       y = 'subject age (months)') + 
+  theme_bw() + 
+  stat_smooth(aes(x = deg.within, y = age), method = 'lm')
+b.plot <- ggplot(ase.stats.df) + 
+  geom_point(aes(x = deg.between, 
+                 y = age)) + 
+  labs(x = 'average degree between hemispheres',
+       y = 'subject age (months)') + 
+  theme_bw() + 
+  stat_smooth(aes(x = deg.between, y = age), method = 'lm')
+gridExtra::grid.arrange(w.plot, b.plot, nrow = 1)
+```
+
+The ASE suggests a latent structure comprised of two line segments, one for each hemisphere of the brain, that meet at the origin. 
+One possible parameter is the angle between the two lines. 
+The estimated angles were observed to correlate with the subject's age, with wider angles corresponding to older subjects (figure \ref{fig:hcp-age-regression}). 
+A linear regression setting aside half of the brain connectivity graphs as test data achieves an RMSE of $`r round(rmse.angle, 2)`$ months.
+
+Other plausible statistics from the graphs are the average degrees within and between hemispheres. 
+When analyzing these as stochastic block models (SBM) or degree corrected block models (DCBM), these statistics would roughly correspond to the expected edge weights within and between communities. 
+Both of these statistics are correlated with the target value (figure \ref{fig:hcp-age-regression-degree}), but a linear model using them (i.e., $age \sim degree_{within} + degree_{between}$) achieves a higher RMSE ($`r round(rmse.deg, 2)`$ months) while adding another parameter to the model.
+Similarly, the assortativity of each graph with respect to hemisphere results in a higher RMSE of $156.5$ when regressed on age.
+
+We also note that this structure is consistent with the DCBM. 
+The multilayer DCBM was previously studied by \citet{agterberg2022joint}
+
+```{r}
+pairwise.angles.df <- readr::read_csv('~/dev/multilayer-rdpg/pairwise-angles.csv')
+
+dti.info.df <- dti.info %>% 
+  dplyr::select(roi = aparc_aseg_region.names,
+                rsn = RSN,
+                hemisphere = hemisphere) %>% 
+  dplyr::arrange(hemisphere, rsn)
+
+pairwise.angles.df %<>%
+  dplyr::mutate(roi1 = factor(roi1, levels = dti.info.df$roi),
+                roi2 = factor(roi2, levels = dti.info.df$roi))
+```
+
+```{r cosine-heatmap, cache = TRUE, fig.width = 20, fig.height = 20, out.width = '100%'}
+# i <- ids[1]
+# 
+# pairwise.angles.df %>% 
+#   dplyr::filter(id == i) %>% 
+#   ggplot() +
+#   geom_tile(aes(x = roi1, y = roi2, fill = angle)) +
+#   viridis::scale_fill_viridis() +
+#   coord_fixed() +
+#   theme_minimal() +
+#   labs(x = NULL, y = NULL) +
+#   theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1))
+
+# 36-46.5
+# 46.5-56.5
+# 56.5-69
+# 69-90
+
+pairwise.angles.df %>%
+  dplyr::group_by(roi1, roi2) %>%
+  dplyr::mutate(mean.angle = mean(angle)) %>%
+  dplyr::ungroup() %>%
+  dplyr::mutate(age = ifelse(age < 46.5, '36-46.5',
+                             ifelse(age < 56.5, '46.5-56.5',
+                                    ifelse(age < 69, '56.5-69',
+                                           '69-90')))) %>%
+  dplyr::group_by(age, roi1, roi2, rsn1, rsn2) %>%
+  dplyr::summarise(angle = mean(angle) - mean.angle) %>%
+  dplyr::ungroup() %>%
+  ggplot() +
+  geom_tile(aes(x = roi1, y = roi2, fill = angle)) +
+  viridis::scale_fill_viridis() +
+  coord_fixed() +
+  theme_minimal() +
+  theme(legend.position = 'bottom') +
+  labs(x = NULL, y = NULL) +
+  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1),
+        legend.key.width = unit(2, 'cm')) +
+  facet_wrap(~ age)
+```
+
+```{r cosine-heatmap-2, cache = TRUE, fig.width = 10, fig.height = 10, out.width = '100%'}
+angles.plot.1 <- pairwise.angles.df %>%
+  dplyr::filter(age < 46.5) %>%
+  dplyr::group_by(roi1, roi2) %>%
+  dplyr::summarise(angle = mean(angle)) %>%
+  dplyr::ungroup() %>%
+  ggplot() +
+  geom_tile(aes(x = roi1, y = roi2, fill = angle)) +
+  viridis::scale_fill_viridis() +
+  coord_fixed() +
+  theme_minimal() + 
+  theme(legend.position = 'bottom') +
+  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1),
+        legend.key.width = unit(2, 'cm')) +
+  labs(x = NULL, y = NULL, title = '36-46.5') +
+  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1))
+angles.plot.2 <- pairwise.angles.df %>%
+  dplyr::filter(age < 56.5, age >= 36) %>%
+  dplyr::group_by(roi1, roi2) %>%
+  dplyr::summarise(angle = mean(angle)) %>%
+  dplyr::ungroup() %>%
+  ggplot() +
+  geom_tile(aes(x = roi1, y = roi2, fill = angle)) +
+  viridis::scale_fill_viridis() +
+  coord_fixed() +
+  theme_minimal() +
+  theme(legend.position = 'bottom') +
+  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1),
+        legend.key.width = unit(2, 'cm')) +
+  labs(x = NULL, y = NULL, title = '46.5-56.5') +
+  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1))
+angles.plot.3 <- pairwise.angles.df %>%
+  dplyr::filter(age < 69, age >= 56.5) %>%
+  dplyr::group_by(roi1, roi2) %>%
+  dplyr::summarise(angle = mean(angle)) %>%
+  dplyr::ungroup() %>%
+  ggplot() +
+  geom_tile(aes(x = roi1, y = roi2, fill = angle)) +
+  viridis::scale_fill_viridis() +
+  coord_fixed() +
+  theme_minimal() +
+  theme(legend.position = 'bottom') +
+  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1),
+        legend.key.width = unit(2, 'cm')) +
+  labs(x = NULL, y = NULL, title = '56.5-69') +
+  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1))
+angles.plot.4 <- pairwise.angles.df %>%
+  dplyr::filter(age >= 69) %>%
+  dplyr::group_by(roi1, roi2) %>%
+  dplyr::summarise(angle = mean(angle)) %>%
+  dplyr::ungroup() %>%
+  ggplot() +
+  geom_tile(aes(x = roi1, y = roi2, fill = angle)) +
+  viridis::scale_fill_viridis() +
+  coord_fixed() +
+  theme_minimal() +
+  theme(legend.position = 'bottom') +
+  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1),
+        legend.key.width = unit(2, 'cm')) +
+  labs(x = NULL, y = NULL, title = '69-90') +
+  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1))
+
+angles.plot.1
+angles.plot.2
+angles.plot.3
+angles.plot.4
+
+```
+
+```{r roi-angles, cache = TRUE, out.width = '100%', fig.height = 10, fig.width = 10}
+# pairwise.angles.df %>% 
+#   dplyr::mutate(hemisphere1 = ifelse(grepl('(rh-)|(Right-)', roi1), 'R', 'L'),
+#                 hemisphere2 = ifelse(grepl('(rh-)|(Right-)', roi2), 'R', 'L')) %>% 
+#   dplyr::mutate(roi1 = gsub('(ctx-rh-)|(Right-)|(ctx-lh-)|(Left-)', '', roi1),
+#                 roi2 = gsub('(ctx-rh-)|(Right-)|(ctx-lh-)|(Left-)', '', roi2)) %>% 
+#   dplyr::filter(roi1 == roi2, hemisphere1 != hemisphere2) %>% 
+#   ggplot() + 
+#   geom_point(aes(x = age, y = angle, groups = roi1),
+#              size = .1) +
+#   facet_wrap(~ roi1) + 
+#   theme_bw()
+```
+
+```{r cache = TRUE}
+assort.df <- foreach(id = ids, .combine = dplyr::bind_rows) %dopar% {
+  A <- dat.NumFibers[[id]] %>% 
+    as.matrix()
+  A <- (A + t(A)) / 2
+  diag(A) <- 0
+  # A <- A / mean(A[upper.tri(A)])
+
+  A.graph <- igraph::graph_from_adjacency_matrix(A) %>% 
+    igraph::set_vertex_attr('hemisphere', value = z)
+
+  assortativity <- igraph::assortativity(A.graph, z)
+  n_triangles <- igraph::count_triangles(A.graph, z) %>% 
+    unique()
+  transitivity <- igraph::transitivity(A.graph)
+  modularity <- igraph::modularity(A.graph, z)
+
+  individual.stats.df <- dplyr::tibble(
+    id = id,
+    assortativity = assortativity,
+    n_tri1 = n_triangles[1],
+    n_tri2 = n_triangles[2],
+    transitivity = transitivity,
+    modularity = modularity
+  )
+
+  return(individual.stats.df)
+} %>% 
+  # dplyr::filter(!(id %in% ids.to.remove)) %>% 
+  dplyr::inner_join(
+    demo.data %>% 
+      dplyr::filter(QC_Issue_Codes == '',
+                    Full_MR_Compl == 1) %>% 
+      dplyr::transmute(id = src_subject_id,
+                       sex = sex,
+                       age = as.numeric(interview_age)) %>% 
+      dplyr::filter(age < 1200), 
+    by = 'id'
+  )
+
+dplyr::tibble(Metric = c('Angle between hemispheres',
+                         'Degree within hemisphere',
+                         'Degree between hemispheres',
+                         'Number of triangles in left hemisphere',
+                         'Number of triangles in right hemisphere',
+                         'Assortativity w.r.t. hemisphere',
+                         'Transitivity',
+                         'Modularity w.r.t. hemisphere'),
+              Correlation = c(cor(ase.stats.df$age, ase.stats.df$angle.diff),
+                              cor(ase.stats.df$age, ase.stats.df$deg.within),
+                              cor(ase.stats.df$age, ase.stats.df$deg.between),
+                              cor(assort.df$age, assort.df$n_tri1),
+                              cor(assort.df$age, assort.df$n_tri2),
+                              cor(assort.df$age, assort.df$assortativity),
+                              cor(assort.df$age, assort.df$transitivity),
+                              cor(assort.df$age, assort.df$modularity))) %>% 
+  knitr::kable(digits = 3, caption = 'Correlation between age and various graph metrics')
 ```
@@
 
+\newpage
+
 # Bibliography

### 2023-11-14 draft.tex
still working on it

@@
 \PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
-\documentclass[
-  11pt,
-]{article}
-\usepackage{amsmath,amssymb}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math} % this also loads fontspec
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-\usepackage{lmodern}
-\ifPDFTeX\else
-  % xetex/luatex font selection
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\usepackage[margin=1in]{geometry}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
+% !TeX program = pdfLaTeX
+\documentclass[12pt]{article}
+\usepackage{amsmath}
+\usepackage{graphicx,psfrag,epsf}
+\usepackage{enumerate}
+\usepackage[]{natbib}
+\usepackage{textcomp}
+
+%\pdfminorversion=4
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{0}
+
+% DON'T change margins - should be 1 inch all around.
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-1in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.7in}%
+\addtolength{\topmargin}{-1in}%
+
+%% load any required packages here
+
+% tightlist command for lists without linebreak
 \providecommand{\tightlist}{%
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
+
+% From pandoc table feature
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+
 \usepackage{setspace}
@@
 \usepackage{comment}
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\usepackage[]{natbib}
-\bibliographystyle{plainnat}
+\usepackage{amsfonts}
+\usepackage{pdfpages}
+\usepackage{xcolor}
+\usepackage{sectsty}
+
 \IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same}
 \hypersetup{
-  colorlinks=true,
-  linkcolor={Maroon},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={blue},
+  pdftitle={Classification and Regression on Random Dot Product Graphs},
+  pdfkeywords={latent structure models, random dot product
+graph, neuroimaging, brain connectivity networks},
+  hidelinks,
   pdfcreator={LaTeX via pandoc}}
 
-\author{}
-\date{\vspace{-2.5em}}
+\begin{document}
 
-\begin{document}
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\if0\blind
+{
+  \title{\bf Classification and Regression on Random Dot Product Graphs}
+
+  \author{
+         \\
+    Department of YYY, University of XXX\\
+      }
+  \maketitle
+} \fi
+
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\LARGE\bf Classification and Regression on Random Dot Product
+Graphs}
+  \end{center}
+  \medskip
+} \fi
+
+\bigskip
+\begin{abstract}
+The random dot product graph (RDPG) has become a powerful modeling tool
+in uncovering latent structures within graphs. In particular, it has
+been shown that the RDPG describes a wide range of popular random graph
+models with rigid latent structures. More recently, joint modeling of
+mutliple random graphs that share common properties or structures across
+graphs have been introduced, such as the multilayer RDPG, multiple RPDG,
+multilayer stochastic block model, etc. In this work, we use these joint
+random graph models in the context of statistical learning, such as
+classification and regression, by introducing the multiple latent
+structure model, in which the graphs share a common latent structure
+with different parameters that correspond to different response
+variables. Then we propose various estimation techniques involving
+manifold learning to estimate these parameters and in turn predict the
+responses, with theorems guaranteeing convergence of the predictions.
+Simulations and applications on functional MRI data verify the
+performance of these methods.
+\end{abstract}
+
+\noindent%
+{\it Keywords:} latent structure models, random dot product
+graph, neuroimaging, brain connectivity networks
+
+\vfill
+
+\newpage
+\spacingset{1.9} % DON'T change the spacing!
 
@@
 \newcommand{\MBM}{\mathrm{MBM}}
+\newcommand{\LSM}{\mathrm{LSM}}
+\newcommand{\MLSM}{\mathrm{MLSM}}
 \newcommand{\Poisson}{\mathrm{Poisson}}
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
+\newtheorem{corollary}{Corollary}
+\newtheorem{proposition}{Proposition}
+\theoremstyle{remark}
+\newtheorem{remark}{Remark}
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}
+\newtheorem{example}{Example}
 
@@
 
+Graph and network data are now as ubiquitous as traditional feature data
+in the fields of sociology (e.g., social networks), neuroimaging (e.g.,
+brain connectivity networks), and deep learning (e.g., graph neural
+networks). As a result, new statistical and machine learning methods
+have recently been developed to analyze network data. One approach is to
+treat the network as a random graph that comes from some probability
+model. In particular, if we sum up the network in an adjacency matrix
+\(A \in \mathbb{R}^{n \times n}\), then one probability model might be
+to draw each element \(A_{ij}\), which represents the existence of an
+edge or the edge weight from vertex \(i\) to \(j\), independently from
+some distribution, perhaps with a unique parameter for the pair
+\((i,j)\), e.g., \(A_{ij} \indep F_{\theta_{ij}}\). The classical
+example of this is the \(\ER\) model \citep{Gilbert:1959}, in which
+every edge is drawn from the same distribution, typically a Bernoulli
+distribution, using the same parameter, i.e.,
+\(A_{ij} \iid \Bernoulli (p)\). The inhomogeneous Bernoulli graph
+extends this by allowing each edge, represented by \(A_{ij}\), to have
+its own parameter, \(P_{ij}\), e.g., in the Bernoulli case,
+\(A_{ij} \indep \Bernoulli(P_{ij})\). Typically, the parameters are
+collected into an edge probability matrix (in the case of unweighted
+graphs) or edge parameter matrix (in the case of weighted graphs),
+denoted as \(P \in \mathbb{R}^{n \times n}\). The network analysis
+problem in this setting is to estimate \(P\) given \(A\).
+
+If we let the parameter matrix \(P\) be unconstrained, the inference
+problem is overparameterized. On the other hand, the classical \(\ER\)
+model is often too restrictive to describe real, observed networks. Much
+work has been done to develop models that are constrained enough for
+robust statistical inference while being generalizable to describe a
+wide range of networks. One such family of random graph models is the
+random dot product graph (RDPG), first proposed by
+\citet{10.1007/978-3-540-77004-6_11}, which is a type of latent space
+graph in which each vertex of the graph has a corresponding latent
+vector in a low-dimensional Euclidean space \(\mathbb{R}^d\), and the
+edge parameter between each pair of vertices is determined by the dot
+product of the corresponding vectors. In this model, the constraint is
+the low rank of the parameter matrix \(P\), assuming that the latent
+dimension \(d\) is less than the number of vertices \(n\). Further
+constraints can be imposed on the RDPG in the form of distributional
+assumptions on the latent vectors or restricting the latent vectors to
+lie on subspaces or manifolds in the latent space
+\citep{athreya2020estimation}.
+
+It has been shown \citep{rubindelanchy2017statistical, Koo_2022} that
+the RDPG (as well as the generalized random dot product graph
+\citep{rubindelanchy2017statistical}) can describe a wide range of
+popular random graph models, such as the \(\ER\) model, the stochastic
+block model (SBM) \citep{doi:10.1080/0022250X.1971.9989788}, degree
+corrected block model (DCBM) \citep{Karrer_2011}, and popularity
+adjusted block model (PABM) \citep{307cbeb9b1be48299388437423d94bf1},
+and describing these models as RDPGs (or its generalized version) is
+useful for parameter estimation. The RDPG captures a wide range of
+phenomena that can be described as networks, and in fact any network
+that can be thought of as being sampled from a parameter matrix \(P\)
+can be described as a (generalized) RDPG, especially if \(P\) is low
+rank. The RPDG representation of these networks
+
 \hypertarget{methods}{%
 \section{Methods}\label{methods}}
+
+\hypertarget{definitions-and-models}{%
+\subsection{Definitions and Models}\label{definitions-and-models}}
+
+We begin by defining the RDPG and the LSM.
+
+\begin{definition}[Random dot product graph (RDPG) \citep{10.1007/978-3-540-77004-6_11}]
+\label{def:rdpg}
+Let $\mathcal{X}$ be a subset of $\mathbb{R}^d$ for some latent space dimension $d \geq 1$ such that for any $x_1, x_2 \in \mathcal{X}$, $x_1^\top x_2 \in [0, 1]$. 
+Let $F_\theta$ be a distribution with support $\mathcal{X}$ and parameters $\theta$, and sample $x_1, ..., x_n \iid F_\theta$. 
+A graph $G$ with adjacency matrix $A$ is a random dot product graph with latent vectors $X = \bigl[x_1 \; \cdots \; x_n\bigr]^\top$ drawn from distribution $F_\theta$ if $A \sim X X^\top$.
+
+We use the notation $A \sim \RDPG(F_\theta)$ to denote a random adjacency matrix $A$ drawn from latent vectors distributed as $F_\theta$. 
+\end{definition}
+
+\begin{remark}
+\label{remark:nonunique}
+The latent vectors of an RDPG are not unique. 
+Suppose that $P = X X^\top$ is the edge parameter matrix of an RDPG with latent positions $X$. 
+Then any orthogonal transformation $W$ on $X$ results in the same edge parameter matrix. 
+More precisely, let $\tilde{X} = X W$. 
+Then it is clear that $\tilde{X} \tilde{X}^\top = X W W^\top X^\top = X X^\top = P$ results in the same edge parameter matrix. 
+Thus, there are infinitely many latent vector configurations that can result in the same $P$, but for any two latent vector configurations, there exists an orthogonal mapping that connects the two. 
+\end{remark}
+
+\begin{definition}[Latent structure model (LSM) \citep{athreya2020estimation}]
+\label{def:lsm}
+Let $\mathcal{C} \subset \mathcal{X} \subset \mathbb{R}^d$ be a smooth, nonintersecting one-dimensional manifold on the domain of a RDPG as defined in definition \ref{def:rdpg}, parameterized by function $p(t) : [0, 1] \to \mathcal{C}$.
+Then if $t_1, ..., t_n \iid F_\theta$ for some distribution $F_\theta$ with support $[0, 1]$ and parameter $\theta$, each $x_i = p(t_i)$, and $ A \sim X X^\top$ for $X = \bigl[x_1 \; \cdots \; x_n\bigr]^\top$, $A$ is the adjacency matrix of a latent structure model on curve $\mathcal{C}$ with parameterization $p$ and underlying distribution $F_\theta$.
+
+We use the notation $A \sim \LSM(\mathcal{C}, F_\theta)$ or $A \sim \LSM(p, F_\theta)$ to denote an adjacency matrix $A$ drawn as an LSM on curve $\mathcal{C}$ or its parameterization $p$ with underlying distribution $F_\theta$.
+\end{definition}
+
+\begin{remark}
+Although \citet{athreya2020estimation} defined the LSM by a single one-dimensional manifold $\mathcal{C}$ in the latent space, in this paper, we will allow for the existence of multiple one-dimensional manifolds, $\mathcal{C}_1, ..., \mathcal{C}_K$, (i.e., mixture of manifolds distribution).
+This type of latent space mixture distribution is observed in networks with community structure (\citep{10.5555/3122009.3242083}). 
+For estimation, if the membership of each latent vector to the manifolds is known, then each manifold can be learned separately using the vectors that belong to that manifold. 
+If the memberships are not known, then we use an iterative algorithm to both cluster the latent vectors to a known number of manifolds and learn the manifolds using the cluster assignments.
+
+In the case of a mixture of $K$ curves, we use the notation $A \sim \LSM(\{\mathcal{C}_k\}_K, F_\theta, \alpha)$ or $A \sim \LSM(\{p_k\}_K, F_\theta, \alpha)$, where $\alpha = (\alpha_1, ..., \alpha_K)$, $\sum_{k=1}^K \alpha_k = 1$ is the mixture parameter.
+For simplicity, we only consider the case where the underlying distribution is the same for each curve. 
+\end{remark}
+
+A plausible inference task in the RDPG is to estimate the original
+latent vectors. The adjacency spectral embedding
+\citep{doi:10.1080/01621459.2012.699795} is a consistent estimator of
+the latent vectors, up to some unknown orthogonal transformation.
+
+\begin{remark}[Sparsity parameter]
+\end{remark}
+
+\begin{definition}[Adjacency spectral embedding (ASE) \citep{doi:10.1080/01621459.2012.699795}]
+\label{def:ase}
+\end{definition}
+
+\begin{theorem}[Consistency of the ASE \citep{lyzinski2014}]
+\label{theorem:ase-consistency}
+\end{theorem}
+
+Theorem \ref{theorem:ase-consistency} implies that the embedding vectors
+of the ASE converge to the original latent vectors, up to some
+unidentifiable orthogonal transformation, and the maximum deviation from
+the original latent vectors after the orthogonal transformation is
+bounded by a value that decays to \(0\). This implies that in the LSM,
+the ASE can lead to consistent estimation of \(F_\theta\), the
+underlying distribution, and in fact, \citet{athreya2020estimation}
+showed exactly this.
+
+\begin{definition}[Multiple latent structure model (MLSM)]
+\label{def:mlsm}
+Let $\mathcal{C}^{(1)}, ..., \mathcal{C}^{(L)} \subset \mathbb{R}^d$ be a sequence of curves defining the latent positions of a sequence of $L$ LSMs as in definition \ref{def:lsm}. 
+Each $\mathcal{C}^{(\ell)}$ is parameterized by function $p^{(\ell)}(t) : [0, 1] \to \mathcal{C}^{(\ell)}$. 
+Let $p^{(1)}, ..., p^{(L)}$ be a sequence of functions with domain $[0, 1]$ that parameterize the curves $\mathcal{C}^{(1)}, ..., \mathcal{C}^{(L)}$, let $F$ be a parametric distribution with support $[0, 1]$, and let $\theta_1, ..., \theta_L$ be a sequence of parameters for distribution $F$.
+For each $p^{(\ell)}$, sample $t_1^{(\ell)}, ..., t_{n_\ell}^{(\ell)} \iid F_{\theta^{(\ell)}}$ for some distribution $F$ parameterized by $\theta_\ell$, and let $x_i^{(\ell)} = p_\ell(t_i^{(\ell)})$ for each $i = 1, ..., n_\ell$, again as in definition \ref{def:lsm}.
+Sample a sequence of $L$ adjacency matrices, each as $A^{(\ell)} \indep \LSM(p_\ell, F_{\theta_\ell})$. 
+Then the sequence $\{A^{(\ell)}\}_L$ is are the adjacency matrices of a multiple latent structure model with curves $\{C^{(\ell)}\}_L$ parameterized by $\{p^{(\ell)}\}$ and underlying distribution $F$ with parameters $\{\theta_\ell\}_L$.
+
+We use the notation $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$ to denote a sequence of adjacency matrices drawn from an MLSM with parameterizations $\{p^{(\ell)}\}$ and parameters $\{\theta_\ell\}$ on underlying distrubtion $F$.
+\end{definition}
+
+\begin{remark}
+\label{remark:mlsm-mixture}
+Again as in the case of a single LSM, we also allow for each $A^{(\ell)}$ to be sampled from a latent structure composed of $K$ curves with mixture parameter $\alpha^{(\ell)}$. 
+In this case, we use the notation $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{\mathcal{C}_k^{(\ell)}\}_{K, L}, F, \{\theta_\ell\}_L, \{\alpha^{(\ell)}\}_L)$ or $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p_k^{(\ell)}\}_{K, L}, F, \{\theta_\ell\}_L, \{\alpha^{(\ell)}\}_L)$.
+\end{remark}
+
+\hypertarget{connections-to-related-models}{%
+\subsubsection{Connections to Related
+Models}\label{connections-to-related-models}}
+
+In the MLSM defined in definition \ref{def:mlsm}, the only commonality
+that we assume from graph to graph is that they are all LSMs with the
+same underlying distribution family. If we impose further restrictions,
+namely that the latent structures are linear, we obtain the multilayer
+random dot product graph \citep{jones2021multilayer}.
+
+\begin{example}[Comparison to the multilayer DCBM \citep{agterberg2022joint}, multi-RDPG \citep{nielsen2018multiple}, and MREG \citep{8889404}]
+In the $K$-community DCBM, the probability of an edge between a pair of vertices is given by
+
+$$A_{ij} \indep \Bernoulli(\omega_i \omega_j B_{z_i, z_j}),$$
+
+where $z_i \in \{1, ..., K\}$ is the community label for vertex $i$, $B_{k, \ell}$ is the block connectivity between communities $k$ and $\ell$, and $\omega_i$ is the degree correction parameter for vertex $i$.
+In order to preserve identifiability and uniqueness, a common constraint on these parameters is to set $\sum_{i : z_i = k} \omega_i^2 = 1$ \citep{Karrer_2011}. 
+As in
+
+The edge parameter matrix of a $K$-community DCBM with $n$ vertices can be decomposed as follows:
+
+$$P = \Omega B \Omega^\top,$$
+
+where $B$ is a $K \times K$ matrix of block connectivities and $\Omega$ is an $n \times K$ matrix such that $\Omega_{ik} = \omega_i$ if vertex $i$ is in community $k$ and $0$ otherwise. 
+Then it is clear that if $B$ is positive semidefinite matrix of rank $K$, $P$ can also be viewed as a $K$-dimensional RDPG, since the rows of $\Omega$ are normalized and can be seen as eigenvectors, and $B$ is full rank and can be decomposed into a diagonal matrix and a rotation matrix, i.e., $P = (\Omega V) \Lambda (\Omega V)^\top$.
+Furthermore, this matrix decomposition implies that the latent vectors lie on one of $K$ line segments that intersect at the origin \citep{rubindelanchy2017statistical}. 
+Thus, the DCBM is a special case of the LSM in which there are $K$ latent ``linear curves'' (i.e., lines) in $\mathbb{R}^K$.
+
+The multilayer DCBM as described by \citet{agterberg2022joint} extends this to a sequence of DCBMs with the same community structure, allowing the block connectivities and degree correction parameters to change for each layer but keeping the same community structure throughout, i.e., each element of $P^{(\ell)} = \Omega^{(\ell)} B^{(\ell)} (\Omega^{(\ell)})^\top$ can vary with $\ell$ but $\Omega^{(\ell)} = 0 \iff \Omega^{(\ell')} = 0$ and similarly, $\Omega^{(\ell)} > 0 \iff \Omega^{(\ell')} > 0$, for every pair $(\ell, \ell')$.
+On the other hand, if the further restriction of $\Omega^{(\ell)} = \Omega^{(\ell')}$ for all $(\ell, \ell')$, i.e., $P^{(\ell)} = \Omega B^{(\ell)} \Omega^\top$, we obtain a special case of the multiple RDPG (multi-RDPG) with the identity link function, as described by \citet{nielsen2018multiple}, or equivalently, a special case of the multiple random eigen graphs model (MREG), as described by \citet{8889404}.
+However, if the community labels are not identical from graph to graph, the sequence of DCBMs cannot be described as a multilayer DCBM, multi-RDPG, or MREG, but it can still be described as an MLSM. 
+\end{example}
+
+\begin{example}[Nonlinear MLSM and comparison to MREG and multi-RDPG]
+\end{example}
+
+\hypertarget{classification-and-regression-on-the-mlsm}{%
+\subsubsection{Classification and Regression on the
+MLSM}\label{classification-and-regression-on-the-mlsm}}
+
+To use the MLSM for classification or regression problems, we assign
+response variables \(y_1, ..., y_L\) to each graph. Then if each
+response \(y_\ell\) depends on \(p^{(\ell)}(t)\) (the parameterization
+of the \(\ell^{th}\) curve) or \(\theta_\ell\) (the parameter of the
+\(\ell^{th}\) distribution), or some combination of the two, there is a
+plausible setup for a predictive modeling task for predicting \(y_\ell\)
+after observing \(A^{(\ell)}\). Four such scenarios are given:
+
+\begin{definition}[MLSM for classification 1]
+\label{def:mlsm-c-1}
+Let $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$, $y_1, ..., y_L \in \{1, ..., M\}$ be labels for each of the $L$ graphs of the MLSM, and each $\theta_\ell$ take on one of $M$ values $\{\phi_1, ..., \phi_M\}$ corresponding to its response, $y_\ell$, i.e., $\theta_\ell = \phi_{y_\ell}$.
+
+We observe the adjacency matrices $A^{(1)}, ..., A^{(L)}$ and the first $r$ labels $y_1, ..., y_r$, and the task is to predict the unknown labels $y_{r+1}, ..., y_L$.
+
+This can be described by the following generative model:
+
+\begin{enumerate}
+  \item Draw labels $y_1, ..., y_L \iid \Categorical(\{1, ..., M\}, \{\pi_1, ..., \pi_M\})$.
+  \item Set each $\theta_\ell = \phi_{y_\ell}$.
+  \item Draw adjacency matrices as $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$.
+\end{enumerate}
+\end{definition}
+
+\begin{definition}[MLSM for classification 2]
+\label{def:mlsm-c-2}
+Let $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{\mathcal{C}^{(\ell)}\}, F, \{\theta_\ell\}_L)$, $y_1, ..., y_L \in \{1, ..., M\}$ be labels for each of the $L$ graphs of the MLSM, and each $\mathcal{C}^{(\ell)}$ take on one of $M$ functional forms $\{p^{(1)}(t), ..., p^{(M)}(t)\}$ corresponding to its response variable, $y_\ell$, i.e., $\mathcal{C}^{(\ell)}$ is parameterized by $p^{(y_\ell)}(t)$.
+
+We observe the adjacency matrices $A^{(1)}, ..., A^{(L)}$ and the first $r$ labels $y_1, ..., y_r$, and the task is to predict the unknown labels $y_{r+1}, ..., y_L$.
+
+This can be described by the following generative model:
+
+\begin{enumerate}
+  \item Draw labels $y_1, ..., y_L \iid \Categorical(\{1, ..., M\}, \{\pi_1, ..., \pi_M\})$.
+  \item Set each $p^{(\ell)}(t) = p^{(y_\ell)}(t)$.
+  \item Draw adjacency matrices as $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$.
+\end{enumerate}
+
+As in remark \ref{remark:mlsm-mixture}, we also allow for each adjacency matrix $A^{(\ell)}$ to be drawn from a latent structure consisting of multiple curves $\{p^{(\ell)}_k\}_K$. 
+\end{definition}
+
+\begin{definition}[MLSM for regression 1]
+\label{def:mlsm-r-1}
+Let $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$, and suppose that for each $\ell = 1, ..., L$, the $\ell^{th}$ graph is coupled with a response variable, $y_\ell$, as $y_\ell \indep \mathcal{N}(\theta_\ell^\top \beta, \sigma^2)$.
+
+We observe the adjacency matrices $A^{(1)}, ..., A^{(L)}$ and the first $r$ response variables $y_1, ..., y_r$. 
+In this setting, there are two plausible inference tasks. 
+The first is to estimate the coefficient vector $\beta$. 
+The second is to predict the unobserved response variables $y_{r+1}, ..., y_L$. 
+\end{definition}
+
+\begin{definition}[MLSM for regression 2]
+\label{def:mlsm-r-2}
+Let $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$ and each $p^{(\ell)}(t) = p(t; \gamma_\ell)$, where $\gamma_\ell$ is the vector of parameters for function $p$. 
+Suppose that for each $\ell = 1, ..., L$, the $\ell^{th}$ graph is coupled with a response variable, $y_\ell$, as $y_\ell \indep \mathcal{N}(\gamma_\ell^\top \beta, \sigma^2)$.
+
+We observe the adjacency matrices $A^{(1)}, ..., A^{(L)}$ and the first $r$ response variables $y_1, ..., y_r$. 
+In this setting, there are two plausible inference tasks. 
+The first is to estimate the coefficient vector $\beta$. 
+The second is to predict the unobserved response variables $y_{r+1}, ..., y_L$.
+
+As in remark \ref{remark:mlsm-mixture}, we also allow for each adjacency matrix $A^{(\ell)}$ to be drawn from a latent structure consisting of multiple curves $\{p^{(\ell)}_k\}_K$. 
+\end{definition}
+
+In all of these settings, the ASE of \(A^{(\ell)}\) provides some
+insight into both \(p^{(\ell)}\) and \(\theta_\ell\).
+\citet{athreya2020estimation} showed that with some additional
+information, the ASE of \(A^{(\ell)}\) can lead to a consistent
+estimator for \(\theta_\ell\).
+
+\hypertarget{main-results}{%
+\subsection{Main Results}\label{main-results}}
 
@@
 In this simulation experiment, the latent vectors were sampled along a
-Bezier curve defined by
-\(g(t) = \begin{bmatrix} t^2 & 2 t (1-t) \end{bmatrix}^\top\). The
-timepoints \(t_i\) were sampled as iid Beta random variables with two
-sets of parameters, \((\alpha_1, \beta_1)\) and \((\alpha_2, \beta_2)\).
-The setup is as follows:
+Bezier curve defined by \(g(t) = \bigl[ t^2 \; 2 t (1-t) \bigr]^\top\).
+The timepoints \(t_i\) were sampled as iid Beta random variables with
+two sets of parameters, \((\alpha_1, \beta_1)\) and
+\((\alpha_2, \beta_2)\). The setup is as follows:
 
@@
   Construct each latent vector as \(x_i = g(t_i)\) and compile them in
-  data matrix
-  \(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\).
+  data matrix \(X = \bigl[ x_1 \; \cdots \; x_n \bigr]^\top\).
 \item
@@
 for testing. We investigated graphs of size \(|V| = 32, 64, 128, 256\).
-The number of graphs for each experiment was set to \(n = 64\). For each
+The number of graphs for each experiment was set to \(L = 64\). For each
 (number of graphs, size of graph) pair, we performed 32 replicates.
@@
 
-\begin{center}\includegraphics{draft_files/figure-latex/class-sim-ase-1} \end{center}
+\begin{figure}[H]
 
-\begin{center}\includegraphics{draft_files/figure-latex/classification-sim-1} \end{center}
+{\centering \includegraphics{draft_files/figure-latex/class-sim-ase-1}
+
+}
+
+\caption{One simulated graph (left) and its ASE (right). The red curve is the fitted quadratic Bezier curve on the ASE.}\label{fig:class-sim-ase}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/classification-sim-1}
+
+}
+
+\caption{Median classification error rate and its IQR vs. number of vertices in each graph.}\label{fig:classification-sim}
+\end{figure}
 
@@
     For each \(i = 1, ..., n\), set
-    \(x_i = \begin{cases} \begin{bmatrix} t_i & 0 \end{bmatrix}^\top & z_i = 1 \\ \begin{bmatrix} t_i \cos \theta_k & t_i \sin \theta_k \end{bmatrix}^\top & z_i = 2 \end{cases}\);
+    \(x_i = \begin{cases} \bigl[ t_i \; 0 \bigr]^\top \; z_i = 1 \\ \bigl[ t_i \cos \theta_\ell \; t_i \sin \theta_\ell \bigr]^\top \; z_i = 2 \end{cases}\);
   \item
-    Collect
-    \(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\) and
-    draw \(A \sim \RDPG(X)\);
+    Collect \(X = \bigl[ x_1 \; \cdots \; x_n \bigr]^\top\) and draw
+    \(A \sim \RDPG(X)\);
   \item
-    Set the response \(y_k = \beta_0 + \beta_1 \theta_k\).
+    Set the response \(y_\ell = \beta_0 + \beta_1 \theta_\ell\)
   \end{enumerate}
@@
 In this simulation, we set \(\beta_0 = \beta_1 = 1\). The number of
-graphs was set to \(N = 64\), and the number of vertices per graph was
+graphs was set to \(L = 64\), and the number of vertices per graph was
 set to \(n = 128, 256, 512, 1024\). For each \(n\), we simulated 32
@@
 \section{Applications}\label{applications}}
+
+\hypertarget{human-connectome-project-aging-study}{%
+\subsection{Human Connectome Project Aging
+Study}\label{human-connectome-project-aging-study}}
 
@@
 
-In this dataset, there are \(N = 136\) graphs (corresponding to
-individual subjects), each with \(n = 68\) vertices (corresponding to
+In this dataset, there are \(N = 516\) graphs (corresponding to
+individual subjects), each with \(n = 84\) vertices (corresponding to
 brain regions). Analyzing these graphs as RDPGs reveals that the DCBM is
-a good candidate for these data (fig \ref{fig:hcp-ase}).
+a good candidate for these data (figure \ref{fig:hcp-ase}).
 
@@
 
-{\centering \includegraphics{draft_files/figure-latex/hcp-ase-1}
+{\centering \includegraphics[width=1\linewidth]{draft_files/figure-latex/hcp-ase-1}
 
@@
 
-\caption{ASE of one graph from the HCP dataset.
-The lines are fitted via $K$-curves clustering using degree = 1.
+\caption{One graph from the HCP dataset (left) and its ASE (right).
+In the ASE, the lines are fitted via $K$-curves clustering using degree = 1.
 The outputted clusters correspond exactly to the left (red) and right (blue) hemispheres.}\label{fig:hcp-ase}
@@
 
-For each graph, the
+\begin{figure}[H]
 
-\begin{center}\includegraphics{draft_files/figure-latex/hcp-gender-1} \end{center}
+{\centering \includegraphics{draft_files/figure-latex/hcp-age-regression-1}
 
-\begin{center}\includegraphics{draft_files/figure-latex/hcp-age-1} \end{center}
+}
+
+\caption{Scatterplot between the subject age (in months) vs. the fitted angle between hemispheres of the brain in the latent space.}\label{fig:hcp-age-regression}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\linewidth]{draft_files/figure-latex/hcp-age-regression-degree-1}
+
+}
+
+\caption{Scatterplot between the subject age (in months) vs. the average degree within (left) and between (right) hemispheres.}\label{fig:hcp-age-regression-degree}
+\end{figure}
+
+The ASE suggests a latent structure comprised of two line segments, one
+for each hemisphere of the brain, that meet at the origin. One possible
+parameter is the angle between the two lines. The estimated angles were
+observed to correlate with the subject's age, with wider angles
+corresponding to older subjects (figure \ref{fig:hcp-age-regression}). A
+linear regression setting aside half of the brain connectivity graphs as
+test data achieves an RMSE of \(11.89\) months.
+
+Other plausible statistics from the graphs are the average degrees
+within and between hemispheres. When analyzing these as stochastic block
+models (SBM) or degree corrected block models (DCBM), these statistics
+would roughly correspond to the expected edge weights within and between
+communities. Both of these statistics are correlated with the target
+value (figure \ref{fig:hcp-age-regression-degree}), but a linear model
+using them (i.e., \(age \sim degree_{within} + degree_{between}\))
+achieves a higher RMSE (\(12.3\) months) while adding another parameter
+to the model. Similarly, the assortativity of each graph with respect to
+hemisphere results in a higher RMSE of \(156.5\) when regressed on age.
+
+We also note that this structure is consistent with the DCBM. The
+multilayer DCBM was previously studied by \citet{agterberg2022joint}
+
+\begin{center}\includegraphics[width=1\linewidth]{draft_files/figure-latex/cosine-heatmap-1} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{draft_files/figure-latex/cosine-heatmap-2-1} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{draft_files/figure-latex/cosine-heatmap-2-2} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{draft_files/figure-latex/cosine-heatmap-2-3} \end{center}
+
+\begin{center}\includegraphics[width=1\linewidth]{draft_files/figure-latex/cosine-heatmap-2-4} \end{center}
+
+\begin{longtable}[]{@{}lr@{}}
+\caption{Correlation between age and various graph
+metrics}\tabularnewline
+\toprule\noalign{}
+Metric & Correlation \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+Metric & Correlation \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Angle between hemispheres & 0.558 \\
+Degree within hemisphere & 0.436 \\
+Degree between hemispheres & -0.499 \\
+Number of triangles in left hemisphere & -0.025 \\
+Number of triangles in right hemisphere & 0.002 \\
+Assortativity w.r.t. hemisphere & 0.436 \\
+Transitivity & -0.261 \\
+Modularity w.r.t. hemisphere & 0.434 \\
+\end{longtable}
 
@@
 
+\newpage
+
+\bibliographystyle{apalike}
 \renewcommand\refname{Bibliography}

### 2023-11-14 hcp-a-eda.R

@@ -1,0 +1,284 @@
+import::from(magrittr, `%>%`, `%<>%`)
+import::from(foreach, foreach, `%do%`, `%dopar%`)
+source('~/dev/pabm-grdpg/functions.R')
+source('~/dev/manifold-block-models/functions.R')
+library(mclust)
+library(ggplot2)
+
+doMC::registerDoMC(parallel::detectCores())
+
+load('~/Downloads/Data/HCP/HCP-A/DTI/Data.RData')
+
+subject.df <- demo.data %>% 
+  dplyr::filter(QC_Issue_Codes == '',
+                Full_MR_Compl == 1) %>% 
+  dplyr::transmute(id = src_subject_id,
+                   sex = sex,
+                   age = as.numeric(interview_age)) %>% 
+  dplyr::filter(age < 1200) %>% 
+  dplyr::mutate(age = age / 12)
+
+z <- dti.info$hemisphere %>% 
+  as.factor()
+
+rois <- dti.info$aparc_aseg_region.names.short
+
+A <- dat.NumFibers[[2]] %>% 
+  as.matrix()
+A <- (A + t(A)) / 2
+diag(A) <- 0
+
+qgraph::qgraph(A, layout = 'spring', groups = dti.info$hemisphere)
+Xhat <- embedding(A, 2, 0, scale = TRUE)
+clustering <- manifold.clustering(Xhat, 
+                                  degree = 1,
+                                  initialization = as.numeric(z),
+                                  normalize = TRUE,
+                                  intercept = FALSE,
+                                  parallel = FALSE)
+plot.estimated.curves(Xhat, clustering) + 
+  ggrepel::geom_label_repel(aes(x = Xhat[, 1],
+                                y = Xhat[, 2],
+                                label = dti.info$aparc_aseg_region.names.short))
+params.l <- EnvStats::ebeta(clustering$t[[1]])$parameters
+params.r <- EnvStats::ebeta(clustering$t[[2]])$parameters
+
+A.within.l <- A[z == 'L', z == 'L']
+A.within.r <- A[z == 'R', z == 'R']
+deg.within <- mean(c(A.within.l[upper.tri(A.within.l)],
+                     A.within.r[upper.tri(A.within.r)]))
+A.between <- A[z == 'L', z == 'R']
[…]

### 2023-11-14 hcp-eda.R

@@ -17,81 +17,114 @@ betadist.mom <- function(x) {
               beta = beta))
 }
 
-data.dir <- '~/Downloads/repeated_10_scale_125/'
-graph.list <- lapply(dir(data.dir), function(x) {
-  graph <- igraph::read.graph(file.path(data.dir, x), format = 'graphml')
-  A <- graph %>% 
-    igraph::as_adjacency_matrix() %>% 
-    as.matrix()
-  z <- graph %>% 
-    igraph::get.vertex.attribute() %>% 
-    {.$dn_hemisphere} %>% 
-    as.factor()
-  return(list(A = A, z = z, id = substr(x, 1, 6)))
-})
+extract.t.norm <- function(roi, A, z, p, t) {
+  roi.l <- paste(roi, 'L', sep = '_')
+  roi.r <- paste(roi, 'R', sep = '_')
+  
+  idx.l <- which(colnames(A)[z == 1] == roi.l)
+  idx.r <- which(colnames(A)[z == 2] == roi.r)
+  
+  t.l <- t[[1]][idx.l]
+  t.r <- t[[2]][idx.r]
+  
+  norm.l <- (p[[1]] * t.l) %>% 
+    magrittr::raise_to_power(2) %>% 
+    sum() %>% 
+    sqrt()
+  norm.r <- (p[[2]] * t.r) %>% 
+    magrittr::raise_to_power(2) %>% 
+    sum() %>% 
+    sqrt()
+  
+  return(list(t = list(l = t.l,
+                       r = t.r),
+              norm = list(l = norm.l,
+                          r = norm.r)))
+}
+
+# data.dir <- '~/Downloads/repeated_10_scale_125/'
+# graph.list <- lapply(dir(data.dir), function(x) {
+#   graph <- igraph::read.graph(file.path(data.dir, x), format = 'graphml')
+#   A <- graph %>% 
+#     igraph::as_adjacency_matrix() %>% 
+#     as.matrix()
+#   z <- graph %>% 
+#     igraph::get.vertex.attribute() %>% 
+#     {.$dn_hemisphere} %>% 
+#     as.factor()
+#   return(list(A = A, z = z, id = substr(x, 1, 6)))
+# })
 
 load('~/Downloads/HCP/HCP-Y/DTI_fMRI/Data.RData')
 
-graph.list <- lapply(graph.list, function(graph) {
-  if (graph$id %in% as.character(dat.demo$Subject)) {
-    return(graph)
-  } else {
[…]

### 2023-11-14 simulations/angle-regression-sim.R

@@ -36,6 +36,7 @@ out.df <- foreach::foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
       clustering <- manifold.clustering(Xhat, 
                                         degree = 1,
                                         initialization = z,
+                                        normalize = TRUE,
                                         intercept = FALSE)
       angle.1 <- atan2(clustering$p[[1]][2, 2],
                        clustering$p[[1]][2, 1])

### 2023-11-17 draft.Rmd
removed large file

@@
 Then we propose various estimation techniques involving manifold learning to estimate these parameters and in turn predict the responses, with theorems guaranteeing convergence of the predictions.
-Simulations and applications on functional MRI data verify the performance of these methods.
+Simulations, as well as applications on brain connectivity networks, verify the performance of our methods.
 ---
@@
 source('~/dev/manifold-block-models/functions.R')
+
+doMC::registerDoMC(parallel::detectCores())
 ```
@@
 Thus, there are infinitely many latent vector configurations that can result in the same $P$, but for any two latent vector configurations, there exists an orthogonal mapping that connects the two. 
+Similarly, if $A \sim \RDPG(F_{\mu, \Sigma})$ and $F_{\mu, \Sigma}$ is the normal distribution with mean vector $\mu$ and covariance matrix $\Sigma$, then $A \sim \RDPG(F_{W \mu, W \Sigma W^\top})$ is an equivalent RDPG for any orthogonal matrix $W$.
 \end{remark}
@@
 \begin{remark}
+\label{rem:lsm-mixture}
 Although \citet{athreya2020estimation} defined the LSM by a single one-dimensional manifold $\mathcal{C}$ in the latent space, in this paper, we will allow for the existence of multiple one-dimensional manifolds, $\mathcal{C}_1, ..., \mathcal{C}_K$, (i.e., mixture of manifolds distribution).
-This type of latent space mixture distribution is observed in networks with community structure (\citep{10.5555/3122009.3242083}). 
+This type of latent space mixture distribution is observed in networks with community structure \citep{10.5555/3122009.3242083}. 
 For estimation, if the membership of each latent vector to the manifolds is known, then each manifold can be learned separately using the vectors that belong to that manifold. 
@@
 \begin{remark}[Sparsity parameter]
+In many real networks, the degree of each vertex often does not grow proportionally with the size of the network. 
+To account for this, a sparsity factor $\rho_n \in (0, 1]$ is introduced in the edge probabilities, i.e., $P_{ij} \leftarrow \rho_n P_{ij}$, for some sequence $\{\rho_n\}$. 
+Oftentimes the additional constraint of $\lim\limits_{n \to \infty} \rho_n = 0$ is included. 
+For example, a sparse SBM has edge probabilities $P_{ij} = \rho_n \theta_{z_i, z_j}$, for which we use the notation $A \sim \SBM(z, \{\theta_{k \ell}\}_K; \rho_n)$ or $A \sim \SBM(\alpha, \{\theta_{k \ell}\}_K; \rho_n)$, depending on whether we treat the labels as random or fixed.
+Then the expected degree grows as $O(n \rho_n)$ instead of linearly as $O(n)$. 
+For the sake of unifying the sparse and dense regimes, we also allow for the special case $\rho_n = 1$ and include the sparsity factor throughout, unless otherwise stated. 
+Finally, we also note that while $\rho_n$ limits the rate of growth of the expected degree, our theoretical results still require $n \rho_n$ to diverge to infinity, albeit at a slower rate than $O(n)$.
+For example, if $\rho_n \propto 1 / \sqrt{n}$, then the expected degree for each vertex of the SBM is $O(\sqrt{n})$. 
+In general, for consistency in estimation and inference, most results on Bernoulli random graphs require $n \rho_n = \omega \big( (\log n)^c \big)$ for some $c > 1$. 
+(See \citet{JMLR:v18:16-480}, \citet{https://doi.org/10.48550/arxiv.2106.09840}, and \citet{rubindelanchy2017statistical} for further discussion.)
 \end{remark}
@@
 \label{def:ase}
+Let $A = V \Lambda V^\top$ be the spectral decomposition of $A$. 
+Define $\lambda_i$ as the $i^{th}$ largest eigenvalue of $A$ and $v_i$ by its corresponding eigenvector, and let $V_d = \bigl[ v_1 \; \cdots \; v_d \bigr]$ and $\Lambda_d = \diag(\lambda_1, ..., \lambda_d)$.
+Then $\hat{X} = V_d |\Lambda|^{1/2}$ is the $d$-dimensional adjacency spectral embedding of $A$. 
 \end{definition}
@@
 \label{theorem:ase-consistency}
+Suppose the sparsity parameter is such that $n \rho_n = \omega (\log^{4 c} n)$ for some constant $c > 1$. 
+Then for some orthogonal transformation $W \in \mathbb{O}(d)$, 
+$$\max_i \|W \hat{x}_i - x_i \| = O_P \bigg( \frac{\log^c n}{n^{1/2}} \bigg),$$
+where $x_i^\top$ and $\hat{x}_i^\top$ are the rows of $X$ and $\hat{X}$, respectively. 
 \end{theorem}
@@
 
+We propose our methodology for classification and regression in three steps:
+
+1. Curve-fitting algorithm to recover the LSM curves from the ASEs.
+2. Quasi-maximum likelihood estimation to recover the parameters of the underlying distributions. 
+3. Classification or regression to determine the relationship between the parameters and the responses.
+
+### Estimation of the LSM Curves
+
+For curve-fitting, we will restrict the LSM curves to non-self-intersecting Bezier polynomials \citep{10.5555/320367}. 
+A polynomial curve of degree $R$ in $\mathbb{R}^d$ is uniquely defined by a Bezier polynomial within the space of Bezier polynomials of degree at most $R$, up to reverse order, although there are infinitely many Bezier polynomials of degree greater than $R$ that can describe the curve \citep{SANCHEZREYES2022102118}.
+We further restrict the curves to the case $p(0) = 0$, i.e., the curve begins at the origin. 
+This is consistent with well-known models such as the DCBM and PABM when viewed as LSMs or mixtures of LSMs \citep{Koo_2022, rubindelanchy2017statistical}, as well as real networks that can adequately be described as LSMs \citep{athreya2020estimation}.
+This constraint also removes the reverse order nonidentifiability. 
+Thus, if $R$, the degree of the polynomial, can be determined, and the curve begins at the origin, a collection of $n \geq R$ unique vectors on the curve and not at the origin can determine the unique Bezier polynomial that describes the curve.
+
+A Bezier polynomial of this form, is defined as 
+\begin{equation}
+\label{eq:bezier}
+p(t; b_1, ..., b_R) = \sum_{r=1}^R \binom{R}{r} b_r (1-t)^r t^{R-r},
+\end{equation}
+where $b_r \in \mathbb{R}^d$ are Bezier coefficients (control points). 
+In the scenario in which an adjacency matrix $A$ sampled from an LSM with curve $p$ that can be described by a Bezier polynomial, the ASE of $A$ consists of embedding vectors that lie on or near a rotation of the curve.
+Then if embedding vectors $x_1, ..., x_n \in \mathbb{R}^d$ are such that $x_i = p(t_i) + \epsilon_i$ (for simplicity, redefine $p$ to be the rotated version of the original curve), the squared loss function for the curve fit is of the form
+\begin{equation}
+\label{eq:bezier-loss-full}
+L(t_1, ..., t_n; b_1, ..., b_R) = \sum_{i=1}^n \big( x_i - p(t_i; b_1, ..., b_R) \big)^2.
+\end{equation}
+
+Letting $X = \bigl[ x_1 \; \cdots \; x_n \bigr]^\top \in \mathbb{R}^{n \times p}$, $T \in \mathbb{R}^{n \times R}$ such that $T_{ir} = \binom{R}{r} (1-t_i)^{R-r} t_i^r$, $b = \bigl[b_1 \; \cdots \; b_R\bigr]^\top \in \mathbb{R}^{R \times d}$, the loss function can be rewritten as
+\begin{equation}
+\label{eq:bezier-loss}
+L(T, b) = \|X - T b\|_F^2,
+\end{equation}
+where $\|\cdot\|_F$ is the Frobenius norm. Then if $t_1, ..., t_n$ are known, the solution to the least squares best fit Bezier coefficients is given by the ordinary least squares estimate, 
+\begin{equation}
+\label{eq:bezier-ols}
+\hat{b} = (T^\top T)^{-1} T^\top X.
+\end{equation}
+
+On the other hand, if the coefficients are known but the timepoints $t_1, ..., t_n$ are not, this turns into $n$ individual minimization problems for polynomials of degree $2 R$. 
+More precisely, $L(t_1, ..., t_n; b_1, ..., b_R) = \sum_{i=1}^n L(t_i; b_1, ..., b_R)$ where 
+\begin{equation}
+\label{eq:bezier-ls}
+L(t_i; b_1, ..., b_R) = x_i - \sum_r \binom{R}{r} b_r (1-t_i)^{R-r} t_i^r. 
+\end{equation}
+To solve, we find the at most $2 R - 1$ roots of its derivative, 
+\begin{equation}
+\label{eq:dbezier}
+\dot{L}(t_i; b_1, ..., b_R) = \bigg(\sum_{r=1}^R \binom{R}{r} (-1)^r c_r t_i^r \bigg) \bigg(\sum_{r=0}^{R-1} \binom{R-1}{r} (-1)^r c_{r+1} t_i^r \bigg) = 0,
+\end{equation}
+where $c_r = \sum_{s=1}^r (-1)^{r-s} \binom{r}{s} b_s$. 
+Since each $t_i$ can be solved separately, this method is highly parallelizable.
+
+For initialization, the ordering of $t_1, ..., t_n$ is determined via a one-dimensional Isomap embedding \citep{Tenenbaum2000-ff}.
+
+Combining equation \ref{eq:bezier-ols} with the solutions to equation \ref{eq:dbezier} for each $i = 1, ..., n$ provides an alternating coordinate descent algorithm, as outlined in algorithm \ref{alg:bezier-fit}.
+
+\begin{algorithm}[h]
+\label{alg:bezier-fit}
+\DontPrintSemicolon
+\SetAlgoLined
+\setstretch{1.1}
+\KwData{Adjacency matrix $A$, embedding dimension $d$, polynomial degree $R$, Isomap neighborhood parameter $\lambda$, stopping criterion $\epsilon$.}
+\KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$.}
+Compute $\hat{X} \in \mathbb{R}^{n \times d}$, the ASE of $A$.\;
+Initialize timepoints $\hat{t}_1, ..., \hat{t}_n$, first via a one-dimensional Isomap embedding of $\hat{X}$ using neighborhood parameter $\lambda$, then by normalizing the embedding to $[0, 1]$.\;
+\Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
+  Fit $\hat{b}$ by equation \ref{eq:bezier-ols}.\;
+  \For {$i = 1, ..., n$} {
+    Fit $\hat{t}_i$ by equation by finding the roots of equation \ref{eq:dbezier} and choosing the roots that minimizes equation \ref{eq:bezier-ls}.\;
+  }
+}
+\caption{Procedure for estimating an LSM curve as a Bezier polynomial from an adjacency matrix.}
+\end{algorithm}
+
+In the case of a mixture of LSMs (as in remark \ref{rem:lsm-mixture}), algorithm \ref{alg:bezier-fit} is modified by fitting Bezier polynomials for each vertex label (if the labels are known) or by estimating the labels via an alternating coordinate descent algorithm outlined in algorithm \ref{alg:bezier-fit-mixture}.
+
+\begin{algorithm}[H]
+\label{alg:bezier-fit-mixture}
+\DontPrintSemicolon
+\SetAlgoLined
+\setstretch{1.1}
+\KwData{Adjacency matrix $A$, embedding dimension $d$, polynomial degree $R$, number of latent structures $K$, Isomap neighborhood parameter $\lambda$, stopping criterion $\epsilon$.}
+\KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$, curve memberships $\hat{z}_1, ..., \hat{z}_n$.}
+Compute $\hat{X} \in \mathbb{R}^{n \times d}$, the ASE of $A$.\;
+Initialize $\hat{z}_1, ..., \hat{z}_n \in \{1, ..., K\}$, the membership of each vertex to each latent curve.\;
+\For {$k = 1, ..., K$} {
+  Subset the rows of $\hat{X}$ by $z_i = k$ to obtain $\hat{x}_{k_1}, ..., \hat{x}_{k_{n_k}}$, where $n_k$ is the number of vertices with label $k$.\;
+  Initialize timepoints $\hat{t}_{k_1}, ..., \hat{t}_{k_{n_k}}$, first via a one-dimensional Isomap embedding of $\hat{x}_{k_1}, ..., \hat{x}_{k_{n_k}}$ using neighborhood parameter $\lambda$, then by normalizing the embedding to $[0, 1]$.\;
+}
+\Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
+  \For {$k = 1, ..., K$} {
+    Fit $\hat{b}^{(k)}$ by equation \ref{eq:bezier-ols}, using the row subsets of $T$ and $X$ such that $z_i = k$.\;
+    \For {$i = 1, ..., n_k$} {
+      Fit $\hat{t}_{k_i}$ by equation by finding the roots of equation \ref{eq:dbezier} and choosing the roots that minimizes equation \ref{eq:bezier-ls}.\;
+    }
+  }
+  \For {$k = 1, ..., n$} {
+    Reassign $z_i$ based on the latent curve closest to $\hat{x}_i$. 
+  }
+}
+\caption{Procedure for estimating multiple LSM curves as a Bezier polynomial from an adjacency matrix.}
+\end{algorithm}
+
+\begin{lemma}[Recovery of LSM curves in noiseless case]
+\label{lem:lsm-noiseless}
+Suppose $t_1, ..., t_n \iid F_\theta$ such that $F_\theta$ has probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
+Let each $x_i = p(t_i)$ such that $p: [0, 1] \to \mathbb{R}^d$ is a non-self-intersecting Bezier polynomial of degree $R$ that begins at the origin with unknown coefficients $b_1, ..., b_R$. 
+Then there exists a sequence of Isomap neighborhood parameters $\lambda_n \to 0$ such that as $n \to \infty$, algorithm \ref{alg:bezier-fit}, $L(t_1, ..., t_n; b_1, ..., b_R) \stackrel{p}{\to} 0$.
+\end{lemma}
+
+\begin{proof}
+Let $z_1, ..., z_n$ be the one-dimensional Isomap embedding of $x_1, ..., x_n$ using neighborhood parameter $\lambda_n$, and let $l_{ij}$ be the arc length along the curve defined by $p$ between $x_i$ and $x_j$.
+Then by theorem 5 of \citet{acharyya2023}, $|z_i - z_j| - l_{ij} \stackrel{p}{\to} 0$ $\forall i, j \in \{1, ...,n\}$. 
+Let $z_{\min} = \min_i z_i$ and $z_{\max} = \max_i z_i$, and initialize each $\hat{t}_i = (z_i - z_{\min}) / (z_{\max} - z_{\min})$. 
+\end{proof}
+
+\begin{theorem}[Recovery of timepoints of the LSM]
+Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a bezier polynomial of degree $R$ and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
+Let $\hat{X}$ be the ASE of $A$ with embedding vectors $\hat{x}_1, ..., \hat{x}_n \in \mathbb{R}^d$. 
+Let $\hat{t}_1, ..., \hat{t}_n$ be the timepoint estimates outputted by algorithm \ref{alg:bezier-fit}. 
+Then there exists a sequence of Isomap parameters $\lambda_n \to 0$ such that as $n \to \infty$, $|t_i - \hat{t}_i| \stackrel{p}{\to} 0$ for each $i, j \in \{1, ..., n\}$. 
+\end{theorem}
+
+### Recovery of the Underlying Distributions
+
+\begin{algorithm}[H]
+\label{alg:bezier-param-fit}
+\DontPrintSemicolon
+\SetAlgoLined
+\setstretch{1.1}
+\KwData{Adjacency matrix $A$, embedding dimension $d$, probability distribution family $F$, polynomial degree $R$, Isomap neighborhood parameter $\lambda$, stopping criterion $\epsilon$.}
+\KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$.}
+Compute $\hat{X} \in \mathbb{R}^{n \times d}$, the $d$-dimensional ASE of $A$.\;
+Initialize timepoints $\hat{t}_1, ..., \hat{t}_n$, first via a one-dimensional Isomap embedding of $\hat{X}$ using neighborhood parameter $\lambda$, then by normalizing the embedding to $[0, 1]$.\;
+\Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
+  Fit $\hat{b}$ by equation \ref{eq:bezier-ols}.\;
+  \For {$i = 1, ..., n$} {
+    Fit $\hat{t}_i$ by equation by finding the roots of equation \ref{eq:dbezier} and choosing the roots that minimizes equation \ref{eq:bezier-ls}.\;
+  }
+}
+Estimate $\hat{\theta}$ from $\hat{t}_1, ..., \hat{t}_n$ via maximum likelihood estimation.\; 
+\caption{Procedure for estimating the underlying distribution of a Bezier LSM curve from an adjacency matrix.}
+\end{algorithm}
+
+\begin{theorem}[Consistency of algorithm \ref{alg:bezier-param-fit}]
+\end{theorem}
+
+### Classification and Regression
+
+\begin{algorithm}[H]
+\label{alg:mlsm-fit}
+\DontPrintSemicolon
+\SetAlgoLined
+\setstretch{1.1}
+\KwData{Adjacency matrices $A^{(1)}, ..., A^{(L)}$, response variables $y_1, ..., y_L$, embedding dimension $d$, probability distribution family $F$, polynomial degree $R$, Isomap neighborhood parameters $\lambda_1, ..., \lambda_L$, stopping criterion $\epsilon$.}
+\KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$.}
+\For {$\ell = \{1, ..., L\}$} {
+  Compute $\hat{X}^{(\ell)} \in \mathbb{R}^{n_\ell \times d}$, the $d$-dimensional ASE of $A^{(\ell)} \in \mathbb{R}^{n_\ell \times n_\ell}$.\;
+  Initialize timepoints $\hat{t}_1, ..., \hat{t}_{n_\ell}$, first via a one-dimensional Isomap embedding of $\hat{X}^{(\ell)}$ using neighborhood parameter $\lambda_\ell$, then by normalizing the embedding to $[0, 1]$.\;
+  \Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
+    Fit $\hat{b}$ by equation \ref{eq:bezier-ols}.\;
+    \For {$i = 1, ..., {n_\ell}$} {
+      Fit $\hat{t}_i$ by equation by finding the roots of equation \ref{eq:dbezier} and choosing the roots that minimizes equation \ref{eq:bezier-ls}.\;
+    }
+  }
+  Estimate $\hat{\theta}_\ell$ from $\hat{t}_1, ..., \hat{t}_{n_\ell}$ via maximum likelihood estimation.\;
+}
+Fit a model for predicting $y_1, ..., y_L$ using the estimated parameters $\theta_1, ..., \theta_L$ (e.g., via linear regression).\;
+\caption{Procedure for fitting a classification or regression model for an MLSM.}
+\end{algorithm}
+
+\begin{theorem}
+\end{theorem}
+
 # Simulation Study
@@
 1. Draw response variables $y_1, ..., y_n \iid \Multinomial(1/2, 1/2)$.
-2. For each $i = 1, ..., n$, draw $t_i \mid y_i \indep \Betadist(\alpha_{y_i}, \beta_{y_i})$, such that
-    * $\alpha_1 = 1$
-    * $\beta_1 = 2$
-    * $\alpha_2 = 2$
-    * $\beta_2 = 1$
+2. For each $i = 1, ..., n$, draw $t_i \mid y_i \indep \Betadist(\alpha_{y_i}, \beta_{y_i})$, such that $\alpha_1 = 1$, $\beta_1 = 2$, $\alpha_2 = 2$, and $\beta_2 = 1$.
 3. Construct each latent vector as $x_i = g(t_i)$ and compile them in data matrix $X = \bigl[ x_1 \; \cdots \; x_n \bigr]^\top$.
@@
 
-In this dataset, there are $N = `r N`$ graphs (corresponding to individual subjects), each with $n = `r n`$ vertices (corresponding to brain regions). 
+In this dataset, there are $L = `r N`$ graphs (corresponding to individual subjects), each with $n_\ell = n = `r n`$ vertices (corresponding to brain regions). 
 Analyzing these graphs as RDPGs reveals that the DCBM is a good candidate for these data (figure \ref{fig:hcp-ase}).
@@
 
+dcbm.lm <- lm(age ~ B.ll + B.rr + B.lr, data = train.df)
+y.hat.dcbm <- predict(dcbm.lm, newdata = test.df)
+
 rmse.angle <- mean((test.df$age - y.hat.angle) ^ 2) ^ .5
 rmse.deg <- mean((test.df$age - y.hat.deg) ^ 2) ^ .5
+rmse.dcbm <- mean((test.df$age - y.hat.dcbm) ^ 2) ^ .5
 ```
@@
 the average degree within (left) and between (right) hemispheres.', fig.width = 10, fig.height = 3, out.width = '100%'}
-w.plot <- ggplot(ase.stats.df) + 
-  geom_point(aes(x = deg.within, 
-                 y = age)) + 
-  labs(x = 'average degree within hemispheres',
-       y = 'subject age (months)') + 
-  theme_bw() + 
-  stat_smooth(aes(x = deg.within, y = age), method = 'lm')
-b.plot <- ggplot(ase.stats.df) + 
-  geom_point(aes(x = deg.between, 
-                 y = age)) + 
-  labs(x = 'average degree between hemispheres',
-       y = 'subject age (months)') + 
-  theme_bw() + 
-  stat_smooth(aes(x = deg.between, y = age), method = 'lm')
-gridExtra::grid.arrange(w.plot, b.plot, nrow = 1)
+# w.plot <- ggplot(ase.stats.df) + 
+#   geom_point(aes(x = deg.within, 
+#                  y = age)) + 
+#   labs(x = 'average degree within hemispheres',
+#        y = 'subject age (months)') + 
+#   theme_bw() + 
+#   stat_smooth(aes(x = deg.within, y = age), method = 'lm')
+# b.plot <- ggplot(ase.stats.df) + 
+#   geom_point(aes(x = deg.between, 
+#                  y = age)) + 
+#   labs(x = 'average degree between hemispheres',
+#        y = 'subject age (months)') + 
+#   theme_bw() + 
+#   stat_smooth(aes(x = deg.between, y = age), method = 'lm')
+# gridExtra::grid.arrange(w.plot, b.plot, nrow = 1)
 ```
 
-The ASE suggests a latent structure comprised of two line segments, one for each hemisphere of the brain, that meet at the origin. 
-One possible parameter is the angle between the two lines. 
+The ASE suggests a latent structure comprised of two Bezier curves of degree 1 (i.e., lines), one for each hemisphere of the brain, that meet at the origin. 
+One possible parameter when analyzing these data as a MLSM is the angle between the two lines. 
 The estimated angles were observed to correlate with the subject's age, with wider angles corresponding to older subjects (figure \ref{fig:hcp-age-regression}). 
-A linear regression setting aside half of the brain connectivity graphs as test data achieves an RMSE of $`r round(rmse.angle, 2)`$ months.
+A linear regression setting aside half of the brain connectivity graphs as test data achieves an RMSE of $`r round(rmse.angle, 3)`$ months.
 
-Other plausible statistics from the graphs are the average degrees within and between hemispheres. 
-When analyzing these as stochastic block models (SBM) or degree corrected block models (DCBM), these statistics would roughly correspond to the expected edge weights within and between communities. 
-Both of these statistics are correlated with the target value (figure \ref{fig:hcp-age-regression-degree}), but a linear model using them (i.e., $age \sim degree_{within} + degree_{between}$) achieves a higher RMSE ($`r round(rmse.deg, 2)`$ months) while adding another parameter to the model.
-Similarly, the assortativity of each graph with respect to hemisphere results in a higher RMSE of $156.5$ when regressed on age.
+To compare this parameter as a covariate for age against other network statistics, we analyzed these data as a multilayer DCBM, first studied by \citep{agterberg2022joint}, who proposed the degree corrected multiple adjacency spectral embedding (DC-MASE) algorithm.
+Since these graphs come with hemisphere labels, we did not apply DC-MASE for community detection but instead used the estimators for the three edge connectivity parameters, $B_{LL}$, $B_{RR}$, and $B_{LR}$.
+A linear model trained on these parameter estimates achieves a higher RMSE of $`r round(rmse.dcbm, 3)`$ months, despite using three covariates instead of one. 
+In addition, since the angles between the latent structures under the MLSM depend primarily on the shape of the latent structures rather than the exact community memberships, it does not depend on recovery of the original community labels.
+Ultimately, these two methods are estimators for transformations of the same parameters, and while we cannot determine how close these estimates are on the "true" edge connectivity parameters since they are unknown, we observe that the MLSM estimate is a better linear predictor than the rest.
 
-We also note that this structure is consistent with the DCBM. 
-The multilayer DCBM was previously studied by \citet{agterberg2022joint}
+Other graph statistics were extracted for each brain network, and their correlations to age are reported in table \ref{tab:cor-comparison}. 
+Note that aside from the angle between the latent structures under the MLSM model, these statistics assume that it is known which vertex belongs to which hemisphere.
 
@@
 
-```{r cache = TRUE}
+```{r cor-comparison, cache = TRUE}
 assort.df <- foreach(id = ids, .combine = dplyr::bind_rows) %dopar% {
@@
 
-dplyr::tibble(Metric = c('Angle between hemispheres',
+dplyr::tibble(
+  Metric = c('Angle between hemispheres',
+             'Edge connectivity within left hemisphere',
+             'Edge connectivity within right hemisphere',
+             'Edge connectivity between hemispheres',
              'Degree within hemisphere',
@@
              'Assortativity w.r.t. hemisphere',
-                         'Transitivity',
+             # 'Transitivity',
              'Modularity w.r.t. hemisphere'),
   Correlation = c(cor(ase.stats.df$age, ase.stats.df$angle.diff),
+                  cor(ase.stats.df$age, ase.stats.df$B.ll),
+                  cor(ase.stats.df$age, ase.stats.df$B.rr),
+                  cor(ase.stats.df$age, ase.stats.df$B.lr),
                   cor(ase.stats.df$age, ase.stats.df$deg.within),
@@
                   cor(assort.df$age, assort.df$assortativity),
-                              cor(assort.df$age, assort.df$transitivity),
-                              cor(assort.df$age, assort.df$modularity))) %>% 
+                  # cor(assort.df$age, assort.df$transitivity),
+                  cor(assort.df$age, assort.df$modularity)),
+  `95% conf. int.` = c(cor.test(ase.stats.df$age, ase.stats.df$angle.diff)$conf.int %>% 
+                         round(3) %>% 
+                         paste(collapse = ', ') %>% 
+                         paste0('(', ., ')'),
+                       cor.test(ase.stats.df$age, ase.stats.df$B.ll)$conf.int %>% 
+                         round(3) %>% 
+                         paste(collapse = ', ') %>% 
+                         paste0('(', ., ')'),
+                       cor.test(ase.stats.df$age, ase.stats.df$B.rr)$conf.int %>% 
+                         round(3) %>% 
+                         paste(collapse = ', ') %>% 
+                         paste0('(', ., ')'),
+                       cor.test(ase.stats.df$age, ase.stats.df$B.lr)$conf.int %>% 
+                         round(3) %>% 
+                         paste(collapse = ', ') %>% 
+                         paste0('(', ., ')'),
+                       cor.test(ase.stats.df$age, ase.stats.df$deg.within)$conf.int %>% 
+                         round(3) %>% 
+                         paste(collapse = ', ') %>% 
+                         paste0('(', ., ')'),
+                       cor.test(ase.stats.df$age, ase.stats.df$deg.between)$conf.int %>% 
+                         round(3) %>% 
+                         paste(collapse = ', ') %>% 
+                         paste0('(', ., ')'),
+                       cor.test(assort.df$age, assort.df$n_tri1)$conf.int %>% 
+                         round(3) %>% 
+                         paste(collapse = ', ') %>% 
+                         paste0('(', ., ')'),
+                       cor.test(assort.df$age, assort.df$n_tri2)$conf.int %>% 
+                         round(3) %>% 
+                         paste(collapse = ', ') %>% 
+                         paste0('(', ., ')'),
+                       cor.test(assort.df$age, assort.df$assortativity)$conf.int %>% 
+                         round(3) %>% 
+                         paste(collapse = ', ') %>% 
+                         paste0('(', ., ')'),
+                       # cor.test(assort.df$age, assort.df$transitivity)$conf.int %>% 
+                       #   round(3) %>% 
+                       #   paste(collapse = ', ') %>% 
+                       #   paste0('(', ., ')'),
+                       cor.test(assort.df$age, assort.df$modularity)$conf.int %>% 
+                         round(3) %>% 
+                         paste(collapse = ', ') %>% 
+                         paste0('(', ., ')'))) %>% 
   knitr::kable(digits = 3, caption = 'Correlation between age and various graph metrics')

### 2023-11-17 draft.tex
removed large file

@@
 responses, with theorems guaranteeing convergence of the predictions.
-Simulations and applications on functional MRI data verify the
-performance of these methods.
+Simulations, as well as applications on brain connectivity networks,
+verify the performance of our methods.
 \end{abstract}
@@
 
+We propose our methodology for classification and regression in three
+steps:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Curve-fitting algorithm to recover the LSM curves from the ASEs.
+\item
+  Quasi-maximum likelihood estimation to recover the parameters of the
+  underlying distributions.
+\item
+  Classification or regression to determine the relationship between the
+  parameters and the responses.
+\end{enumerate}
+
+\hypertarget{estimation-of-the-lsm-curves}{%
+\subsubsection{Estimation of the LSM
+Curves}\label{estimation-of-the-lsm-curves}}
+
+For curve-fitting, we will restrict the LSM curves to
+non-self-intersecting Bezier polynomials \citep{10.5555/320367}. A
+polynomial curve of degree \(R\) in \(\mathbb{R}^d\) is uniquely defined
+by a Bezier polynomial within the space of Bezier polynomials of degree
+at most \(R\), up to reverse order, although there are infinitely many
+Bezier polynomials of degree greater than \(R\) that can describe the
+curve \citep{SANCHEZREYES2022102118}. We further restrict the curves to
+the case \(p(0) = 0\), i.e., the curve begins at the origin. This is
+consistent with well-known models such as the DCBM and PABM when viewed
+as LSMs or mixtures of LSMs
+\citep{Koo_2022, rubindelanchy2017statistical}, as well as real networks
+that can adequately be described as LSMs \citep{athreya2020estimation}.
+This constraint also removes the reverse order nonidentifiability. Thus,
+if \(R\), the degree of the polynomial, can be determined, and the curve
+begins at the origin, a collection of \(n \geq R\) unique vectors on the
+curve and not at the origin can determine the unique Bezier polynomial
+that describes the curve.
+
+A Bezier polynomial of this form, is defined as \begin{equation}
+\label{eq:bezier}
+p(t; b_1, ..., b_R) = \sum_{r=1}^R \binom{R}{r} b_r (1-t)^r t^{R-r},
+\end{equation} where \(b_r \in \mathbb{R}^d\) are Bezier coefficients
+(control points). In the scenario in which an adjacency matrix \(A\)
+sampled from an LSM with curve \(p\) that can be described by a Bezier
+polynomial, the ASE of \(A\) consists of embedding vectors that lie on
+or near a rotation of the curve. Then if embedding vectors
+\(x_1, ..., x_n \in \mathbb{R}^d\) are such that
+\(x_i = p(t_i) + \epsilon_i\) (for simplicity, redefine \(p\) to be the
+rotated version of the original curve), the squared loss function for
+the curve fit is of the form \begin{equation}
+\label{eq:bezier-loss-full}
+L(t_1, ..., t_n; b_1, ..., b_R) = \sum_{i=1}^n \big( x_i - p(t_i; b_1, ..., b_R) \big)^2.
+\end{equation}
+
+Letting
+\(X = \bigl[ x_1 \; \cdots \; x_n \bigr]^\top \in \mathbb{R}^{n \times p}\),
+\(T \in \mathbb{R}^{n \times R}\) such that
+\(T_{ir} = \binom{R}{r} (1-t_i)^{R-r} t_i^r\),
+\(b = \bigl[b_1 \; \cdots \; b_R\bigr]^\top \in \mathbb{R}^{R \times d}\),
+the loss function can be rewritten as \begin{equation}
+\label{eq:bezier-loss}
+L(T, b) = \|X - T b\|_F^2,
+\end{equation} where \(\|\cdot\|_F\) is the Frobenius norm. Then if
+\(t_1, ..., t_n\) are known, the solution to the least squares best fit
+Bezier coefficients is given by the ordinary least squares estimate,
+\begin{equation}
+\label{eq:bezier-ols}
+\hat{b} = (T^\top T)^{-1} T^\top X.
+\end{equation}
+
+On the other hand, if the coefficients are known but the timepoints
+\(t_1, ..., t_n\) are not, this turns into \(n\) individual minimization
+problems for polynomials of degree \(2 R\). More precisely,
+\(L(t_1, ..., t_n; b_1, ..., b_R) = \sum_{i=1}^n L(t_i; b_1, ..., b_R)\)
+where \begin{equation}
+\label{eq:bezier-ls}
+L(t_i; b_1, ..., b_R) = x_i - \sum_r \binom{R}{r} b_r (1-t_i)^{R-r} t_i^r. 
+\end{equation} To solve, we find the at most \(2 R - 1\) roots of its
+derivative, \begin{equation}
+\label{eq:dbezier}
+\dot{L}(t_i; b_1, ..., b_R) = \bigg(\sum_{r=1}^R \binom{R}{r} (-1)^r c_r t_i^r \bigg) \bigg(\sum_{r=0}^{R-1} \binom{R-1}{r} (-1)^r c_{r+1} t_i^r \bigg) = 0,
+\end{equation} where \(c_r = \sum_{s=1}^r (-1)^{r-s} \binom{r}{s} b_s\).
+Since each \(t_i\) can be solved separately, this method is highly
+parallelizable.
+
+For initialization, the ordering of \(t_1, ..., t_n\) is determined via
+a one-dimensional Isomap embedding \citep{Tenenbaum2000-ff}.
+
+Combining equation \ref{eq:bezier-ols} with the solutions to equation
+\ref{eq:dbezier} for each \(i = 1, ..., n\) provides an alternating
+coordinate descent algorithm, as outlined in algorithm
+\ref{alg:bezier-fit}.
+
+\begin{algorithm}[h]
+\label{alg:bezier-fit}
+\DontPrintSemicolon
+\SetAlgoLined
+\setstretch{1.1}
+\KwData{Adjacency matrix $A$, embedding dimension $d$, polynomial degree $R$, Isomap neighborhood parameter $\lambda$, stopping criterion $\epsilon$.}
+\KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$.}
+Compute $\hat{X} \in \mathbb{R}^{n \times d}$, the ASE of $A$.\;
+Initialize timepoints $\hat{t}_1, ..., \hat{t}_n$, first via a one-dimensional Isomap embedding of $\hat{X}$ using neighborhood parameter $\lambda$, then by normalizing the embedding to $[0, 1]$.\;
+\Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
+  Fit $\hat{b}$ by equation \ref{eq:bezier-ols}.\;
+  \For {$i = 1, ..., n$} {
+    Fit $\hat{t}_i$ by equation by finding the roots of equation \ref{eq:dbezier} and choosing the roots that minimizes equation \ref{eq:bezier-ls}.\;
+  }
+}
+\caption{Procedure for estimating an LSM curve as a Bezier polynomial from an adjacency matrix.}
+\end{algorithm}
+
+In the case of a mixture of LSMs (as in remark \ref{rem:lsm-mixture}),
+algorithm \ref{alg:bezier-fit} is modified by fitting Bezier polynomials
+for each vertex label (if the labels are known) or by estimating the
+labels via an alternating coordinate descent algorithm outlined in
+algorithm \ref{alg:bezier-fit-mixture}.
+
+\begin{algorithm}[H]
+\label{alg:bezier-fit-mixture}
+\DontPrintSemicolon
+\SetAlgoLined
+\setstretch{1.1}
+\KwData{Adjacency matrix $A$, embedding dimension $d$, polynomial degree $R$, number of latent structures $K$, Isomap neighborhood parameter $\lambda$, stopping criterion $\epsilon$.}
+\KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$, curve memberships $\hat{z}_1, ..., \hat{z}_n$.}
+Compute $\hat{X} \in \mathbb{R}^{n \times d}$, the ASE of $A$.\;
+Initialize $\hat{z}_1, ..., \hat{z}_n \in \{1, ..., K\}$, the membership of each vertex to each latent curve.\;
+\For {$k = 1, ..., K$} {
+  Subset the rows of $\hat{X}$ by $z_i = k$ to obtain $\hat{x}_{k_1}, ..., \hat{x}_{k_{n_k}}$, where $n_k$ is the number of vertices with label $k$.\;
+  Initialize timepoints $\hat{t}_{k_1}, ..., \hat{t}_{k_{n_k}}$, first via a one-dimensional Isomap embedding of $\hat{x}_{k_1}, ..., \hat{x}_{k_{n_k}}$ using neighborhood parameter $\lambda$, then by normalizing the embedding to $[0, 1]$.\;
+}
+\Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
+  \For {$k = 1, ..., K$} {
+    Fit $\hat{b}^{(k)}$ by equation \ref{eq:bezier-ols}, using the row subsets of $T$ and $X$ such that $z_i = k$.\;
+    \For {$i = 1, ..., n_k$} {
+      Fit $\hat{t}_{k_i}$ by equation by finding the roots of equation \ref{eq:dbezier} and choosing the roots that minimizes equation \ref{eq:bezier-ls}.\;
+    }
+  }
+  \For {$k = 1, ..., n$} {
+    Reassign $z_i$ based on the latent curve closest to $\hat{x}_i$. 
+  }
+}
+\caption{Procedure for estimating multiple LSM curves as a Bezier polynomial from an adjacency matrix.}
+\end{algorithm}
+
+\begin{lemma}[Recovery of LSM curves in noiseless case]
+\label{lem:lsm-noiseless}
+Suppose $t_1, ..., t_n \iid F_\theta$ such that $F_\theta$ has probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
+Let each $x_i = p(t_i)$ such that $p: [0, 1] \to \mathbb{R}^d$ is a non-self-intersecting Bezier polynomial of degree $R$ that begins at the origin with unknown coefficients $b_1, ..., b_R$. 
+Then there exists a sequence of Isomap neighborhood parameters $\lambda_n \to 0$ such that as $n \to \infty$, algorithm \ref{alg:bezier-fit}, $L(t_1, ..., t_n; b_1, ..., b_R) \stackrel{p}{\to} 0$.
+\end{lemma}
+
+\begin{proof}
+Let $z_1, ..., z_n$ be the one-dimensional Isomap embedding of $x_1, ..., x_n$ using neighborhood parameter $\lambda_n$, and let $l_{ij}$ be the arc length along the curve defined by $p$ between $x_i$ and $x_j$.
+Then by theorem 5 of \citet{acharyya2023}, $|z_i - z_j| - l_{ij} \stackrel{p}{\to} 0$ $\forall i, j \in \{1, ...,n\}$. 
+Let $z_{\min} = \min_i z_i$ and $z_{\max} = \max_i z_i$, and initialize each $\hat{t}_i = (z_i - z_{\min}) / (z_{\max} - z_{\min})$. 
+\end{proof}
+
+\begin{theorem}[Recovery of timepoints of the LSM]
+Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a bezier polynomial of degree $R$ and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
+Let $\hat{X}$ be the ASE of $A$ with embedding vectors $\hat{x}_1, ..., \hat{x}_n \in \mathbb{R}^d$. 
+Let $\hat{t}_1, ..., \hat{t}_n$ be the timepoint estimates outputted by algorithm \ref{alg:bezier-fit}. 
+Then there exists a sequence of Isomap parameters $\lambda_n \to 0$ such that as $n \to \infty$, $|t_i - \hat{t}_i| \stackrel{p}{\to} 0$ for each $i, j \in \{1, ..., n\}$. 
+\end{theorem}
+
+\hypertarget{recovery-of-the-underlying-distributions}{%
+\subsubsection{Recovery of the Underlying
+Distributions}\label{recovery-of-the-underlying-distributions}}
+
+\begin{algorithm}[H]
+\label{alg:bezier-param-fit}
+\DontPrintSemicolon
+\SetAlgoLined
+\setstretch{1.1}
+\KwData{Adjacency matrix $A$, embedding dimension $d$, probability distribution family $F$, polynomial degree $R$, Isomap neighborhood parameter $\lambda$, stopping criterion $\epsilon$.}
+\KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$.}
+Compute $\hat{X} \in \mathbb{R}^{n \times d}$, the $d$-dimensional ASE of $A$.\;
+Initialize timepoints $\hat{t}_1, ..., \hat{t}_n$, first via a one-dimensional Isomap embedding of $\hat{X}$ using neighborhood parameter $\lambda$, then by normalizing the embedding to $[0, 1]$.\;
+\Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
+  Fit $\hat{b}$ by equation \ref{eq:bezier-ols}.\;
+  \For {$i = 1, ..., n$} {
+    Fit $\hat{t}_i$ by equation by finding the roots of equation \ref{eq:dbezier} and choosing the roots that minimizes equation \ref{eq:bezier-ls}.\;
+  }
+}
+Estimate $\hat{\theta}$ from $\hat{t}_1, ..., \hat{t}_n$ via maximum likelihood estimation.\; 
+\caption{Procedure for estimating the underlying distribution of a Bezier LSM curve from an adjacency matrix.}
+\end{algorithm}
+
+\begin{theorem}[Consistency of algorithm \ref{alg:bezier-param-fit}]
+\end{theorem}
+
+\hypertarget{classification-and-regression}{%
+\subsubsection{Classification and
+Regression}\label{classification-and-regression}}
+
+\begin{algorithm}[H]
+\label{alg:mlsm-fit}
+\DontPrintSemicolon
+\SetAlgoLined
+\setstretch{1.1}
+\KwData{Adjacency matrices $A^{(1)}, ..., A^{(L)}$, response variables $y_1, ..., y_L$, embedding dimension $d$, probability distribution family $F$, polynomial degree $R$, Isomap neighborhood parameters $\lambda_1, ..., \lambda_L$, stopping criterion $\epsilon$.}
+\KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$.}
+\For {$\ell = \{1, ..., L\}$} {
+  Compute $\hat{X}^{(\ell)} \in \mathbb{R}^{n_\ell \times d}$, the $d$-dimensional ASE of $A^{(\ell)} \in \mathbb{R}^{n_\ell \times n_\ell}$.\;
+  Initialize timepoints $\hat{t}_1, ..., \hat{t}_{n_\ell}$, first via a one-dimensional Isomap embedding of $\hat{X}^{(\ell)}$ using neighborhood parameter $\lambda_\ell$, then by normalizing the embedding to $[0, 1]$.\;
+  \Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
+    Fit $\hat{b}$ by equation \ref{eq:bezier-ols}.\;
+    \For {$i = 1, ..., {n_\ell}$} {
+      Fit $\hat{t}_i$ by equation by finding the roots of equation \ref{eq:dbezier} and choosing the roots that minimizes equation \ref{eq:bezier-ls}.\;
+    }
+  }
+  Estimate $\hat{\theta}_\ell$ from $\hat{t}_1, ..., \hat{t}_{n_\ell}$ via maximum likelihood estimation.\;
+}
+Fit a model for predicting $y_1, ..., y_L$ using the estimated parameters $\theta_1, ..., \theta_L$ (e.g., via linear regression).\;
+\caption{Procedure for fitting a classification or regression model for an MLSM.}
+\end{algorithm}
+
+\begin{theorem}
+\end{theorem}
+
 \hypertarget{simulation-study}{%
@@
   \(t_i \mid y_i \indep \Betadist(\alpha_{y_i}, \beta_{y_i})\), such
-  that
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \(\alpha_1 = 1\)
-  \item
-    \(\beta_1 = 2\)
-  \item
-    \(\alpha_2 = 2\)
-  \item
-    \(\beta_2 = 1\)
-  \end{itemize}
+  that \(\alpha_1 = 1\), \(\beta_1 = 2\), \(\alpha_2 = 2\), and
+  \(\beta_2 = 1\).
 \item
@@
 
-In this dataset, there are \(N = 516\) graphs (corresponding to
-individual subjects), each with \(n = 84\) vertices (corresponding to
-brain regions). Analyzing these graphs as RDPGs reveals that the DCBM is
-a good candidate for these data (figure \ref{fig:hcp-ase}).
+In this dataset, there are \(L = 516\) graphs (corresponding to
+individual subjects), each with \(n_\ell = n = 84\) vertices
+(corresponding to brain regions). Analyzing these graphs as RDPGs
+reveals that the DCBM is a good candidate for these data (figure
+\ref{fig:hcp-ase}).
 
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\linewidth]{draft_files/figure-latex/hcp-age-regression-degree-1}
-
-}
-
-\caption{Scatterplot between the subject age (in months) vs. the average degree within (left) and between (right) hemispheres.}\label{fig:hcp-age-regression-degree}
-\end{figure}
-
-The ASE suggests a latent structure comprised of two line segments, one
-for each hemisphere of the brain, that meet at the origin. One possible
-parameter is the angle between the two lines. The estimated angles were
+The ASE suggests a latent structure comprised of two Bezier curves of
+degree 1 (i.e., lines), one for each hemisphere of the brain, that meet
+at the origin. One possible parameter when analyzing these data as a
+MLSM is the angle between the two lines. The estimated angles were
 observed to correlate with the subject's age, with wider angles
@@
 linear regression setting aside half of the brain connectivity graphs as
-test data achieves an RMSE of \(11.89\) months.
+test data achieves an RMSE of \(11.889\) months.
 
-Other plausible statistics from the graphs are the average degrees
-within and between hemispheres. When analyzing these as stochastic block
-models (SBM) or degree corrected block models (DCBM), these statistics
-would roughly correspond to the expected edge weights within and between
-communities. Both of these statistics are correlated with the target
-value (figure \ref{fig:hcp-age-regression-degree}), but a linear model
-using them (i.e., \(age \sim degree_{within} + degree_{between}\))
-achieves a higher RMSE (\(12.3\) months) while adding another parameter
-to the model. Similarly, the assortativity of each graph with respect to
-hemisphere results in a higher RMSE of \(156.5\) when regressed on age.
+To compare this parameter as a covariate for age against other network
+statistics, we analyzed these data as a multilayer DCBM, first studied
+by \citep{agterberg2022joint}, who proposed the degree corrected
+multiple adjacency spectral embedding (DC-MASE) algorithm. Since these
+graphs come with hemisphere labels, we did not apply DC-MASE for
+community detection but instead used the estimators for the three edge
+connectivity parameters, \(B_{LL}\), \(B_{RR}\), and \(B_{LR}\). A
+linear model trained on these parameter estimates achieves a higher RMSE
+of \(12.889\) months, despite using three covariates instead of one. In
+addition, since the angles between the latent structures under the MLSM
+depend primarily on the shape of the latent structures rather than the
+exact community memberships, it does not depend on recovery of the
+original community labels. Ultimately, these two methods are estimators
+for transformations of the same parameters, and while we cannot
+determine how close these estimates are on the ``true'' edge
+connectivity parameters since they are unknown, we observe that the MLSM
+estimate is a better linear predictor than the rest.
 
-We also note that this structure is consistent with the DCBM. The
-multilayer DCBM was previously studied by \citet{agterberg2022joint}
+Other graph statistics were extracted for each brain network, and their
+correlations to age are reported in table \ref{tab:cor-comparison}. Note
+that aside from the angle between the latent structures under the MLSM
+model, these statistics assume that it is known which vertex belongs to
+which hemisphere.
 
@@
 
-\begin{longtable}[]{@{}lr@{}}
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5915}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1690}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2394}}@{}}
 \caption{Correlation between age and various graph
@@
 \toprule\noalign{}
-Metric & Correlation \\
+\begin{minipage}[b]{\linewidth}\raggedright
+Metric
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Correlation
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+95\% conf. int.
+\end{minipage} \\
 \midrule\noalign{}
@@
 \endlastfoot
-Angle between hemispheres & 0.558 \\
-Degree within hemisphere & 0.436 \\
-Degree between hemispheres & -0.499 \\
-Number of triangles in left hemisphere & -0.025 \\
-Number of triangles in right hemisphere & 0.002 \\
-Assortativity w.r.t. hemisphere & 0.436 \\
-Transitivity & -0.261 \\
-Modularity w.r.t. hemisphere & 0.434 \\
+Angle between hemispheres & 0.558 & (0.496, 0.615) \\
+Edge connectivity within left hemisphere & 0.241 & (0.158, 0.321) \\
+Edge connectivity within right hemisphere & 0.243 & (0.16, 0.322) \\
+Edge connectivity between hemispheres & -0.436 & (-0.503, -0.363) \\
+Degree within hemisphere & 0.436 & (0.363, 0.503) \\
+Degree between hemispheres & -0.499 & (-0.561, -0.431) \\
+Number of triangles in left hemisphere & -0.025 & (-0.111, 0.062) \\
+Number of triangles in right hemisphere & 0.002 & (-0.084, 0.088) \\
+Assortativity w.r.t. hemisphere & 0.436 & (0.363, 0.503) \\
+Modularity w.r.t. hemisphere & 0.434 & (0.362, 0.502) \\
 \end{longtable}

### 2023-11-17 hcp-a-eda.R

@@ -96,6 +96,13 @@ ase.stats.df <- foreach(id = ids, .combine = dplyr::bind_rows) %dopar% {
   
   A.eigen <- eigen(A, symmetric = TRUE)
   
+  degrees <- colSums(A)
+  n.l <- sum(z == 'L')
+  n.r <- sum(z == 'R')
+  B.ll <- 1 / n.l / n.l * sum(A[z == 'L', z == 'L'])
+  B.rr <- 1 / n.r / n.r * sum(A[z == 'R', z == 'R'])
+  B.lr <- 1 / n.l / n.r * sum(A[z == 'L', z == 'R'])
+  
   individual.stats.df <- dplyr::tibble(
     id = id,
     a.l = params.l[1],
@@ -108,7 +115,10 @@ ase.stats.df <- foreach(id = ids, .combine = dplyr::bind_rows) %dopar% {
     deg.within = deg.within,
     deg.between = deg.between,
     lambda1 = A.eigen$values[1],
-    lambda2 = A.eigen$values[2]
+    lambda2 = A.eigen$values[2],
+    B.ll = B.ll,
+    B.rr = B.rr,
+    B.lr = B.lr
   )
   
   return(individual.stats.df)
@@ -119,7 +129,7 @@ ase.stats.df <- foreach(id = ids, .combine = dplyr::bind_rows) %dopar% {
     by = 'id'
   )
 
-GGally::ggpairs(ase.stats.df, columns = c(2:14), 
+GGally::ggpairs(ase.stats.df, columns = c(5:17), 
                 aes(colour = sex)) + 
   theme_bw()
 

### 2023-11-20 hcp-a-eda.R
syncing stuff

@@ -105,13 +105,13 @@ ase.stats.df <- foreach(id = ids, .combine = dplyr::bind_rows) %dopar% {
   
   individual.stats.df <- dplyr::tibble(
     id = id,
-    a.l = params.l[1],
-    b.l = params.l[2],
-    a.r = params.r[1],
-    b.r = params.r[2],
+    # a.l = params.l[1],
+    # b.l = params.l[2],
+    # a.r = params.r[1],
+    # b.r = params.r[2],
     angle.diff = angle.diff,
-    radius.l = radius.l,
-    radius.r = radius.r,
+    # radius.l = radius.l,
+    # radius.r = radius.r,
     deg.within = deg.within,
     deg.between = deg.between,
     lambda1 = A.eigen$values[1],
@@ -129,7 +129,7 @@ ase.stats.df <- foreach(id = ids, .combine = dplyr::bind_rows) %dopar% {
     by = 'id'
   )
 
-GGally::ggpairs(ase.stats.df, columns = c(5:17), 
+GGally::ggpairs(ase.stats.df, columns = c(2:10), 
                 aes(colour = sex)) + 
   theme_bw()
 

### 2024-01-12 draft.Rmd
minor updates

@@
 
-\begin{lemma}[Recovery of LSM curves in noiseless case]
-\label{lem:lsm-noiseless}
-Suppose $t_1, ..., t_n \iid F_\theta$ such that $F_\theta$ has probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
-Let each $x_i = p(t_i)$ such that $p: [0, 1] \to \mathbb{R}^d$ is a non-self-intersecting Bezier polynomial of degree $R$ that begins at the origin with unknown coefficients $b_1, ..., b_R$. 
-Then there exists a sequence of Isomap neighborhood parameters $\lambda_n \to 0$ such that as $n \to \infty$, algorithm \ref{alg:bezier-fit}, $L(t_1, ..., t_n; b_1, ..., b_R) \stackrel{p}{\to} 0$.
+\begin{lemma}
+\label{lem:bezier-consistency}
+Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R$ with coefficients $b$, and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
+Let $\hat{T}$ and $\hat{b}$ be the minimizers of the loss function in equation \ref{eq:bezier-ls} from $A$, and let $\tilde{X} = \hat{T} \hat{b}$, the estimated vectors along the fitted Bezier curve. 
+Then as $n \to \infty$, $\|X - \tilde{X} W\|_{2,\infty} \stackrel{p}{\to} 0$ for some $W \in \mathbb{O}(d)$.
 \end{lemma}
@@
 \begin{proof}
-Let $z_1, ..., z_n$ be the one-dimensional Isomap embedding of $x_1, ..., x_n$ using neighborhood parameter $\lambda_n$, and let $l_{ij}$ be the arc length along the curve defined by $p$ between $x_i$ and $x_j$.
-Then by theorem 5 of \citet{acharyya2023}, $|z_i - z_j| - l_{ij} \stackrel{p}{\to} 0$ $\forall i, j \in \{1, ...,n\}$. 
-Let $z_{\min} = \min_i z_i$ and $z_{\max} = \max_i z_i$, and initialize each $\hat{t}_i = (z_i - z_{\min}) / (z_{\max} - z_{\min})$. 
+Let $X$ be the true latent positions of the LSM and $\hat{X}$ be the ASE of $A$ drawn from $X$. 
+Then $X = T b$, and $\hat{X} = \hat{T} \hat{b} + \hat{E}$, where $\hat{E}$ is the matrix of distances from the estimated positions along the fitted Bezier curves and the embedding vectors. 
+Then
+$$
+\begin{aligned}
+\|X - \tilde{X} W\| & = \|T b - \hat{T} \hat{b} W\|_{2,\infty} \\
+& = \|T b - (\hat{T} \hat{b} + \hat{E} - \hat{E}) W \|_{2,\infty} \\
+& = \|T b - (\hat{T} \hat{b} + \hat{E}) W + \hat{E} W\|_{2,\infty} \\
+& \leq \|X - \hat{X} W\|_{2,\infty} + \|\hat{E} W\|_{2,\infty}.
+\end{aligned}
+$$
+The first term $\|X - \hat{X} W\|_{2,\infty}$ converges to $0$ in probability \citep{lyzinski2014}. 
+Since the ASE is consistent, the embedding vectors will converge in probability to a Bezier curve, so $\|\hat{E} W\|_{2,\infty} \stackrel{p}{\to} 0$. 
 \end{proof}
 
-\begin{theorem}[Recovery of timepoints of the LSM]
-Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a bezier polynomial of degree $R$ and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
-Let $\hat{X}$ be the ASE of $A$ with embedding vectors $\hat{x}_1, ..., \hat{x}_n \in \mathbb{R}^d$. 
-Let $\hat{t}_1, ..., \hat{t}_n$ be the timepoint estimates outputted by algorithm \ref{alg:bezier-fit}. 
-Then there exists a sequence of Isomap parameters $\lambda_n \to 0$ such that as $n \to \infty$, $|t_i - \hat{t}_i| \stackrel{p}{\to} 0$ for each $i, j \in \{1, ..., n\}$. 
+\begin{theorem}[Consistency of algorithm \ref{alg:bezier-fit}]
+\label{thm:bezier-consistency}
+Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R$ with coefficients $b$, and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
+Let $\hat{T}$ and $\hat{b}$ be the minimizers of the loss function in equation \ref{eq:bezier-ls} from $A$. 
+Then as $n \to \infty$, $\|T - \hat{T}\|_{2, \infty} \stackrel{p}{\to} 0$ where $T$ is defined as in equation \ref{eq:bezier-loss}, and $\|b - \hat{b} W \|_F \stackrel{p}{\to} 0$ for some $W \in \mathbb{O}(d)$.
 \end{theorem}
+
+\begin{proof}
+By lemma \ref{lem:bezier-consistency}, $\|T b - \hat{T} \hat{b} W\|_{2,\infty} \stackrel{p}{\to} 0$ as $n \to \infty$. 
+Then
+
+$$
+\begin{aligned}
+\|T b - \hat{T} \hat{b} W\|_{2,\infty} & = \|T b - \hat{T} \hat{b} W + \hat{T} b - \hat{T} b\|_{2,\infty} \\
+& = \|(T - \hat{T}) b - \hat{T} (\hat{b} W - b)\|_{2,\infty} \\
+& \leq \|(T - \hat{T}) b\|_{2,\infty} + \|\hat{T} (\hat{b} W - b)\|_{2,\infty} \\
+& \leq \|T - \hat{T}\|_{2,\infty} \|b\|_F + \|\hat{T}\|_{2,\infty} \|b - \hat{b} W\|_F \\
+& \stackrel{p}{\to} 0.
+\end{aligned}
+$$
+$\|b\|_F$ and $\|\hat{T}\|_{2,\infty}$ are finite, so both $\|T - \hat{T}\|_{2,\infty}$ and $\|b - \hat{b} W\|_F$ must converge to $0$ in probability. 
+\end{proof}
 
@@
 \begin{theorem}[Consistency of algorithm \ref{alg:bezier-param-fit}]
+Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R$ with coefficients $b$, and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
+Let $\hat{t}_1, ..., \hat{t}_n$ and $\hat{b}$ be the minimizers of the loss function in equation \ref{eq:bezier-ls} from $A$, and let $\tilde{theta}$ be the maximum likelihood estimator for $\theta$ using $\hat{t}_1, ..., \hat{t}_n$.
+Then $\tilde{\theta} \stackrel{p}{\to} \theta$. 
 \end{theorem}

### 2024-01-12 draft.Rmd
minor updates

@@
 & \leq \|(T - \hat{T}) b\|_{2,\infty} + \|\hat{T} (\hat{b} W - b)\|_{2,\infty} \\
-& \leq \|T - \hat{T}\|_{2,\infty} \|b\|_F + \|\hat{T}\|_{2,\infty} \|b - \hat{b} W\|_F \\
-& \stackrel{p}{\to} 0.
+& \leq \|T - \hat{T}\|_{2,\infty} \|b\|_F + \|\hat{T}\|_{2,\infty} \|b - \hat{b} W\|_F.
 \end{aligned}
 $$
-$\|b\|_F$ and $\|\hat{T}\|_{2,\infty}$ are finite, so both $\|T - \hat{T}\|_{2,\infty}$ and $\|b - \hat{b} W\|_F$ must converge to $0$ in probability. 
+Since $\|T b - \hat{T} \hat{b} W\|_{2,\infty} \stackrel{p}{to} 0$, $\|T - \hat{T}\|_{2,\infty} \|b\|_F + \|\hat{T}\|_{2,\infty} \|b - \hat{b} W\|_F \stackrel{p}{\to} 0$. 
+Then since $\|b\|_F$ and $\|\hat{T}\|_{2,\infty}$ are finite, both $\|T - \hat{T}\|_{2,\infty}$ and $\|b - \hat{b} W\|_F$ must converge to $0$ in probability. 
 \end{proof}

### 2024-01-26 draft.Rmd
found issues with proofs--working on fixing them

--- 2024-01-12 draft.Rmd
@@
 
+\begin{lemma}
+Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R$ with coefficients $b$, and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
+Let $\hat{T}$ and $\hat{b}$ be the minimizers of the loss function in equation \ref{eq:bezier-ls} from $A$. 
+Then $\|T - \hat{T}\|_{2, \infty} \stackrel{p}{\to} 0$ as if and only if $\|b - \hat{b} W\|_F \stackrel{p}{\to} 0$. 
+\end{lemma}
+
 \begin{theorem}[Consistency of algorithm \ref{alg:bezier-fit}]
 \label{thm:bezier-consistency}
-Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R$ with coefficients $b$, and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
+Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R$ with finite coefficients $b$, and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
 Let $\hat{T}$ and $\hat{b}$ be the minimizers of the loss function in equation \ref{eq:bezier-ls} from $A$. 
@@
 \begin{proof}
-By lemma \ref{lem:bezier-consistency}, $\|T b - \hat{T} \hat{b} W\|_{2,\infty} \stackrel{p}{\to} 0$ as $n \to \infty$. 
-Then
-
-$$
-\begin{aligned}
-\|T b - \hat{T} \hat{b} W\|_{2,\infty} & = \|T b - \hat{T} \hat{b} W + \hat{T} b - \hat{T} b\|_{2,\infty} \\
-& = \|(T - \hat{T}) b - \hat{T} (\hat{b} W - b)\|_{2,\infty} \\
-& \leq \|(T - \hat{T}) b\|_{2,\infty} + \|\hat{T} (\hat{b} W - b)\|_{2,\infty} \\
-& \leq \|T - \hat{T}\|_{2,\infty} \|b\|_F + \|\hat{T}\|_{2,\infty} \|b - \hat{b} W\|_F \\
-& \stackrel{p}{\to} 0.
-\end{aligned}
-$$
-$\|b\|_F$ and $\|\hat{T}\|_{2,\infty}$ are finite, so both $\|T - \hat{T}\|_{2,\infty}$ and $\|b - \hat{b} W\|_F$ must converge to $0$ in probability. 
+
 \end{proof}
@@
 \end{theorem}
+
+\begin{proof}
+By theorem \ref{thm:bezier-consistency}, 
+\end{proof}
 

### 2024-01-26 draft.tex
found issues with proofs--working on fixing them

--- 2024-01-12 draft.tex
@@
 \begin{proof}
-By lemma \ref{lem:bezier-consistency}, $\|T b - \hat{T} \hat{b} W\|_{2,\infty} \stackrel{p}{\to} 0$ as $n \to \infty$. 
-Then
 
-$$
-\begin{aligned}
-\|T b - \hat{T} \hat{b} W\|_{2,\infty} & = \|T b - \hat{T} \hat{b} W + \hat{T} b - \hat{T} b\|_{2,\infty} \\
-& = \|(T - \hat{T}) b - \hat{T} (\hat{b} W - b)\|_{2,\infty} \\
-& \leq \|(T - \hat{T}) b\|_{2,\infty} + \|\hat{T} (\hat{b} W - b)\|_{2,\infty} \\
-& \leq \|T - \hat{T}\|_{2,\infty} \|b\|_F + \|\hat{T}\|_{2,\infty} \|b - \hat{b} W\|_F \\
-& \stackrel{p}{\to} 0.
-\end{aligned}
-$$
-$\|b\|_F$ and $\|\hat{T}\|_{2,\infty}$ are finite, so both $\|T - \hat{T}\|_{2,\infty}$ and $\|b - \hat{b} W\|_F$ must converge to $0$ in probability. 
 \end{proof}

### 2024-02-22 hcp-a-eda.R
adding a few things overdue

@@ -105,13 +105,13 @@ ase.stats.df <- foreach(id = ids, .combine = dplyr::bind_rows) %dopar% {
   
   individual.stats.df <- dplyr::tibble(
     id = id,
-    # a.l = params.l[1],
-    # b.l = params.l[2],
-    # a.r = params.r[1],
-    # b.r = params.r[2],
+    a.l = params.l[1],
+    b.l = params.l[2],
+    a.r = params.r[1],
+    b.r = params.r[2],
     angle.diff = angle.diff,
-    # radius.l = radius.l,
-    # radius.r = radius.r,
+    radius.l = radius.l,
+    radius.r = radius.r,
     deg.within = deg.within,
     deg.between = deg.between,
     lambda1 = A.eigen$values[1],
@@ -292,3 +292,18 @@ A.list <- lapply(ids, function(id) {
   
   return(A)
 })
+
+ase.stats.long.df <- dplyr::tibble(
+  a = c(ase.stats.df$a.l, ase.stats.df$a.r),
+  b = c(ase.stats.df$b.l, ase.stats.df$b.r),
+  radius = c(ase.stats.df$radius.l, ase.stats.df$radius.r),
+  hemisphere = c(rep('left', nrow(ase.stats.df)), rep('right', nrow(ase.stats.df)))
+)
+ggplot(ase.stats.long.df) + 
+  geom_point(aes(x = a, y = b, colour = hemisphere)) + 
+  coord_fixed()
+ggplot(ase.stats.long.df) + 
+  geom_density(aes(x = radius, group = hemisphere))
+
+zhat <- mclust::Mclust(ase.stats.long.df[c('a', 'b')], G = 2)$classification
+plot(a ~ b, data = ase.stats.long.df, col = zhat)

### 2024-02-22 simulations/beta-regression-sim.R

@@ -6,17 +6,18 @@ library(ggplot2)
 
 doMC::registerDoMC(parallel::detectCores())
 
-nrep <- 32
+nrep <- 50
 
 K <- 2
 beta0 <- 1
 beta1 <- 2
 beta2 <- -1
+sigma_eps <- 1
 
 N <- 2 ^ 6
 n.vec <- 2 ^ c(10, 9, 8, 7, 6)
 
-sim.dir <- '~/dev/multilayer-rdpg/simulations'
+sim.dir <- '~/dev/multilayer-rdpg/simulations/beta-regression'
 
 parallelize.outer <- TRUE
 if (parallelize.outer) {
@@ -37,7 +38,8 @@ out.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
                                     progress = FALSE,
                                     show_col_types = FALSE)
     } else {
-      datamat.df <- foreach(i = seq(N), .combine = dplyr::bind_rows, .errorhandling = 'remove') %DO% {
+      datamat.df <- plyr::ldply(seq(N), function(i) {
+        set.seed(i + (rep - 1) * N)
         print(paste(i, '/', N))
         iter.filename <- paste0('beta-reg-n-', n, '-rep-', rep, '-iter-', i, '.csv')
         if (iter.filename %in% dir(sim.dir)) {
@@ -45,37 +47,76 @@ out.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
                                          progress = FALSE,
                                          show_col_types = FALSE)
         } else {
-        a <- runif(1, 0, 2)
-        b <- runif(1, 0, 2)
-        y <- beta0 + beta1 * a + beta2 * b
-        t. <- rbeta(n, a, b)
-        x1 <- t. ^ 2
-        x2 <- 2 * t. * (1 - t.)
-        X <- cbind(x1, x2)
-        P <- X %*% t(X)
-        A <- draw.graph(P)
-        Xhat <- embedding(A, 2, 0)
-        curve.est <- estimate.bezier.curve.2(Xhat, 
-                                             degree = 2, 
-                                             intercept = FALSE, 
-                                             initialization = 'isomap', 
-                                             # min.t = 0, max.t = 1,
-                                             normalize = TRUE,
-                                             parallel = parallelize.curvefit)
-        plot(Xhat); points(curve.est$X, col = 2)
-        t.hat <- curve.est$t
-        plot(t.hat, t.)
[…]

### 2024-02-22 draft.Rmd
forgot to add a few things

@@
 In particular, it has been shown that the RDPG describes a wide range of popular random graph models with rigid latent structures.
-More recently, joint modeling of mutliple random graphs that share common properties or structures across graphs have been introduced, such as the multilayer RDPG, multiple RPDG, multilayer stochastic block model, etc.
+More recently, joint modeling of mutliple random graphs that share common properties or structures across graphs have been introduced, such as the multilayer RDPG, multiple RPDG, and multilayer stochastic block model.
 In this work, we use these joint random graph models in the context of statistical learning, such as classification and regression, by introducing the multiple latent structure model, in which the graphs share a common latent structure with different parameters that correspond to different response variables.
@@
 library(ggraph)
+library(mclust)
 source('~/dev/pabm-grdpg/functions.R')
@@
 
+Next, we considered each hemisphere separately and fit a Beta distribution to the embedding of each half-network. 
+The scatterplot of the fitted $\alpha$ and $\beta$ parameters suggests that there are two clusters of brain networks (figure \ref{fig:hcp-clustering}). 
+Greater $\alpha$ values correspond to more embedding vectors farther away from the origin, corresponding to more nodes of higher degree, which in turn correspond to more brain regions with a greater number of connections.
+Conversely, greater $\beta$ values correspond to more embedding vectors closer to the origin, which in turn correspond to more brain regions with fewer connections. 
+This may suggest that there are underlying groups, one of which tends to have higher connectivity in the left hemisphere, and the other having higher connectivity in the right hemisphere.
+
+```{r hcp-clustering, cache = TRUE, fig.cap = 'Scatterplot of fitted parameters of the Beta distribution.
+The left image are the parameters fitted on the embedding of the left hemisphere, and the right image are the parameters fitted on the embedding of the right hemisphere.', fig.width = 8, fig.height = 4}
+beta.df <- ase.stats.df[c('a.l', 'a.r', 'b.l', 'b.r')]
+
+zhat.l <- mclust::Mclust(beta.df[, c('a.l', 'b.l')], G = 2)$classification
+plot.l <- ggplot(beta.df) + 
+  geom_point(aes(x = a.l, y = b.l, colour = factor(zhat.l))) + 
+  coord_fixed() + 
+  theme_bw() + 
+  labs(x = expression(hat(alpha)[l]), y = expression(hat(beta)[l]), colour = 'cluster') + 
+  scale_colour_brewer(palette = 'Set1')
+
+zhat.r <- mclust::Mclust(beta.df[, c('a.r', 'b.r')], G = 2)$classification
+plot.r <- ggplot(beta.df) + 
+  geom_point(aes(x = a.r, y = b.r, colour = factor(zhat.r))) + 
+  coord_fixed() +
+  theme_bw() + 
+  labs(x = expression(hat(alpha)[r]), y = expression(hat(beta)[r]), colour = 'cluster') + 
+  scale_colour_brewer(palette = 'Set1')
+
+gridExtra::grid.arrange(plot.l, plot.r, nrow = 1)
+```
+
 # Discussion

### 2024-02-22 draft.tex
forgot to add a few things

@@
 graphs have been introduced, such as the multilayer RDPG, multiple RPDG,
-multilayer stochastic block model, etc. In this work, we use these joint
+and multilayer stochastic block model. In this work, we use these joint
 random graph models in the context of statistical learning, such as
@@
 
+Next, we considered each hemisphere separately and fit a Beta
+distribution to the embedding of each half-network. The scatterplot of
+the fitted \(\alpha\) and \(\beta\) parameters suggests that there are
+two clusters of brain networks (figure \ref{fig:hcp-clustering}).
+Greater \(\alpha\) values correspond to more embedding vectors farther
+away from the origin, corresponding to more nodes of higher degree,
+which in turn correspond to more brain regions with a greater number of
+connections. Conversely, greater \(\beta\) values correspond to more
+embedding vectors closer to the origin, which in turn correspond to more
+brain regions with fewer connections. This may suggest that there are
+underlying groups, one of which tends to have higher connectivity in the
+left hemisphere, and the other having higher connectivity in the right
+hemisphere.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/hcp-clustering-1}
+
+}
+
+\caption{Scatterplot of fitted parameters of the Beta distribution.
+The left image are the parameters fitted on the embedding of the left hemisphere, and the right image are the parameters fitted on the embedding of the right hemisphere.}\label{fig:hcp-clustering}
+\end{figure}
+
 \hypertarget{discussion}{%

### 2024-02-22 simulations/beta-regression-sim.R
fixed error in simulation code

@@ -8,14 +8,13 @@ doMC::registerDoMC(parallel::detectCores())
 
 nrep <- 50
 
-K <- 2
 beta0 <- 1
 beta1 <- 2
 beta2 <- -1
 sigma_eps <- 1
 
 N <- 2 ^ 6
-n.vec <- 2 ^ c(10, 9, 8, 7, 6)
+n.vec <- 2 ^ c(9, 8, 7, 6)
 
 sim.dir <- '~/dev/multilayer-rdpg/simulations/beta-regression'
 
@@ -49,7 +48,7 @@ out.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
         } else {
           a <- runif(1, 1, 2)
           b <- runif(1, 1, 2)
-          y <- beta0 + beta1 * a + beta2 * b + rnorm(1, sigma_eps)
+          y <- beta0 + beta1 * a + beta2 * b + rnorm(1, 0, sigma_eps)
           t. <- rbeta(n, a, b)
           x1 <- t. ^ 2
           x2 <- 2 * t. * (1 - t.)

### 2024-02-23 simulations/angle-regression-sim.R
adding a couple more simulation results

@@ -6,62 +6,86 @@ library(ggplot2)
 
 doMC::registerDoMC(parallel::detectCores())
 
-nrep <- 32
+nrep <- 50
 
 a.vec <- c(1, 1)
 b.vec <- c(1, 1)
 theta.min <- pi / 6
 theta.max <- pi / 3
 beta1 <- 1
-beta0 <- 1
+beta0 <- 2
+sigma_eps <- 1
 
 N <- 2 ^ 6
-n.vec <- 2 ^ c(10, 9, 8, 7)
+n.vec <- 2 ^ c(10, 9, 8, 7, 6)
+
+sim.dir <- '~/dev/multilayer-rdpg/simulations/angle-regression'
 
 out.df <- foreach::foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
   print(paste('n =', n))
   foreach::foreach(rep = seq(nrep), .combine = dplyr::bind_rows) %do% {
     print(paste('rep', rep, 'out of', nrep))
-    datamat.df <- foreach::foreach(i = seq(N), .combine = dplyr::bind_rows, .errorhandling = 'remove') %dopar% {
-      z <- sample(seq(2), n, replace = TRUE)
-      t. <- rbeta(n, a.vec[z], b.vec[z])
-      theta <- runif(1, theta.min, theta.max)
-      y <- beta0 + beta1 * theta
-      x1 <- ifelse(z == 1, t., cos(theta) * t.)
-      x2 <- ifelse(z == 1, 0, sin(theta) * t.)
-      X <- cbind(x1, x2)
-      P <- X %*% t(X)
-      A <- draw.graph(P)
-      Xhat <- embedding(A, 2, 0)
-      clustering <- manifold.clustering(Xhat, 
-                                        degree = 1,
-                                        initialization = z,
-                                        normalize = TRUE,
-                                        intercept = FALSE)
-      angle.1 <- atan2(clustering$p[[1]][2, 2],
-                       clustering$p[[1]][2, 1])
-      angle.2 <- atan2(clustering$p[[2]][2, 2],
-                       clustering$p[[2]][2, 1])
-      theta.hat <- abs(angle.2 - angle.1)
-      if (theta.hat > pi) {
-        theta.hat <- 2 * pi - theta.hat
+    rep.filename <- paste0('angle-reg-n-', n, '-rep-', rep, '.csv')
+    if (rep.filename %in% dir(sim.dir)) {
+      one.rep.df <- readr::read_csv(file.path(sim.dir, rep.filename),
[…]

### 2024-02-24 simulations/beta-regression-sim.R
adding more (probably bad) simulation results

@@ -56,28 +56,29 @@ out.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
           P <- X %*% t(X)
           A <- draw.graph(P)
           Xhat <- embedding(A, 2, 0)
-          curve.est <- estimate.bezier.curve.2(Xhat, 
-                                               degree = 2, 
-                                               intercept = FALSE, 
-                                               initialization = 'isomap', 
+          curve.est <- estimate.bezier.curve.2(Xhat,
+                                               degree = 2,
+                                               intercept = FALSE,
+                                               initialization = 'isomap',
                                                # min.t = 0, max.t = 1,
                                                normalize = TRUE,
                                                parallel = parallelize.curvefit)
           t.hat <- curve.est$t
           param.est <- EnvStats::ebeta(t.hat)
-          one.iter.df <- dplyr::tibble(y = y, 
-                                       a = a, 
+          one.iter.df <- dplyr::tibble(n = n,
+                                       y = y,
+                                       a = a,
                                        b = b,
-                                       a.hat = param.est$parameters[1], 
+                                       a.hat = param.est$parameters[1],
                                        b.hat = param.est$parameters[2])
-          readr::write_csv(one.iter.df,
-                           file.path(sim.dir, iter.filename))
+          # readr::write_csv(one.iter.df,
+          #                  file.path(sim.dir, iter.filename))
         }
         return(one.iter.df)
       }, .parallel = parallelize.outer)
       
       # datamat.df <- foreach(i = seq(N), .combine = dplyr::bind_rows, .errorhandling = 'remove') %DO% {
-      #   set.seed(i)
+      #   set.seed(i * (rep - 1) * N)
       #   print(paste(i, '/', N))
       #   iter.filename <- paste0('beta-reg-n-', n, '-rep-', rep, '-iter-', i, '.csv')
       #   if (iter.filename %in% dir(sim.dir)) {
@@ -87,7 +88,7 @@ out.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
       #   } else {
       #   a <- runif(1, 0, 2)
       #   b <- runif(1, 0, 2)
-      #   y <- beta0 + beta1 * a + beta2 * b + rnorm(1, sigma_eps)
+      #   y <- beta0 + beta1 * a + beta2 * b + rnorm(1, 0, sigma_eps)
       #   t. <- rbeta(n, a, b)
       #   x1 <- t. ^ 2
       #   x2 <- 2 * t. * (1 - t.)
[…]

### 2024-03-22 draft.Rmd
added another data analysis

@@
 
+In the first regression simulation,
+
+In the next regression simulation, we applied
+
 1. Draw angles $\theta_1, ..., \theta_N \iid \Uniform(\pi / 6, \pi / 3)$.
@@
 
+## Drosophila Connectome
+
+In the second example, we analyzed the larval *Drosophila* mushroom body connectome \citep{Eichler141762}, which has been studied as a GRDPG by \citet{athreya2020estimation}. 
+This dataset consists of two graphs representing two networks of neurons, one for each hemisphere of the *Drosophila* brain. 
+In these graphs, each vertex is a neuron, and the labels correspond to one of four neuron types (Kenyon Cells, Input Neurons, Output Neurons, and Projection Neurons). 
+The resulting graphs are illustrated in figure \ref{fig:mbconnectome-graph}.
+
+```{r mbconnectome-graph, fig.cap = 'Graphs of the Drosophila connectomes.
+The left and right are of the left and right hemispheres, respectively.
+Each vertex represents a neuron, which are labeled by neuron type.', fig.height = 2, figh.width = 12, cache = TRUE}
+library(mbstructure)
+data(MBconnectome)
+
+graph.r <- generate.graph(newrdat, vdf.right)
+A.r <- as.matrix(igraph::as_adjacency_matrix(graph.r$g, type = 'both'))
+z.r <- as.numeric(graph.r$vdf$type)
+n.r <- length(z.r)
+A.r <- sign(A.r + t(A.r))
+diag(A.r) <- 0
+graph.r <- ggraph(A.r, layout = 'stress') + 
+  geom_edge_link(alpha = 1 / sqrt(n.r) / 2) + 
+  geom_node_point(aes(colour = factor(z.r)), 
+                  show.legend = FALSE) + 
+  labs(colour = NULL) + 
+  theme_void() + 
+  scale_colour_brewer(palette = 'Set1')
+
+graph.l <- generate.graph(newldat, vdf.left)
+A.l <- as.matrix(igraph::as_adjacency_matrix(graph.l$g, type = 'both'))
+z.l <- as.numeric(graph.l$vdf$type)
+n.l <- length(z.l)
+A.l <- sign(A.l + t(A.l))
+diag(A.l) <- 0
+graph.l <- ggraph(A.l, layout = 'stress') + 
+  geom_edge_link(alpha = 1 / sqrt(n.l) / 2) + 
+  geom_node_point(aes(colour = factor(z.l)), 
+                  show.legend = FALSE) + 
+  labs(colour = NULL) + 
+  theme_void() + 
+  scale_colour_brewer(palette = 'Set1')
+
+gridExtra::grid.arrange(graph.l, graph.r, nrow = 1)
+```
+
+```{r mbconnectome-ase, fig.cap = 'ASEs of the Drosophila connectome graphs. The embedding vectors are labeled by neuron type. ', fig.height = 2, fig.width = 8}
+Xhat.l <- embedding(A.l, 2, 1)
+ase.l <- ggplot() + 
+  geom_point(aes(x = Xhat.l[, 1], y = Xhat.l[, 2],
+                 colour = factor(z.l))) + 
+  coord_fixed(.5) +
+  theme_bw() + 
+  theme(legend.position = 'none') +
+  scale_colour_brewer(palette = 'Set1') + 
+  labs(x = NULL, y = NULL, colour = NULL)
+
+Xhat.r <- embedding(A.r, 2, 1)
+ase.r <- ggplot() + 
+  geom_point(aes(x = Xhat.r[, 1], y = Xhat.r[, 2],
+                 colour = factor(z.r))) + 
+  coord_fixed(.5) +
+  theme_bw() + 
+  theme(legend.position = 'none') +
+  scale_colour_brewer(palette = 'Set1') + 
+  labs(x = NULL, y = NULL, colour = NULL)
+
+gridExtra::grid.arrange(ase.l, ase.r, nrow = 1)
+```
+
+When analyzed as a GRDPG, the ASE of each hemisphere suggests that nodes of each type falls along a curve in the latent space (figure \ref{fig:mbconnectome-ase}). 
+This matches observations by \citet{athreya2020estimation}. 
+In our analysis, we set the embedding dimension to $d = 4$, with 3 assortative dimensions and 1 disassortative dimension. 
+We fit three latent structure Bezier curves, one for Kenyon Cells, one for Input and Output Neurons, and one for Projection Neurons, to the embedding for each hemisphere.
+Then we fit a Beta distribution to to the timepoints along each curve and extracted the two Beta parameter estimates (via likelihood maximization) for each curve. 
+If these Beta parameters are informative, we would expect the parameters for each hemisphere to match by neuron type, which is what we observe in these data (fig \ref{fig:mbconnectome-beta}).
+
+```{r mbconnectome-beta, fig.cap = 'Beta parameter estimates for each curve and hemisphere.', fig.height = 3, fig.width = 6}
+param.est.df <- readr::read_csv('~/dev/multilayer-rdpg/drosophila.csv')
+
+param.est.df %>% 
+ggplot() + 
+  geom_point(aes(x = shape1, y = shape2, colour = region, shape = hemisphere)) +
+  labs(x = expression(hat(alpha)),
+       y = expression(hat(beta)),
+       colour = 'Neuron type',
+       shape = 'Hemisphere') +
+  coord_fixed() + 
+  theme_bw()
+```
+
 ## Human Connectome Project Aging Study
@@
 
-In the first example, we analyzed fiber count data between brain regions from the Human Connectome Project (HCP). 
+In the second example, we analyzed fiber count data between brain regions from the Human Connectome Project (HCP). 
 When analyzing these data as graphs, we denote the regions as vertices and the fiber counts between pairs of regions as weighted edges. 
@@
 
+# https://github.com/jesusdaniel/JEG
+source('~/dev/JEG/Code/joint_embedding.R')
+A.list <- lapply(ids, function(id) {
+  A <- dat.NumFibers[[id]] %>% 
+    as.matrix()
+  A <- (A + t(A)) / 2
+  diag(A) <- 0
+  return(A)
+})
+jeg.out <- multidembed(A.list, 2)
+jeg.lm <- lm(ase.stats.df$age ~ jeg.out$lambda[, 1] + jeg.out$lambda[, 2])
+jeg.yhat <- jeg.lm$fitted.values
+
 dplyr::tibble(
@@
              # 'Transitivity',
-             'Modularity w.r.t. hemisphere'),
+             'Modularity w.r.t. hemisphere',
+             'Joint Embedding'),
   Correlation = c(cor(ase.stats.df$age, ase.stats.df$angle.diff),
@@
                   # cor(assort.df$age, assort.df$transitivity),
-                  cor(assort.df$age, assort.df$modularity)),
+                  cor(assort.df$age, assort.df$modularity),
+                  cor(assort.df$age, jeg.yhat)),
   `95% conf. int.` = c(cor.test(ase.stats.df$age, ase.stats.df$angle.diff)$conf.int %>% 
@@
                          paste(collapse = ', ') %>% 
+                         paste0('(', ., ')'),
+                       cor.test(assort.df$age, jeg.yhat)$conf.int %>% 
+                         round(3) %>% 
+                         paste(collapse = ', ') %>% 
                          paste0('(', ., ')'))) %>% 

### 2024-03-22 draft.tex
added another data analysis

@@
 
+In the first regression simulation,
+
+In the next regression simulation, we applied
+
 \begin{enumerate}
@@
 
+\hypertarget{drosophila-connectome}{%
+\subsection{Drosophila Connectome}\label{drosophila-connectome}}
+
+In the second example, we analyzed the larval \emph{Drosophila} mushroom
+body connectome \citep{Eichler141762}, which has been studied as a GRDPG
+by \citet{athreya2020estimation}. This dataset consists of two graphs
+representing two networks of neurons, one for each hemisphere of the
+\emph{Drosophila} brain. In these graphs, each vertex is a neuron, and
+the labels correspond to one of four neuron types (Kenyon Cells, Input
+Neurons, Output Neurons, and Projection Neurons). The resulting graphs
+are illustrated in figure \ref{fig:mbconnectome-graph}.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/mbconnectome-graph-1}
+
+}
+
+\caption{Graphs of the Drosophila connectomes.
+The left and right are of the left and right hemispheres, respectively.
+Each vertex represents a neuron, which are labeled by neuron type.}\label{fig:mbconnectome-graph}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/mbconnectome-ase-1}
+
+}
+
+\caption{ASEs of the Drosophila connectome graphs. The embedding vectors are labeled by neuron type. }\label{fig:mbconnectome-ase}
+\end{figure}
+
+When analyzed as a GRDPG, the ASE of each hemisphere suggests that nodes
+of each type falls along a curve in the latent space (figure
+\ref{fig:mbconnectome-ase}). This matches observations by
+\citet{athreya2020estimation}. In our analysis, we set the embedding
+dimension to \(d = 4\), with 3 assortative dimensions and 1
+disassortative dimension. We fit three latent structure Bezier curves,
+one for Kenyon Cells, one for Input and Output Neurons, and one for
+Projection Neurons, to the embedding for each hemisphere. Then we fit a
+Beta distribution to to the timepoints along each curve and extracted
+the two Beta parameter estimates (via likelihood maximization) for each
+curve. If these Beta parameters are informative, we would expect the
+parameters for each hemisphere to match by neuron type, which is what we
+observe in these data (fig \ref{fig:mbconnectome-beta}).
+
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/mbconnectome-beta-1}
+
+}
+
+\caption{Beta parameter estimates for each curve and hemisphere.}\label{fig:mbconnectome-beta}
+\end{figure}
+
 \hypertarget{human-connectome-project-aging-study}{%
@@
 
-In the first example, we analyzed fiber count data between brain regions
-from the Human Connectome Project (HCP). When analyzing these data as
-graphs, we denote the regions as vertices and the fiber counts between
-pairs of regions as weighted edges. A plausible statistical model for
-these data is to assume that the edge weights between pairs of vertices
-is Poisson distributed, i.e., the adjacency matrix is sampled as
-\(A_{ij} \indep \Poisson(\Theta_{ij})\), where
+In the second example, we analyzed fiber count data between brain
+regions from the Human Connectome Project (HCP). When analyzing these
+data as graphs, we denote the regions as vertices and the fiber counts
+between pairs of regions as weighted edges. A plausible statistical
+model for these data is to assume that the edge weights between pairs of
+vertices is Poisson distributed, i.e., the adjacency matrix is sampled
+as \(A_{ij} \indep \Poisson(\Theta_{ij})\), where
 \(\Theta \in \mathbb{R}_+^{n \times n}\) is a symmetric matrix of
@@
 Modularity w.r.t. hemisphere & 0.434 & (0.362, 0.502) \\
+Joint Embedding & 0.556 & (0.493, 0.613) \\
 \end{longtable}

### 2024-03-22 draft.Rmd
minor edits

@@
 In these graphs, each vertex is a neuron, and the labels correspond to one of four neuron types (Kenyon Cells, Input Neurons, Output Neurons, and Projection Neurons). 
+The number of neurons in each hemisphere is not equal (209 in the left hemisphere and 213 in the right hemisphere).
 The resulting graphs are illustrated in figure \ref{fig:mbconnectome-graph}.
@@
 When analyzed as a GRDPG, the ASE of each hemisphere suggests that nodes of each type falls along a curve in the latent space (figure \ref{fig:mbconnectome-ase}). 
+In our analysis, we set the embedding dimension to $d = 4$, with 3 assortative dimensions and 1 disassortative dimension, and the figure only shows the first two assortative dimensions. 
 This matches observations by \citet{athreya2020estimation}. 
-In our analysis, we set the embedding dimension to $d = 4$, with 3 assortative dimensions and 1 disassortative dimension. 
 We fit three latent structure Bezier curves, one for Kenyon Cells, one for Input and Output Neurons, and one for Projection Neurons, to the embedding for each hemisphere.

### 2024-03-22 draft.tex
minor edits

@@
 the labels correspond to one of four neuron types (Kenyon Cells, Input
-Neurons, Output Neurons, and Projection Neurons). The resulting graphs
-are illustrated in figure \ref{fig:mbconnectome-graph}.
+Neurons, Output Neurons, and Projection Neurons). The number of neurons
+in each hemisphere is not equal (209 in the left hemisphere and 213 in
+the right hemisphere). The resulting graphs are illustrated in figure
+\ref{fig:mbconnectome-graph}.
 
@@
 of each type falls along a curve in the latent space (figure
-\ref{fig:mbconnectome-ase}). This matches observations by
-\citet{athreya2020estimation}. In our analysis, we set the embedding
+\ref{fig:mbconnectome-ase}). In our analysis, we set the embedding
 dimension to \(d = 4\), with 3 assortative dimensions and 1
-disassortative dimension. We fit three latent structure Bezier curves,
-one for Kenyon Cells, one for Input and Output Neurons, and one for
-Projection Neurons, to the embedding for each hemisphere. Then we fit a
-Beta distribution to to the timepoints along each curve and extracted
-the two Beta parameter estimates (via likelihood maximization) for each
-curve. If these Beta parameters are informative, we would expect the
-parameters for each hemisphere to match by neuron type, which is what we
-observe in these data (fig \ref{fig:mbconnectome-beta}).
+disassortative dimension, and the figure only shows the first two
+assortative dimensions. This matches observations by
+\citet{athreya2020estimation}. We fit three latent structure Bezier
+curves, one for Kenyon Cells, one for Input and Output Neurons, and one
+for Projection Neurons, to the embedding for each hemisphere. Then we
+fit a Beta distribution to to the timepoints along each curve and
+extracted the two Beta parameter estimates (via likelihood maximization)
+for each curve. If these Beta parameters are informative, we would
+expect the parameters for each hemisphere to match by neuron type, which
+is what we observe in these data (fig \ref{fig:mbconnectome-beta}).
 

### 2024-04-19 draft.Rmd
still typing up paper

@@
 
-If we let the parameter matrix $P$ be unconstrained, the inference problem is overparameterized. 
+If the parameter matrix $P$ is unconstrained, the inference problem is overparameterized. 
 On the other hand, the classical $\ER$ model is often too restrictive to describe real, observed networks. 
@@
 
-It has been shown \citep{rubindelanchy2017statistical, Koo_2022} that the RDPG (as well as the generalized random dot product graph \citep{rubindelanchy2017statistical}) can describe a wide range of popular random graph models, such as the $\ER$ model, the stochastic block model (SBM) \citep{doi:10.1080/0022250X.1971.9989788}, degree corrected block model (DCBM) \citep{Karrer_2011}, and popularity adjusted block model (PABM) \citep{307cbeb9b1be48299388437423d94bf1}, and describing these models as RDPGs (or its generalized version) is useful for parameter estimation.
-The RDPG captures a wide range of phenomena that can be described as networks, and in fact any network that can be thought of as being sampled from a parameter matrix $P$ can be described as a (generalized) RDPG, especially if $P$ is low rank.
-The RPDG representation of these networks
+It has been shown \citep{Koo_2022, rubindelanchy2017statistical} that the RDPG (as well as the generalized random dot product graph, or GRDPG, \citep{rubindelanchy2017statistical}) can describe a wide range of popular random graph models, such as the $\ER$ model, the stochastic block model (SBM) \citep{doi:10.1080/0022250X.1971.9989788}, degree corrected block model (DCBM) \citep{Karrer_2011}, and popularity adjusted block model (PABM) \citep{307cbeb9b1be48299388437423d94bf1}.
+In these examples, the (G)RPDG representation of these networks allows for inference and analysis in more familiar Euclidean space rather than on graphs, and the geometry of the configuration of latent vectors correspond to the parameters of these graph models.
+More specifically, when these network models are described as (G)RDPGs, each latent vector lies on a subspace. 
+\citet{athreya2020estimation} and \citet{Passino_2020} extended this to the latent structure model (LSM) and latent structure block model (LSBM) to include latent configurations that consist of nonlinear manifolds.
+
+In this paper, we apply these highly structured GRDPG models to supervised learning problems in which there are multiple graphs, $G_1, ..., G_L$, each with a response $y^{(1)}, ..., y^{(L)}$. 
+Each graph, represented by adjacency matrix $A^{(l)}$, is presumed to be drawn from a GRDPG with parameter set $\theta_l$, and in turn, the response $y_l$ is drawn from a univariate distribution with parameter $\theta_l$.
+Since the latent vectors of each $A^{(l)}$ are presumed to be distributed on a structured subset of the latent space, e.g., a manifold, we call this the multiple latent structure model (MLSM). 
+Much previous work has been done on multilayer networks, such as the multilayer degree corrected block model of \citet{agterberg2022joint}, multiple random eigen graphs of \citet{8889404}, and the multiple common subspace independent edge graphs of \citet{arroyo2020inference}.
+In each of these models, the main inference problem is to identify parameters that are shared across multiple graphs. 
+Our contribution is in applying similar methods to identify parameters specific to each graph, which are in turn used as features for supervised learning.
+
+The following sections are organized as follows: Section 2 introduces the structured and parameterized
 
 # Methods
+
+## Notation and Scope
+
+Let $G_1, ..., G_L$ be a collection of undirected graphs where each $G_l = (V_l, E_l)$. 
+In this setup, each graph can have its own vertex and edge sets. 
+Each graph is represented by adjacency matrices $A^{(1)}, ..., A^{(L)}$, where each $A^{(l)} \in \mathbb{R}^{n_l \times n_l}$ is a symmetric and hollow matrix such that $A^{(l)}_{ij}$ represents the edge weight between vertices $i$ and $j$, and $n_l$ represents the size of the vertex set of the $l$^th^ graph.
+Let $P^{(1)}, ..., P^{(L)}$ be the corresponding collection of edge parameter matrices such that each $A^{(l)}_{ij} \indep H(P^{(l)}_{ij})$ for each $1 \leq i < j \leq n_l$ and $H$ is a specified probability distribution such that $E[A^{(l)}_{ij}] = P^{(l)}_{ij}$.
+For ease of notation, we use $A^{(l)} \sim H(P^{(l)})$ for an adjacency matrix $A^{(l)}$ drawn from edge parameter matrix $P^{(l)}$ in this way.
+Denote $X^{(1)}, ..., X^{(L)}$ as the sets of latent vectors corresponding to each graph under the GRDPG framework.
+Each $X^{(l)} \in \mathbb{R}^{n_l \times d}$, i.e., the graphs share the same latent dimension, and $x_i^{(l)} \in \mathbb{R}^d$ is the latent vector corresponding to the $i$^th^ vertex in the $l$^th^ graph.
 
@@
 Let $F_\theta$ be a distribution with support $\mathcal{X}$ and parameters $\theta$, and sample $x_1, ..., x_n \iid F_\theta$. 
-A graph $G$ with adjacency matrix $A$ is a random dot product graph with latent vectors $X = \bigl[x_1 \; \cdots \; x_n\bigr]^\top$ drawn from distribution $F_\theta$ if $A \sim X X^\top$.
+A graph $G$ with adjacency matrix $A$ is a random dot product graph with latent vectors $X = \bigl[x_1 \; \cdots \; x_n\bigr]^\top$ drawn from distribution $F_\theta$ if $A \sim H(X X^\top)$.
 
@@
 
+\begin{remark}
+\label{remark:grdpg}
+The RDPG requires that $P = X X^\top$ be a positive semidefinite matrix. 
+To include matrices that are not positive semidefinite, the generalized random dot product graph (GRDPG) was introduced by \citet{rubindelanchy2017statistical}, in which $P = X I_{p,q} X^\top$. 
+Here, $I_{p,q}$ is a diagonal matrix of $p$ $1$'s followed by $q$ $-1$'s. 
+While the focus of this work is on the RDPG, the theoretical results also apply to the GRDPG. 
+\end{remark}
+
 \begin{definition}[Latent structure model (LSM) \citep{athreya2020estimation}]
@@
 Let $\mathcal{C} \subset \mathcal{X} \subset \mathbb{R}^d$ be a smooth, nonintersecting one-dimensional manifold on the domain of a RDPG as defined in definition \ref{def:rdpg}, parameterized by function $p(t) : [0, 1] \to \mathcal{C}$.
-Then if $t_1, ..., t_n \iid F_\theta$ for some distribution $F_\theta$ with support $[0, 1]$ and parameter $\theta$, each $x_i = p(t_i)$, and $ A \sim X X^\top$ for $X = \bigl[x_1 \; \cdots \; x_n\bigr]^\top$, $A$ is the adjacency matrix of a latent structure model on curve $\mathcal{C}$ with parameterization $p$ and underlying distribution $F_\theta$.
+Then if $t_1, ..., t_n \iid F_\theta$ for some distribution $F_\theta$ with support $[0, 1]$ and parameter $\theta$, each $x_i = p(t_i)$, and $ A \sim H(X X^\top)$ for $X = \bigl[x_1 \; \cdots \; x_n\bigr]^\top$, $A$ is the adjacency matrix of a latent structure model on curve $\mathcal{C}$ with parameterization $p$ and underlying distribution $F_\theta$.
 
@@
 If these Beta parameters are informative, we would expect the parameters for each hemisphere to match by neuron type, which is what we observe in these data (fig \ref{fig:mbconnectome-beta}). 
+Using the three subgraphs of the left hemisphere as a training set to classify which subgraph of the right hemisphere belongs to which neuron type using the extracted Beta distribution parameter estimates as predictors results in 100\% accuracy.
 
@@
 
-```{r cosine-heatmap, cache = TRUE, fig.width = 20, fig.height = 20, out.width = '100%'}
+```{r cosine-heatmap, cache = TRUE, fig.width = 20, fig.height = 20, out.width = '100%', eval=FALSE}
 # i <- ids[1]
@@
 
-```{r cosine-heatmap-2, cache = TRUE, fig.width = 10, fig.height = 10, out.width = '100%'}
+```{r cosine-heatmap-2, cache = TRUE, fig.width = 10, fig.height = 10, out.width = '100%', eval=FALSE}
 angles.plot.1 <- pairwise.angles.df %>%
@@
   Metric = c('Angle between hemispheres',
-             'Edge connectivity within left hemisphere',
-             'Edge connectivity within right hemisphere',
-             'Edge connectivity between hemispheres',
+             # 'Edge connectivity within left hemisphere',
+             # 'Edge connectivity within right hemisphere',
+             # 'Edge connectivity between hemispheres',
              'Degree within hemisphere',
              'Degree between hemispheres',
-             'Number of triangles in left hemisphere',
-             'Number of triangles in right hemisphere',
-             'Assortativity w.r.t. hemisphere',
+             # 'Number of triangles in left hemisphere',
+             # 'Number of triangles in right hemisphere',
+             # Assortativity w.r.t. hemisphere',
              # 'Transitivity',
@@
   Correlation = c(cor(ase.stats.df$age, ase.stats.df$angle.diff),
-                  cor(ase.stats.df$age, ase.stats.df$B.ll),
-                  cor(ase.stats.df$age, ase.stats.df$B.rr),
-                  cor(ase.stats.df$age, ase.stats.df$B.lr),
+                  # cor(ase.stats.df$age, ase.stats.df$B.ll),
+                  # cor(ase.stats.df$age, ase.stats.df$B.rr),
+                  # cor(ase.stats.df$age, ase.stats.df$B.lr),
                   cor(ase.stats.df$age, ase.stats.df$deg.within),
                   cor(ase.stats.df$age, ase.stats.df$deg.between),
-                  cor(assort.df$age, assort.df$n_tri1),
-                  cor(assort.df$age, assort.df$n_tri2),
-                  cor(assort.df$age, assort.df$assortativity),
+                  # cor(assort.df$age, assort.df$n_tri1),
+                  # cor(assort.df$age, assort.df$n_tri2),
+                  # cor(assort.df$age, assort.df$assortativity),
                   # cor(assort.df$age, assort.df$transitivity),
@@
                          paste0('(', ., ')'),
-                       cor.test(ase.stats.df$age, ase.stats.df$B.ll)$conf.int %>% 
-                         round(3) %>% 
-                         paste(collapse = ', ') %>% 
-                         paste0('(', ., ')'),
-                       cor.test(ase.stats.df$age, ase.stats.df$B.rr)$conf.int %>% 
-                         round(3) %>% 
-                         paste(collapse = ', ') %>% 
-                         paste0('(', ., ')'),
-                       cor.test(ase.stats.df$age, ase.stats.df$B.lr)$conf.int %>% 
-                         round(3) %>% 
-                         paste(collapse = ', ') %>% 
-                         paste0('(', ., ')'),
+                       # cor.test(ase.stats.df$age, ase.stats.df$B.ll)$conf.int %>% 
+                       #   round(3) %>% 
+                       #   paste(collapse = ', ') %>% 
+                       #   paste0('(', ., ')'),
+                       # cor.test(ase.stats.df$age, ase.stats.df$B.rr)$conf.int %>% 
+                       #   round(3) %>% 
+                       #   paste(collapse = ', ') %>% 
+                       #   paste0('(', ., ')'),
+                       # cor.test(ase.stats.df$age, ase.stats.df$B.lr)$conf.int %>% 
+                       #   round(3) %>% 
+                       #   paste(collapse = ', ') %>% 
+                       #   paste0('(', ., ')'),
                        cor.test(ase.stats.df$age, ase.stats.df$deg.within)$conf.int %>% 
@@
                          paste0('(', ., ')'),
-                       cor.test(assort.df$age, assort.df$n_tri1)$conf.int %>% 
-                         round(3) %>% 
-                         paste(collapse = ', ') %>% 
-                         paste0('(', ., ')'),
-                       cor.test(assort.df$age, assort.df$n_tri2)$conf.int %>% 
-                         round(3) %>% 
-                         paste(collapse = ', ') %>% 
-                         paste0('(', ., ')'),
-                       cor.test(assort.df$age, assort.df$assortativity)$conf.int %>% 
-                         round(3) %>% 
-                         paste(collapse = ', ') %>% 
-                         paste0('(', ., ')'),
+                       # cor.test(assort.df$age, assort.df$n_tri1)$conf.int %>% 
+                       #   round(3) %>% 
+                       #   paste(collapse = ', ') %>% 
+                       #   paste0('(', ., ')'),
+                       # cor.test(assort.df$age, assort.df$n_tri2)$conf.int %>% 
+                       #   round(3) %>% 
+                       #   paste(collapse = ', ') %>% 
+                       #   paste0('(', ., ')'),
+                       # cor.test(assort.df$age, assort.df$assortativity)$conf.int %>% 
+                       #   round(3) %>% 
+                       #   paste(collapse = ', ') %>% 
+                       #   paste0('(', ., ')'),
                        # cor.test(assort.df$age, assort.df$transitivity)$conf.int %>% 
@@
                          paste0('(', ., ')'))) %>% 
-  knitr::kable(digits = 3, caption = 'Correlation between age and various graph metrics')
+  knitr::kable(digits = 3, caption = 'Correlation between age and various graph metrics.\\label{tab:cor-comparison}')
 ```
@@
 
+\section*{Appendix A: Proofs of Theorems}
+
 \newpage

### 2024-04-19 draft.tex
still typing up paper

@@
 
-If we let the parameter matrix \(P\) be unconstrained, the inference
-problem is overparameterized. On the other hand, the classical \(\ER\)
-model is often too restrictive to describe real, observed networks. Much
-work has been done to develop models that are constrained enough for
-robust statistical inference while being generalizable to describe a
-wide range of networks. One such family of random graph models is the
-random dot product graph (RDPG), first proposed by
+If the parameter matrix \(P\) is unconstrained, the inference problem is
+overparameterized. On the other hand, the classical \(\ER\) model is
+often too restrictive to describe real, observed networks. Much work has
+been done to develop models that are constrained enough for robust
+statistical inference while being generalizable to describe a wide range
+of networks. One such family of random graph models is the random dot
+product graph (RDPG), first proposed by
 \citet{10.1007/978-3-540-77004-6_11}, which is a type of latent space
@@
 
-It has been shown \citep{rubindelanchy2017statistical, Koo_2022} that
-the RDPG (as well as the generalized random dot product graph
+It has been shown \citep{Koo_2022, rubindelanchy2017statistical} that
+the RDPG (as well as the generalized random dot product graph, or GRDPG,
 \citep{rubindelanchy2017statistical}) can describe a wide range of
@@
 corrected block model (DCBM) \citep{Karrer_2011}, and popularity
-adjusted block model (PABM) \citep{307cbeb9b1be48299388437423d94bf1},
-and describing these models as RDPGs (or its generalized version) is
-useful for parameter estimation. The RDPG captures a wide range of
-phenomena that can be described as networks, and in fact any network
-that can be thought of as being sampled from a parameter matrix \(P\)
-can be described as a (generalized) RDPG, especially if \(P\) is low
-rank. The RPDG representation of these networks
+adjusted block model (PABM) \citep{307cbeb9b1be48299388437423d94bf1}. In
+these examples, the (G)RPDG representation of these networks allows for
+inference and analysis in more familiar Euclidean space rather than on
+graphs, and the geometry of the configuration of latent vectors
+correspond to the parameters of these graph models. More specifically,
+when these network models are described as (G)RDPGs, each latent vector
+lies on a subspace. \citet{athreya2020estimation} and
+\citet{Passino_2020} extended this to the latent structure model (LSM)
+and latent structure block model (LSBM) to include latent configurations
+that consist of nonlinear manifolds.
+
+In this paper, we apply these highly structured GRDPG models to
+supervised learning problems in which there are multiple graphs,
+\(G_1, ..., G_L\), each with a response \(y^{(1)}, ..., y^{(L)}\). Each
+graph, represented by adjacency matrix \(A^{(l)}\), is presumed to be
+drawn from a GRDPG with parameter set \(\theta_l\), and in turn, the
+response \(y_l\) is drawn from a univariate distribution with parameter
+\(\theta_l\). Since the latent vectors of each \(A^{(l)}\) are presumed
+to be distributed on a structured subset of the latent space, e.g., a
+manifold, we call this the multiple latent structure model (MLSM). Much
+previous work has been done on multilayer networks, such as the
+multilayer degree corrected block model of \citet{agterberg2022joint},
+multiple random eigen graphs of \citet{8889404}, and the multiple common
+subspace independent edge graphs of \citet{arroyo2020inference}. In each
+of these models, the main inference problem is to identify parameters
+that are shared across multiple graphs. Our contribution is in applying
+similar methods to identify parameters specific to each graph, which are
+in turn used as features for supervised learning.
+
+The following sections are organized as follows: Section 2 introduces
+the structured and parameterized
 
@@
 \section{Methods}\label{methods}}
+
+\hypertarget{notation-and-scope}{%
+\subsection{Notation and Scope}\label{notation-and-scope}}
+
+Let \(G_1, ..., G_L\) be a collection of undirected graphs where each
+\(G_l = (V_l, E_l)\). In this setup, each graph can have its own vertex
+and edge sets. Each graph is represented by adjacency matrices
+\(A^{(1)}, ..., A^{(L)}\), where each
+\(A^{(l)} \in \mathbb{R}^{n_l \times n_l}\) is a symmetric and hollow
+matrix such that \(A^{(l)}_{ij}\) represents the edge weight between
+vertices \(i\) and \(j\), and \(n_l\) represents the size of the vertex
+set of the \(l\)\textsuperscript{th} graph. Let
+\(P^{(1)}, ..., P^{(L)}\) be the corresponding collection of edge
+parameter matrices such that each
+\(A^{(l)}_{ij} \indep H(P^{(l)}_{ij})\) for each
+\(1 \leq i < j \leq n_l\) and \(H\) is a specified probability
+distribution such that \(E[A^{(l)}_{ij}] = P^{(l)}_{ij}\). For ease of
+notation, we use \(A^{(l)} \sim H(P^{(l)})\) for an adjacency matrix
+\(A^{(l)}\) drawn from edge parameter matrix \(P^{(l)}\) in this way.
+Denote \(X^{(1)}, ..., X^{(L)}\) as the sets of latent vectors
+corresponding to each graph under the GRDPG framework. Each
+\(X^{(l)} \in \mathbb{R}^{n_l \times d}\), i.e., the graphs share the
+same latent dimension, and \(x_i^{(l)} \in \mathbb{R}^d\) is the latent
+vector corresponding to the \(i\)\textsuperscript{th} vertex in the
+\(l\)\textsuperscript{th} graph.
 
@@
 is what we observe in these data (fig \ref{fig:mbconnectome-beta}).
+Using the three subgraphs of the left hemisphere as a training set to
+classify which subgraph of the right hemisphere belongs to which neuron
+type using the extracted Beta distribution parameter estimates as
+predictors results in 100\% accuracy.
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{draft_files/figure-latex/cosine-heatmap-1} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{draft_files/figure-latex/cosine-heatmap-2-1} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{draft_files/figure-latex/cosine-heatmap-2-2} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{draft_files/figure-latex/cosine-heatmap-2-3} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{draft_files/figure-latex/cosine-heatmap-2-4} \end{center}
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5915}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1690}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2394}}@{}}
+\begin{longtable}[]{@{}lrl@{}}
 \caption{Correlation between age and various graph
-metrics}\tabularnewline
+metrics.\label{tab:cor-comparison}}\tabularnewline
 \toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Metric
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-Correlation
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-95\% conf. int.
-\end{minipage} \\
+Metric & Correlation & 95\% conf. int. \\
 \midrule\noalign{}
@@
 \toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-Metric
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-Correlation
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-95\% conf. int.
-\end{minipage} \\
+Metric & Correlation & 95\% conf. int. \\
 \midrule\noalign{}
@@
 Angle between hemispheres & 0.558 & (0.496, 0.615) \\
-Edge connectivity within left hemisphere & 0.241 & (0.158, 0.321) \\
-Edge connectivity within right hemisphere & 0.243 & (0.16, 0.322) \\
-Edge connectivity between hemispheres & -0.436 & (-0.503, -0.363) \\
 Degree within hemisphere & 0.436 & (0.363, 0.503) \\
 Degree between hemispheres & -0.499 & (-0.561, -0.431) \\
-Number of triangles in left hemisphere & -0.025 & (-0.111, 0.062) \\
-Number of triangles in right hemisphere & 0.002 & (-0.084, 0.088) \\
-Assortativity w.r.t. hemisphere & 0.436 & (0.363, 0.503) \\
 Modularity w.r.t. hemisphere & 0.434 & (0.362, 0.502) \\

### 2024-05-02 simulations/angle-regression-sim.R
fixed issues with some simulations

@@ -11,10 +11,10 @@ nrep <- 50
 a.vec <- c(1, 1)
 b.vec <- c(1, 1)
 theta.min <- pi / 6
-theta.max <- pi / 3
+theta.max <- pi / 2
 beta1 <- 1
 beta0 <- 2
-sigma_eps <- 1
+sigma_eps <- .5
 
 N <- 2 ^ 6
 n.vec <- 2 ^ c(10, 9, 8, 7, 6)
@@ -34,53 +34,62 @@ out.df <- foreach::foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
       datamat.df <- foreach::foreach(i = seq(N), .combine = dplyr::bind_rows, .errorhandling = 'remove') %dopar% {
         set.seed(i + (nrep - 1) * rep)
         print(paste(i, '/', N))
-        iter.filename <- paste0('angle-reg-n-', n, '-rep-', rep, '-iter-', i, '.csv')
-        if (iter.filename %in% dir(sim.dir)) {
-          one.iter.df <- readr::read_csv(file.path(sim.dir, iter.filename),
-                                         progress = FALSE,
-                                         show_col_types = FALSE)
-        } else {
-          z <- sample(seq(2), n, replace = TRUE)
-          t. <- rbeta(n, a.vec[z], b.vec[z])
-          theta <- runif(1, theta.min, theta.max)
-          y <- beta0 + beta1 * theta + rnorm(1, 0, sigma_eps)
-          x1 <- ifelse(z == 1, t., cos(theta) * t.)
-          x2 <- ifelse(z == 1, 0, sin(theta) * t.)
-          X <- cbind(x1, x2)
-          P <- X %*% t(X)
-          A <- draw.graph(P)
-          Xhat <- embedding(A, 2, 0)
-          clustering <- manifold.clustering(Xhat, 
-                                            degree = 1,
-                                            initialization = z,
-                                            normalize = TRUE,
-                                            intercept = FALSE)
-          angle.1 <- atan2(clustering$p[[1]][2, 2],
-                           clustering$p[[1]][2, 1])
-          angle.2 <- atan2(clustering$p[[2]][2, 2],
-                           clustering$p[[2]][2, 1])
-          theta.hat <- abs(angle.2 - angle.1)
-          if (theta.hat > pi) {
-            theta.hat <- 2 * pi - theta.hat
-          }
-          one.iter.df <- dplyr::tibble(y = y,
-                                       theta = theta,
[…]

### 2024-05-03 simulations/beta-regression-sim.R
updated simulations

@@ -11,10 +11,10 @@ nrep <- 50
 beta0 <- 1
 beta1 <- 2
 beta2 <- -1
-sigma_eps <- 1
+sigma_eps <- .5
 
 N <- 2 ^ 6
-n.vec <- 2 ^ c(9, 8, 7, 6)
+n.vec <- 2 ^ c(10, 9, 8, 7, 6)
 
 sim.dir <- '~/dev/multilayer-rdpg/simulations/beta-regression'
 
@@ -31,109 +31,72 @@ out.df <- foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
   print(paste('n =', n))
   foreach(rep = seq(nrep), .combine = dplyr::bind_rows) %do% {
     print(paste('rep', rep, 'out of', nrep))
-    rep.filename <- paste0('beta-reg-n-', n, '-rep-', rep, '.csv')
-    if (rep.filename %in% dir(sim.dir)) {
-      one.rep.df <- readr::read_csv(file.path(sim.dir, rep.filename),
-                                    progress = FALSE,
-                                    show_col_types = FALSE)
-    } else {
-      datamat.df <- plyr::ldply(seq(N), function(i) {
-        set.seed(i + (rep - 1) * N)
-        print(paste(i, '/', N))
-        iter.filename <- paste0('beta-reg-n-', n, '-rep-', rep, '-iter-', i, '.csv')
-        if (iter.filename %in% dir(sim.dir)) {
-          one.iter.df <- readr::read_csv(file.path(sim.dir, iter.filename),
-                                         progress = FALSE,
-                                         show_col_types = FALSE)
-        } else {
-          a <- runif(1, 1, 2)
-          b <- runif(1, 1, 2)
-          y <- beta0 + beta1 * a + beta2 * b + rnorm(1, 0, sigma_eps)
-          t. <- rbeta(n, a, b)
-          x1 <- t. ^ 2
-          x2 <- 2 * t. * (1 - t.)
-          X <- cbind(x1, x2)
-          P <- X %*% t(X)
-          A <- draw.graph(P)
-          Xhat <- embedding(A, 2, 0)
-          curve.est <- estimate.bezier.curve.2(Xhat,
-                                               degree = 2,
-                                               intercept = FALSE,
-                                               initialization = 'isomap',
-                                               # min.t = 0, max.t = 1,
-                                               normalize = TRUE,
-                                               parallel = parallelize.curvefit)
-          t.hat <- curve.est$t
[…]

### 2024-05-03 draft.Rmd
updated draft with new simulation results

@@
 \begin{theorem}
+\ref{thm:supervised}
 \end{theorem}
@@
 # Simulation Study
+
+In this section, we evaluate algorithm \ref{alg:mlsm-fit} and theorem \ref{thm:supervised} empirically via simulation and compare to existing methods. 
+Two settings are examined. In the first, the data generating model reduces to a simpler model and can be adequately described by the COSIE model or multilayer DCBM. 
+The second setting involves a nonlinear latent structure that can be described as a polynomial curve and thus does not reduce to a simpler model or be analyzed by their corresponding methods.
+
+The scenarios are as follows:
+
+1. In this setting, there are two latent structures, described by two rays emanating from the origin.
+
+2.
+
+## Regression
+
+In this scenario, the responses $y_\ell$ are generated as linear responses of $\theta_\ell$ (in scenario 1) or $a_\ell$ and $b_\ell$ (in scenario 2). 
+More specifically, in scenario 1, each $y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 \theta_\ell, \sigma^2)$, and in scenario 2, each $y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 a_\ell + \beta_2 b_\ell, \sigma^2)$.
+
+```{r angle-regression-sim, fig.width = 6, cache = TRUE, fig.cap = 'The average MSE over 50 simulations in scenario 1. The errorbars are the standard error.'}
+
+angle.reg.df <- readr::read_csv('~/dev/multilayer-rdpg/simulations/angle-regression/angle-regression-sim.csv')
+
+n.vec <- unique(angle.reg.df$n)
+
+angle.reg.df %>% 
+  dplyr::group_by(n) %>% 
+  dplyr::summarise(mse.theta = mean(mse),
+                   se.theta = sd(mse) / dplyr::n(),
+                   se.m = sd(mse.m) / dplyr::n(),
+                   mse.m = mean(mse.m),
+                   se.a = sd(mse.a) / dplyr::n(),
+                   mse.a = mean(mse.a)) %>% 
+  ggplot() + 
+  theme_bw() + 
+  scale_colour_brewer(palette = 'Set1') +
+  labs(x = 'vertices per graph', y = 'MSE', colour = 'predictor') +
+  geom_line(aes(x = n, y = mse.theta, colour = 'estimated angle')) + 
+  geom_errorbar(aes(x = n, 
+                    ymin = mse.theta - se.theta, 
+                    ymax = mse.theta + se.theta,
+                    colour = 'estimated angle'), 
+                width = .1) + 
+  geom_line(aes(x = n, y = mse.m, colour = 'modularity')) + 
+  geom_errorbar(aes(x = n,
+                    ymin = mse.m - se.m,
+                    ymax = mse.m + se.m,
+                    colour = 'modularity'), 
+                width = .1) +
+
+  geom_line(aes(x = n, y = mse.a, colour = 'assortativity')) + 
+  geom_errorbar(aes(x = n,
+                    ymin = mse.a - se.a,
+                    ymax = mse.a + se.a,
+                    colour = 'assortativity'), 
+                width = .1) + 
+  scale_x_log10(breaks = sort(n.vec))
+```
+
+```{r theta-mse, fig.width = 6, cache = TRUE, fig.cap = 'The average MSE of the estimated angles in scenario 1 over 50 simulations. The errorbars are the standard error.'}
+angle.reg.df %>% 
+  dplyr::group_by(n) %>% 
+  dplyr::summarise(theta.se = sd(theta.mse) / dplyr::n(),
+                   theta.mse = mean(theta.mse)) %>% 
+  ggplot() + 
+  geom_line(aes(x = n, y = theta.mse)) + 
+  geom_errorbar(aes(x = n, ymin = theta.mse - theta.se, ymax = theta.mse + theta.se),
+                width = .1) + 
+  labs(x = 'number of vertices', y = expression('MSE('~hat(theta)~')')) + 
+  theme_bw() + 
+  scale_x_log10(breaks = sort(n.vec))
+```
+
+```{r beta-regression-sim, fig.width = 6, cache = TRUE, fig.cap = 'The average MSE over 50 simulations in scenario 2. The errorbars are the standard error.'}
+
+beta.reg.df <- readr::read_csv('~/dev/multilayer-rdpg/simulations/beta-regression/beta-regression-sim.csv')
+
+n.vec <- unique(beta.reg.df$n)
+
+beta.reg.df %>% 
+  dplyr::group_by(n) %>% 
+  dplyr::summarise(mse.beta = mean(mse),
+                   se.beta = sd(mse) / sqrt(dplyr::n()),
+                   se.d = sd(mse.d) / sqrt(dplyr::n()),
+                   mse.d = mean(mse.d),
+                   se.t = sd(mse.t) / sqrt(dplyr::n()),
+                   mse.t = mean(mse.t)) %>% 
+  ggplot() + 
+  geom_line(aes(x = n, y = mse.beta, colour = 'estimated beta parameters')) + 
+  geom_errorbar(aes(x = n, 
+                    ymin = mse.beta - se.beta, 
+                    ymax = mse.beta + se.beta,
+                    colour = 'estimated beta parameters'),
+                width = .1) + 
+  geom_line(aes(x = n, y = mse.d, colour = 'average degree')) + 
+  geom_errorbar(aes(x = n,
+                    ymin = mse.d - se.d,
+                    ymax = mse.d + se.d,
+                    colour = 'average degree'),
+                width = .1) +
+
+  geom_line(aes(x = n, y = mse.t, colour = 'transitivity')) + 
+  geom_errorbar(aes(x = n,
+                    ymin = mse.t - se.t,
+                    ymax = mse.t + se.t,
+                    colour = 'transitivity'),
+                width = .1) +
+  theme_bw() + 
+  scale_colour_brewer(palette = 'Set1') +
+  scale_x_log10(breaks = sort(n.vec)) + 
+  labs(x = 'number of vertices', y = 'MSE', colour = 'predictor')
+```
+
+```{r beta-mse, fig.width = 6, cache = TRUE, fig.cap = 'The average MSE of the estimated beta parameters, $(a, b)$, in scenario 2 over 50 simulations. The errorbars are the standard error.'}
+beta.reg.df %>% 
+  dplyr::group_by(n) %>% 
+  dplyr::summarise(a.se = sd(a.mse) / sqrt(dplyr::n()),
+                   b.se = sd(b.mse) / sqrt(dplyr::n()),
+                   a.mse = mean(a.mse),
+                   b.mse = mean(b.mse)) %>% 
+  ggplot() + 
+  # scale_y_log10() + 
+  scale_x_log10() + 
+  geom_line(aes(x = n, y = a.mse, colour = 'a')) + 
+  geom_errorbar(aes(x = n, ymin = a.mse - a.se, ymax = a.mse + a.se, colour = 'a'),
+                width = .1) + 
+  geom_line(aes(x = n, y = b.mse, colour = 'b')) + 
+  geom_errorbar(aes(x = n, ymin = b.mse - b.se, ymax = b.mse + b.se, colour = 'b'),
+                width = .1) + 
+  theme_bw() + 
+  labs(x = 'number of vertices', y = 'MSE of Beta parameter estimates', 
+       colour = 'Beta \ndistribution \nparameter') + 
+  scale_colour_brewer(palette = 'Set1') +
+  scale_x_log10(breaks = sort(n.vec))
+```
 

### 2024-05-03 draft.tex
updated draft with new simulation results

@@
 \section{Simulation Study}\label{simulation-study}}
+
+In this section, we evaluate algorithm \ref{alg:mlsm-fit} and theorem
+\ref{thm:supervised} empirically via simulation and compare to existing
+methods. Two settings are examined. In the first, the data generating
+model reduces to a simpler model and can be adequately described by the
+COSIE model or multilayer DCBM. The second setting involves a nonlinear
+latent structure that can be described as a polynomial curve and thus
+does not reduce to a simpler model or be analyzed by their corresponding
+methods.
+
+The scenarios are as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  In this setting, there are two latent structures, described by two
+  rays emanating from the origin.
+\item
+\end{enumerate}
+
+\hypertarget{regression}{%
+\subsection{Regression}\label{regression}}
+
+In this scenario, the responses \(y_\ell\) are generated as linear
+responses of \(\theta_\ell\) (in scenario 1) or \(a_\ell\) and
+\(b_\ell\) (in scenario 2). More specifically, in scenario 1, each
+\(y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 \theta_\ell, \sigma^2)\),
+and in scenario 2, each
+\(y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 a_\ell + \beta_2 b_\ell, \sigma^2)\).
+
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/angle-regression-sim-1}
+
+}
+
+\caption{The average MSE over 50 simulations in scenario 1. The errorbars are the standard error.}\label{fig:angle-regression-sim}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/theta-mse-1}
+
+}
+
+\caption{The average MSE of the estimated angles in scenario 1 over 50 simulations. The errorbars are the standard error.}\label{fig:theta-mse}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/beta-regression-sim-1}
+
+}
+
+\caption{The average MSE over 50 simulations in scenario 2. The errorbars are the standard error.}\label{fig:beta-regression-sim}
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{draft_files/figure-latex/beta-mse-1}
+
+}
+
+\caption{The average MSE of the estimated beta parameters, $(a, b)$, in scenario 2 over 50 simulations. The errorbars are the standard error.}\label{fig:beta-mse}
+\end{figure}
 
@@
 
-\hypertarget{regression}{%
-\subsection{Regression}\label{regression}}
+\hypertarget{regression-1}{%
+\subsection{Regression}\label{regression-1}}
 

### 2024-05-03 simulations/angle-regression-sim.R
minor changes to simulation

@@ -113,10 +113,10 @@ ggplot(out.df) +
 out.df %>% 
   dplyr::group_by(n) %>% 
   dplyr::summarise(mse.theta = mean(mse),
-                   se.theta = sd(mse) / dplyr::n(),
-                   se.m = sd(mse.m) / dplyr::n(),
+                   se.theta = sd(mse) / sqrt(dplyr::n()),
+                   se.m = sd(mse.m) / sqrt(dplyr::n()),
                    mse.m = mean(mse.m),
-                   se.a = sd(mse.a) / dplyr::n(),
+                   se.a = sd(mse.a) / sqrt(dplyr::n()),
                    mse.a = mean(mse.a)) %>% 
   ggplot() + 
   geom_line(aes(x = n, y = mse.theta, colour = 'theta')) + 

### 2024-05-03 draft.Rmd
minor updates

--- 2024-04-19 draft.Rmd
@@
 # Simulation Study
+
+In this section, we evaluate algorithm \ref{alg:mlsm-fit} and theorem \ref{thm:supervised} empirically via simulation and compare to existing methods. 
+Two settings are examined. In the first, the data generating model reduces to a simpler model and can be adequately described by the COSIE model or multilayer DCBM. 
+The second setting involves a nonlinear latent structure that can be described as a polynomial curve and thus does not reduce to a simpler model or be analyzed by their corresponding methods.
+
+The scenarios are as follows:
+
+1. In this setting, there are two latent structures, described by two rays emanating from the origin.
+
+2.
+
+## Regression
+
+In this scenario, the responses $y_\ell$ are generated as linear responses of $\theta_\ell$ (in scenario 1) or $a_\ell$ and $b_\ell$ (in scenario 2). 
+More specifically, in scenario 1, each $y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 \theta_\ell, \sigma^2)$, and in scenario 2, each $y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 a_\ell + \beta_2 b_\ell, \sigma^2)$.
+
+```{r angle-regression-sim, fig.width = 6, cache = FALSE, fig.cap = 'The average MSE over 50 simulations in scenario 1. The errorbars are the standard error.'}
+n.sigma <- 1
+
+angle.reg.df <- readr::read_csv('~/dev/multilayer-rdpg/simulations/angle-regression/angle-regression-sim.csv')
+
+n.vec <- unique(angle.reg.df$n)
+
+angle.reg.df %>% 
+  dplyr::group_by(n) %>% 
+  dplyr::summarise(mse.theta = mean(mse),
+                   se.theta = sd(mse) / sqrt(dplyr::n()) * n.sigma,
+                   se.m = sd(mse.m) / sqrt(dplyr::n()) * n.sigma,
+                   mse.m = mean(mse.m),
+                   se.a = sd(mse.a) / sqrt(dplyr::n()) * n.sigma,
+                   mse.a = mean(mse.a)) %>% 
+  ggplot() + 
+  theme_bw() + 
+  scale_colour_brewer(palette = 'Set1') +
+  labs(x = 'vertices per graph', y = 'MSE', colour = 'predictor') +
+  geom_line(aes(x = n, y = mse.theta, colour = 'estimated angle')) + 
+  geom_errorbar(aes(x = n, 
+                    ymin = mse.theta - se.theta, 
+                    ymax = mse.theta + se.theta,
+                    colour = 'estimated angle'), 
+                width = .1) + 
+  geom_line(aes(x = n, y = mse.m, colour = 'modularity')) + 
+  geom_errorbar(aes(x = n,
+                    ymin = mse.m - se.m,
+                    ymax = mse.m + se.m,
+                    colour = 'modularity'), 
+                width = .1) +
+
+  geom_line(aes(x = n, y = mse.a, colour = 'assortativity')) + 
+  geom_errorbar(aes(x = n,
+                    ymin = mse.a - se.a,
+                    ymax = mse.a + se.a,
+                    colour = 'assortativity'), 
+                width = .1) + 
+  scale_x_log10(breaks = sort(n.vec))
+```
+
+```{r theta-mse, fig.width = 6, cache = FALSE, fig.cap = 'The average MSE of the estimated angles in scenario 1 over 50 simulations. The errorbars are the standard error.'}
+angle.reg.df %>% 
+  dplyr::group_by(n) %>% 
+  dplyr::summarise(theta.se = sd(theta.mse) / sqrt(dplyr::n()) * n.sigma,
+                   theta.mse = mean(theta.mse)) %>% 
+  ggplot() + 
+  geom_line(aes(x = n, y = theta.mse)) + 
+  geom_errorbar(aes(x = n, ymin = theta.mse - theta.se, ymax = theta.mse + theta.se),
+                width = .1) + 
+  labs(x = 'number of vertices', y = expression('MSE('~hat(theta)~')')) + 
+  theme_bw() + 
+  scale_y_log10() + 
+  scale_x_log10(breaks = sort(n.vec))
+```
+
+```{r beta-regression-sim, fig.width = 6, cache = FALSE, fig.cap = 'The average MSE over 50 simulations in scenario 2. The errorbars are the standard error.'}
+
+beta.reg.df <- readr::read_csv('~/dev/multilayer-rdpg/simulations/beta-regression/beta-regression-sim.csv')
+
+n.vec <- unique(beta.reg.df$n)
+
+beta.reg.df %>% 
+  dplyr::group_by(n) %>% 
+  dplyr::summarise(mse.beta = mean(mse),
+                   se.beta = sd(mse) / sqrt(dplyr::n()) * n.sigma,
+                   se.d = sd(mse.d) / sqrt(dplyr::n()) * n.sigma,
+                   mse.d = mean(mse.d),
+                   se.t = sd(mse.t) / sqrt(dplyr::n()) * n.sigma,
+                   mse.t = mean(mse.t)) %>% 
+  ggplot() + 
+  geom_line(aes(x = n, y = mse.beta, colour = 'estimated beta parameters')) + 
+  geom_errorbar(aes(x = n, 
+                    ymin = mse.beta - se.beta, 
+                    ymax = mse.beta + se.beta,
+                    colour = 'estimated beta parameters'),
+                width = .1) + 
+  geom_line(aes(x = n, y = mse.d, colour = 'average degree')) + 
+  geom_errorbar(aes(x = n,
+                    ymin = mse.d - se.d,
+                    ymax = mse.d + se.d,
+                    colour = 'average degree'),
+                width = .1) +
+
+  geom_line(aes(x = n, y = mse.t, colour = 'transitivity')) + 
+  geom_errorbar(aes(x = n,
+                    ymin = mse.t - se.t,
+                    ymax = mse.t + se.t,
+                    colour = 'transitivity'),
+                width = .1) +
+  theme_bw() + 
+  scale_colour_brewer(palette = 'Set1') +
+  scale_x_log10(breaks = sort(n.vec)) + 
+  labs(x = 'number of vertices', y = 'MSE', colour = 'predictor')
+```
+
+```{r beta-mse, fig.width = 6, cache = FALSE, fig.cap = 'The average MSE of the estimated beta parameters, $(a, b)$, in scenario 2 over 50 simulations. The errorbars are the standard error.'}
+beta.reg.df %>% 
+  dplyr::group_by(n) %>% 
+  dplyr::summarise(a.se = sd(a.mse) / sqrt(dplyr::n()) * n.sigma,
+                   b.se = sd(b.mse) / sqrt(dplyr::n()) * n.sigma,
+                   a.mse = mean(a.mse),
+                   b.mse = mean(b.mse)) %>% 
+  ggplot() + 
+  # scale_y_log10() + 
+  scale_x_log10() + 
+  geom_line(aes(x = n, y = a.mse, colour = 'a')) + 
+  geom_errorbar(aes(x = n, ymin = a.mse - a.se, ymax = a.mse + a.se, colour = 'a'),
+                width = .1) + 
+  geom_line(aes(x = n, y = b.mse, colour = 'b')) + 
+  geom_errorbar(aes(x = n, ymin = b.mse - b.se, ymax = b.mse + b.se, colour = 'b'),
+                width = .1) + 
+  theme_bw() + 
+  labs(x = 'number of vertices', y = 'MSE of Beta parameter estimates', 
+       colour = 'Beta \ndistribution \nparameter') + 
+  scale_colour_brewer(palette = 'Set1') +
+  scale_y_log10() + 
+  scale_x_log10(breaks = sort(n.vec))
+```
 

### 2024-09-19 draft.Rmd
minor changes

@@
 angle.reg.df %>% 
+  dplyr::filter(n >= 128) %>% 
   dplyr::group_by(n) %>% 
@@
                 width = .1) + 
-  geom_line(aes(x = n, y = mse.m, colour = 'modularity')) + 
-  geom_errorbar(aes(x = n,
-                    ymin = mse.m - se.m,
-                    ymax = mse.m + se.m,
-                    colour = 'modularity'), 
-                width = .1) +
-
-  geom_line(aes(x = n, y = mse.a, colour = 'assortativity')) + 
+  # geom_line(aes(x = n, y = mse.m, colour = 'modularity')) + 
+  # geom_errorbar(aes(x = n,
+  #                   ymin = mse.m - se.m,
+  #                   ymax = mse.m + se.m,
+  #                   colour = 'modularity'), 
+  #               width = .1) + 
+  geom_line(aes(x = n, y = mse.a, colour = 'graph statistics')) + 
   geom_errorbar(aes(x = n,
@@
                     ymax = mse.a + se.a,
-                    colour = 'assortativity'), 
+                    colour = 'graph statistics'), 
                 width = .1) + 
@@
                 width = .1) + 
-  geom_line(aes(x = n, y = mse.d, colour = 'average degree')) + 
+  geom_line(aes(x = n, y = mse.d, colour = 'graph statistics')) + 
   geom_errorbar(aes(x = n,
@@
                     ymax = mse.d + se.d,
-                    colour = 'average degree'),
+                    colour = 'graph statistics'),
                 width = .1) + 
-
-  geom_line(aes(x = n, y = mse.t, colour = 'transitivity')) + 
-  geom_errorbar(aes(x = n,
-                    ymin = mse.t - se.t,
-                    ymax = mse.t + se.t,
-                    colour = 'transitivity'),
-                width = .1) +
+  # geom_line(aes(x = n, y = mse.t, colour = 'transitivity')) + 
+  # geom_errorbar(aes(x = n,
+  #                   ymin = mse.t - se.t,
+  #                   ymax = mse.t + se.t,
+  #                   colour = 'transitivity'),
+  #               width = .1) +
   theme_bw() + 

### 2024-09-19 draft.tex
minor changes

@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{notation-and-scope}{%
-\subsection{Notation and Scope}\label{notation-and-scope}}
+\section{Methods}\label{methods}
+
+\subsection{Notation and Scope}\label{notation-and-scope}
 
@@
 
-\hypertarget{definitions-and-models}{%
-\subsection{Definitions and Models}\label{definitions-and-models}}
+\subsection{Definitions and Models}\label{definitions-and-models}
 
@@
 
-\hypertarget{connections-to-related-models}{%
 \subsubsection{Connections to Related
-Models}\label{connections-to-related-models}}
+Models}\label{connections-to-related-models}
 
@@
 
-\hypertarget{classification-and-regression-on-the-mlsm}{%
 \subsubsection{Classification and Regression on the
-MLSM}\label{classification-and-regression-on-the-mlsm}}
+MLSM}\label{classification-and-regression-on-the-mlsm}
 
@@
 
-\hypertarget{main-results}{%
-\subsection{Main Results}\label{main-results}}
+\subsection{Main Results}\label{main-results}
 
@@
 
-\hypertarget{estimation-of-the-lsm-curves}{%
 \subsubsection{Estimation of the LSM
-Curves}\label{estimation-of-the-lsm-curves}}
+Curves}\label{estimation-of-the-lsm-curves}
 
@@
 
-\hypertarget{recovery-of-the-underlying-distributions}{%
 \subsubsection{Recovery of the Underlying
-Distributions}\label{recovery-of-the-underlying-distributions}}
+Distributions}\label{recovery-of-the-underlying-distributions}
 
@@
 
-\hypertarget{classification-and-regression}{%
 \subsubsection{Classification and
-Regression}\label{classification-and-regression}}
+Regression}\label{classification-and-regression}
 
@@
 
-\hypertarget{simulation-study}{%
-\section{Simulation Study}\label{simulation-study}}
+\section{Simulation Study}\label{simulation-study}
 
@@
 
-\hypertarget{regression}{%
-\subsection{Regression}\label{regression}}
+\subsection{Regression}\label{regression}
 
@@
 
-\hypertarget{classification}{%
-\subsection{Classification}\label{classification}}
+\subsection{Classification}\label{classification}
 
@@
 
-\hypertarget{regression-1}{%
-\subsection{Regression}\label{regression-1}}
+\subsection{Regression}\label{regression-1}
 
@@
 
-\hypertarget{applications}{%
-\section{Applications}\label{applications}}
-
-\hypertarget{drosophila-connectome}{%
-\subsection{Drosophila Connectome}\label{drosophila-connectome}}
+\section{Applications}\label{applications}
+
+\subsection{Drosophila Connectome}\label{drosophila-connectome}
 
@@
 
-\hypertarget{human-connectome-project-aging-study}{%
 \subsection{Human Connectome Project Aging
-Study}\label{human-connectome-project-aging-study}}
+Study}\label{human-connectome-project-aging-study}
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\section{Discussion}\label{discussion}
 

### 2024-09-19 hcp-a-eda.R

@@ -28,8 +28,21 @@ A <- dat.NumFibers[[2]] %>%
 A <- (A + t(A)) / 2
 diag(A) <- 0
 
+# A <- log(A)
+# A[is.infinite(A)] <- 0
+# A[A < 0] <- 0
+# 
+# A <- log(A + 1)
+# 
+# A <- sqrt(A)
+
 qgraph::qgraph(A, layout = 'spring', groups = dti.info$hemisphere)
 Xhat <- embedding(A, 2, 0, scale = TRUE)
+plot(Xhat, asp = 1, col = z)
+
+# Xhat <- embedding(A, 6, 3, scale = TRUE)
+# pairs(Xhat, asp = 1, col = z)
+
 clustering <- manifold.clustering(Xhat, 
                                   degree = 1,
                                   initialization = as.numeric(z),

### 2024-09-20 simulations/beta-classification/beta-classification-sim.R
minor updates

@@ -1,0 +1,114 @@
+import::from(magrittr, `%>%`)
+import::from(foreach, foreach, `%do%`, `%dopar%`)
+source('~/dev/pabm-grdpg/functions.R')
+source('~/dev/manifold-block-models/functions.R')
+library(ggplot2)
+
+doMC::registerDoMC(parallel::detectCores())
+
+nrep <- 50
+
+beta0 <- 1
+beta1 <- 2
+beta2 <- -1
+sigma_eps <- .5
+
+K <- 2
+a.vec <- c(2, 1)
+b.vec <- c(2, .839214517857)
+
+N <- 2 ^ 6
+n.vec <- 2 ^ c(11, 10, 9, 8)
+
+sim.dir <- '~/dev/multilayer-rdpg/simulations/beta-classification'
+
+out.df <- foreach::foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
+  print(paste('n = ', n))
+  foreach::foreach(rep = seq(nrep), .combine = dplyr::bind_rows) %do% {
+    print(paste('rep', rep, 'out of', nrep))
+    z.vec <- sample(seq(K), N, replace = TRUE)
+    # a.vec <- runif(2, .25, 3)
+    # b.vec <- runif(2, .25, 3)
+    datamat.df <- foreach::foreach(i = seq(N), .combine = dplyr::bind_rows, .errorhandling = 'remove') %dopar% {
+      z <- z.vec[i]
+      a <- a.vec[z]
+      b <- b.vec[z]
+      n.i <- rpois(1, n)
+      t. <- rbeta(n.i, a, b)
+      x1 <- t. ^ 2
+      x2 <- 2 * t. * (1 - t.)
+      X <- cbind(x1, x2)
+      P <- X %*% t(X)
+      A <- draw.graph(P)
+      Xhat <- embedding(A, 2, 0)
+      
+      p.init <- matrix(c(0, 0, 
+                         Xhat[which.max(apply(Xhat, 1, function(x) sum(x ^ 2))), ]),
+                       ncol = 2, 
+                       byrow = TRUE)
+      curve.est <- estimate.bezier.curve.2(Xhat,
+                                           degree = 2,
+                                           intercept = FALSE,
+                                           # initialization = 'isomap',
+                                           init.params = p.init,
+                                           # initialization = 'x',
+                                           # min.t = 0, max.t = 1,
[…]

### 2024-09-20 simulations/beta-classification/beta-classification-sim.R
minor correction

@@ -111,4 +111,4 @@ out.df %>%
                     ymax = acc.stats.mean + 2 * acc.stats.se,
                     colour = 'stats'))
 
-readr::write_csv(out.df, file.path(sim.dir, 'beta.classification-sim.csv'))
+readr::write_csv(out.df, file.path(sim.dir, 'beta-classification-sim.csv'))

### 2024-11-15 draft.Rmd
first draft of paper

--- 2024-05-03 draft.Rmd
@@
 In particular, it has been shown that the RDPG describes a wide range of popular random graph models with rigid latent structures.
-More recently, joint modeling of mutliple random graphs that share common properties or structures across graphs have been introduced, such as the multilayer RDPG, multiple RPDG, and multilayer stochastic block model.
-In this work, we use these joint random graph models in the context of statistical learning, such as classification and regression, by introducing the multiple latent structure model, in which the graphs share a common latent structure with different parameters that correspond to different response variables.
+More recently, joint modeling of mutliple random graphs that share common properties or structures across graphs have been introduced, such as the multilayer RDPG, multiple RPDG, and common subspace independent edge model.
+In this work, we use these joint random graph models in the context of graph classification and regression by introducing the multiple latent structure model, in which the graphs share a common latent structure with different parameters that correspond to different response variables.
 Then we propose various estimation techniques involving manifold learning to estimate these parameters and in turn predict the responses, with theorems guaranteeing convergence of the predictions.
@@
 One approach is to treat the network as a random graph that comes from some probability model. 
-In particular, if we sum up the network in an adjacency matrix $A \in \mathbb{R}^{n \times n}$, then one probability model might be to draw each element $A_{ij}$, which represents the existence of an edge or the edge weight from vertex $i$ to $j$, independently from some distribution, perhaps with a unique parameter for the pair $(i,j)$, e.g., $A_{ij} \indep F_{\theta_{ij}}$.
+One such approach in particular is the independent edge approach.
+If we sum up the network in an adjacency matrix $A \in \mathbb{R}^{n \times n}$, then this model draws each element $A_{ij}$, which represents the existence of an edge or the edge weight from vertex $i$ to $j$, independently from some distribution, perhaps with a unique parameter for the pair $(i,j)$, e.g., $A_{ij} \indep F_{\theta_{ij}}$.
 The classical example of this is the $\ER$ model \citep{Gilbert:1959}, in which every edge is drawn from the same distribution, typically a Bernoulli distribution, using the same parameter, i.e., $A_{ij} \iid \Bernoulli (p)$.
@@
 On the other hand, the classical $\ER$ model is often too restrictive to describe real, observed networks. 
-Much work has been done to develop models that are constrained enough for robust statistical inference while being generalizable to describe a wide range of networks.
+Various approaches have been taken to restrict $P$, such as by setting the matrix rank of $P$ to a small value. 
 One such family of random graph models is the random dot product graph (RDPG), first proposed by \citet{10.1007/978-3-540-77004-6_11}, which is a type of latent space graph in which each vertex of the graph has a corresponding latent vector in a low-dimensional Euclidean space $\mathbb{R}^d$, and the edge parameter between each pair of vertices is determined by the dot product of the corresponding vectors.
@@
 
-In this paper, we apply these highly structured GRDPG models to supervised learning problems in which there are multiple graphs, $G_1, ..., G_L$, each with a response $y^{(1)}, ..., y^{(L)}$. 
-Each graph, represented by adjacency matrix $A^{(l)}$, is presumed to be drawn from a GRDPG with parameter set $\theta_l$, and in turn, the response $y_l$ is drawn from a univariate distribution with parameter $\theta_l$.
-Since the latent vectors of each $A^{(l)}$ are presumed to be distributed on a structured subset of the latent space, e.g., a manifold, we call this the multiple latent structure model (MLSM). 
-Much previous work has been done on multilayer networks, such as the multilayer degree corrected block model of \citet{agterberg2022joint}, multiple random eigen graphs of \citet{8889404}, and the multiple common subspace independent edge graphs of \citet{arroyo2020inference}.
+In this paper, we apply these highly structured random dot product graph models to supervised learning problems in which there are multiple graphs, $G_1, ..., G_L$, each with a response $y^{(1)}, ..., y^{(L)}$.
+Each graph, represented by adjacency matrix $A^{(\ell)}$, is presumed to be drawn from a RDPG with parameter set $\theta_l$, and in turn, the response $y_l$ is drawn from a univariate distribution with parameter $\theta_l$.
+Since the latent vectors of each $A^{(\ell)}$ are presumed to be distributed on a structured subset of the latent space, e.g., a manifold, we call this the multiple latent structure model (MLSM). 
+Much previous work has been done on multilayer networks, such as the multilayer degree corrected block model of \citet{agterberg2022joint}, multiple random eigen graphs of \citet{8889404}, and the common subspace independent edge graphs of \citet{arroyo2020inference}.
 In each of these models, the main inference problem is to identify parameters that are shared across multiple graphs. 
 Our contribution is in applying similar methods to identify parameters specific to each graph, which are in turn used as features for supervised learning. 
-
-The following sections are organized as follows: Section 2 introduces the structured and parameterized
-
-# Methods
+This further allows us to model graphs with differing vertex sets.
+
+The remainder of this paper is organized as follows: 
+Section 2 builds on the latent structure model to introduce the multiple latent structure model and proposes manifold learning algorithms for estimation. 
+Sections 3 and 4 provide simulation studies and real data analyses. 
+Section 5 concludes. 
+Proofs of theorems are provided in the Appendix.
+
+# Setup and Methodology
 
@@
 
-Let $G_1, ..., G_L$ be a collection of undirected graphs where each $G_l = (V_l, E_l)$. 
+Let $G^{(1)}, ..., G^{(L)}$ be a collection of undirected graphs where each $G^{(\ell)} = (V^{(\ell)}, E^{(\ell)})$. 
 In this setup, each graph can have its own vertex and edge sets. 
-Each graph is represented by adjacency matrices $A^{(1)}, ..., A^{(L)}$, where each $A^{(l)} \in \mathbb{R}^{n_l \times n_l}$ is a symmetric and hollow matrix such that $A^{(l)}_{ij}$ represents the edge weight between vertices $i$ and $j$, and $n_l$ represents the size of the vertex set of the $l$^th^ graph.
-Let $P^{(1)}, ..., P^{(L)}$ be the corresponding collection of edge parameter matrices such that each $A^{(l)}_{ij} \indep H(P^{(l)}_{ij})$ for each $1 \leq i < j \leq n_l$ and $H$ is a specified probability distribution such that $E[A^{(l)}_{ij}] = P^{(l)}_{ij}$.
-For ease of notation, we use $A^{(l)} \sim H(P^{(l)})$ for an adjacency matrix $A^{(l)}$ drawn from edge parameter matrix $P^{(l)}$ in this way.
+These graphs are represented by adjacency matrices $A^{(1)}, ..., A^{(L)}$, where each $A^{(\ell)} \in \mathbb{R}^{n_l \times n_l}$ is a symmetric and hollow matrix such that $A^{(\ell)}_{ij}$ represents the edge weight between vertices $i$ and $j$, and $n_l$ represents the size of the vertex set of the $l$^th^ graph.
+Let $P^{(1)}, ..., P^{(L)}$ be the corresponding collection of edge parameter matrices such that each $A^{(\ell)}_{ij} \indep H(P^{(\ell)}_{ij})$ for each $1 \leq i < j \leq n_l$ and $H$ is a specified probability distribution such that $E[A^{(\ell)}_{ij}] = P^{(\ell)}_{ij}$.
+For ease of notation, we use $A^{(\ell)} \sim H(P^{(\ell)})$ for an adjacency matrix $A^{(\ell)}$ drawn from edge parameter matrix $P^{(\ell)}$ in this way.
 Denote $X^{(1)}, ..., X^{(L)}$ as the sets of latent vectors corresponding to each graph under the GRDPG framework.
-Each $X^{(l)} \in \mathbb{R}^{n_l \times d}$, i.e., the graphs share the same latent dimension, and $x_i^{(l)} \in \mathbb{R}^d$ is the latent vector corresponding to the $i$^th^ vertex in the $l$^th^ graph.
+Each $X^{(\ell)} \in \mathbb{R}^{n_\ell \times d}$, i.e., the graphs share the same latent dimension, and $x_i^{(\ell)} \in \mathbb{R}^d$ is the latent vector corresponding to the $i$^th^ vertex in the $\ell$^th^ graph.
+$\hat{X}^{(\ell)}$ is the adjacency spectral embedding of $A^{(\ell)}$ and is an estimator for $X^{(\ell)}$.
 
@@
 
-We begin by defining the RDPG and the LSM.
+We begin by defining the random dot product graph and the latent structure model, then building the Multiple Latent Structure Model from there.
 
@@
 (See \citet{JMLR:v18:16-480}, \citet{https://doi.org/10.48550/arxiv.2106.09840}, and \citet{rubindelanchy2017statistical} for further discussion.)
+For ease of exposition, we set $\rho_n \equiv 1$ for the remainder of this paper. 
 \end{remark}
@@
 In the MLSM defined in definition \ref{def:mlsm}, the only commonality that we assume from graph to graph is that they are all LSMs with the same underlying distribution family. 
-If we impose further restrictions, namely that the latent structures are linear, we obtain the multilayer random dot product graph \citep{jones2021multilayer}.
+Unlike multilayer models, such as the multilayer random dot product graph \citep{jones2021multilayer} or COmmon Subspace Independent Edge (COSIE) graph \citep{arroyo2020inference}, there is no assumption that the graphs are connected in some way, such as via a shared vertex set.
+In the MLSM, each graph is drawn from the same process with different parameters, whereas in multilayer models the graphs share the same vertex set but the edges are drawn from different distributions.
+The MLSM can describe multilayer graphs if further restrictions are imposed on it.
 
@@
 However, if the community labels are not identical from graph to graph, the sequence of DCBMs cannot be described as a multilayer DCBM, multi-RDPG, or MREG, but it can still be described as an MLSM. 
-\end{example}
-
-\begin{example}[Nonlinear MLSM and comparison to MREG and multi-RDPG]
 \end{example}
@@
 We propose our methodology for classification and regression in three steps:
-
-1. Curve-fitting algorithm to recover the LSM curves from the ASEs.
-2. Quasi-maximum likelihood estimation to recover the parameters of the underlying distributions. 
-3. Classification or regression to determine the relationship between the parameters and the responses.
+First, we propose a curve-fitting algorithm to recover the LSM curves from the ASEs.
+Second, we the parameters of the underlying distributions from the curve fit via maximum likelihood estimation. 
+Third, we apply classification or regression methods to determine the relationship between the estimated parameters and the responses. 
+Finally, we provide theorems that show the consistency of the curve fits and maximum likelihood estimators.
 
@@
 For curve-fitting, we will restrict the LSM curves to non-self-intersecting Bezier polynomials \citep{10.5555/320367}. 
-A polynomial curve of degree $R$ in $\mathbb{R}^d$ is uniquely defined by a Bezier polynomial within the space of Bezier polynomials of degree at most $R$, up to reverse order, although there are infinitely many Bezier polynomials of degree greater than $R$ that can describe the curve \citep{SANCHEZREYES2022102118}.
+A polynomial curve of degree $R$ defined by $p : [0, 1] \to \mathbb{R}^d$ is uniquely defined by a Bezier polynomial within the space of Bezier polynomials of degree at most $R$, up to reverse order, although there are infinitely many Bezier polynomials of degree greater than $R$ that can describe the curve \citep{SANCHEZREYES2022102118}.
 We further restrict the curves to the case $p(0) = 0$, i.e., the curve begins at the origin. 
@@
 
-For initialization, the ordering of $t_1, ..., t_n$ is determined via a one-dimensional Isomap embedding \citep{Tenenbaum2000-ff}.
-
 Combining equation \ref{eq:bezier-ols} with the solutions to equation \ref{eq:dbezier} for each $i = 1, ..., n$ provides an alternating coordinate descent algorithm, as outlined in algorithm \ref{alg:bezier-fit}.
@@
 \setstretch{1.1}
-\KwData{Adjacency matrix $A$, embedding dimension $d$, polynomial degree $R$, Isomap neighborhood parameter $\lambda$, stopping criterion $\epsilon$.}
+\KwData{Adjacency matrix $A$, embedding dimension $d$, polynomial degree $R$, stopping criterion $\epsilon$.}
 \KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$.}
 Compute $\hat{X} \in \mathbb{R}^{n \times d}$, the ASE of $A$.\;
-Initialize timepoints $\hat{t}_1, ..., \hat{t}_n$, first via a one-dimensional Isomap embedding of $\hat{X}$ using neighborhood parameter $\lambda$, then by normalizing the embedding to $[0, 1]$.\;
+Initialize timepoints $\hat{t}_1, ..., \hat{t}_n$, (e.g., via a one-dimensional Isomap embedding \citep{Tenenbaum2000-ff}).\;
 \Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
@@
 \setstretch{1.1}
-\KwData{Adjacency matrix $A$, embedding dimension $d$, polynomial degree $R$, number of latent structures $K$, Isomap neighborhood parameter $\lambda$, stopping criterion $\epsilon$.}
+\KwData{Adjacency matrix $A$, embedding dimension $d$, polynomial degree $R$, number of latent structures $K$, stopping criterion $\epsilon$.}
 \KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$, curve memberships $\hat{z}_1, ..., \hat{z}_n$.}
@@
   Subset the rows of $\hat{X}$ by $z_i = k$ to obtain $\hat{x}_{k_1}, ..., \hat{x}_{k_{n_k}}$, where $n_k$ is the number of vertices with label $k$.\;
-  Initialize timepoints $\hat{t}_{k_1}, ..., \hat{t}_{k_{n_k}}$, first via a one-dimensional Isomap embedding of $\hat{x}_{k_1}, ..., \hat{x}_{k_{n_k}}$ using neighborhood parameter $\lambda$, then by normalizing the embedding to $[0, 1]$.\;
+  Initialize timepoints $\hat{t}_{k_1}, ..., \hat{t}_{k_{n_k}}$, (e.g., via a one-dimensional Isomap embedding \citep{Tenenbaum2000-ff}).\;
 }
@@
 
-\begin{lemma}
-\label{lem:bezier-consistency}
-Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R$ with coefficients $b$, and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
-Let $\hat{T}$ and $\hat{b}$ be the minimizers of the loss function in equation \ref{eq:bezier-ls} from $A$, and let $\tilde{X} = \hat{T} \hat{b}$, the estimated vectors along the fitted Bezier curve. 
-Then as $n \to \infty$, $\|X - \tilde{X} W\|_{2,\infty} \stackrel{p}{\to} 0$ for some $W \in \mathbb{O}(d)$.
-\end{lemma}
-
-\begin{proof}
-Let $X$ be the true latent positions of the LSM and $\hat{X}$ be the ASE of $A$ drawn from $X$. 
-Then $X = T b$, and $\hat{X} = \hat{T} \hat{b} + \hat{E}$, where $\hat{E}$ is the matrix of distances from the estimated positions along the fitted Bezier curves and the embedding vectors. 
-Then
-$$
-\begin{aligned}
-\|X - \tilde{X} W\| & = \|T b - \hat{T} \hat{b} W\|_{2,\infty} \\
-& = \|T b - (\hat{T} \hat{b} + \hat{E} - \hat{E}) W \|_{2,\infty} \\
-& = \|T b - (\hat{T} \hat{b} + \hat{E}) W + \hat{E} W\|_{2,\infty} \\
-& \leq \|X - \hat{X} W\|_{2,\infty} + \|\hat{E} W\|_{2,\infty}.
-\end{aligned}
-$$
-The first term $\|X - \hat{X} W\|_{2,\infty}$ converges to $0$ in probability \citep{lyzinski2014}. 
-Since the ASE is consistent, the embedding vectors will converge in probability to a Bezier curve, so $\|\hat{E} W\|_{2,\infty} \stackrel{p}{\to} 0$. 
-\end{proof}
-
-\begin{lemma}
-Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R$ with coefficients $b$, and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
-Let $\hat{T}$ and $\hat{b}$ be the minimizers of the loss function in equation \ref{eq:bezier-ls} from $A$. 
-Then $\|T - \hat{T}\|_{2, \infty} \stackrel{p}{\to} 0$ as if and only if $\|b - \hat{b} W\|_F \stackrel{p}{\to} 0$. 
-\end{lemma}
-
 \begin{theorem}[Consistency of algorithm \ref{alg:bezier-fit}]
@@
 
-\begin{proof}
-
-\end{proof}
-
-### Recovery of the Underlying Distributions
+### Quasi-Maximum Likelihood Estimation for $\theta$
+
+The estimation of $T$ and the restrictions imposed on the family of Bezier curves allows for the estimation of $\theta$ via maximum likelihood estimation (algorithrm \ref{alg:bezier-param-fit}). 
+Since $\hat{T}$ is a consistent estimator for $T$ by theorem \ref{thm:bezier-consistency}, if $t_1, ..., t_n \sim F_\theta$, a maximum likelihood estimator using $\hat{t}_1, ..., \hat{t}_n$ is a consistent estimator for $\theta$.
+Algorithm \ref{alg:bezier-param-fit} details the procedure.
 
@@
 \setstretch{1.1}
-\KwData{Adjacency matrix $A$, embedding dimension $d$, probability distribution family $F$, polynomial degree $R$, Isomap neighborhood parameter $\lambda$, stopping criterion $\epsilon$.}
+\KwData{Adjacency matrix $A$, embedding dimension $d$, probability distribution family $F$, polynomial degree $R$.}
 \KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$.}
 Compute $\hat{X} \in \mathbb{R}^{n \times d}$, the $d$-dimensional ASE of $A$.\;
-Initialize timepoints $\hat{t}_1, ..., \hat{t}_n$, first via a one-dimensional Isomap embedding of $\hat{X}$ using neighborhood parameter $\lambda$, then by normalizing the embedding to $[0, 1]$.\;
+Initialize timepoints $\hat{t}_1, ..., \hat{t}_n$.\;
 \Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
@@
 
-\begin{proof}
-By theorem \ref{thm:bezier-consistency}, 
-\end{proof}
-
 ### Classification and Regression
+
+Applying algorithm \ref{alg:bezier-param-fit} to supervised learning problems is straightforward. 
+First, we extract each $\hat{\theta}^{(\ell)}$ from each graph, then apply a supervised learning algorithm (e.g., ordinary least squares) for predicting each $y^{(\ell)}$ from each $\hat{\theta}^{(\ell)}$.
+Since \ref{alg:bezier-param-fit} produces consistent estimators, for large graphs the supervised learning algorithm in this regime will perform just as well as if we already knew each $\theta^{(\ell)}$.
+Therefore, $\tilde{\beta}$, the estimator based on $\hat{\theta}^{(1)}, ..., \hat{\theta}^{(L)}$, is consistent for $\beta$ (theorem \ref{thm:supervised}).
 
@@
 \setstretch{1.1}
-\KwData{Adjacency matrices $A^{(1)}, ..., A^{(L)}$, response variables $y_1, ..., y_L$, embedding dimension $d$, probability distribution family $F$, polynomial degree $R$, Isomap neighborhood parameters $\lambda_1, ..., \lambda_L$, stopping criterion $\epsilon$.}
+\KwData{Adjacency matrices $A^{(1)}, ..., A^{(L)}$, response variables $y_1, ..., y_L$, embedding dimension $d$, probability distribution family $F$, polynomial degree $R$, stopping criterion $\epsilon$.}
 \KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$.}
@@
   Compute $\hat{X}^{(\ell)} \in \mathbb{R}^{n_\ell \times d}$, the $d$-dimensional ASE of $A^{(\ell)} \in \mathbb{R}^{n_\ell \times n_\ell}$.\;
-  Initialize timepoints $\hat{t}_1, ..., \hat{t}_{n_\ell}$, first via a one-dimensional Isomap embedding of $\hat{X}^{(\ell)}$ using neighborhood parameter $\lambda_\ell$, then by normalizing the embedding to $[0, 1]$.\;
+  Initialize timepoints $\hat{t}_1, ..., \hat{t}_{n_\ell}$;
   \Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
@@
   }
-  Estimate $\hat{\theta}_\ell$ from $\hat{t}_1, ..., \hat{t}_{n_\ell}$ via maximum likelihood estimation.\;
+  Estimate $\hat{\theta}^{(\ell)}$ from $\hat{t}_1, ..., \hat{t}_{n_\ell}$ via maximum likelihood estimation.\;
 }
-Fit a model for predicting $y_1, ..., y_L$ using the estimated parameters $\theta_1, ..., \theta_L$ (e.g., via linear regression).\;
+Fit a model for predicting $y^{(\ell)}, ..., y^{(\ell)}$ using the estimated parameters $\theta^{(\ell)}, ..., \theta^{(\ell)}$ (e.g., via linear regression).\;
 \caption{Procedure for fitting a classification or regression model for an MLSM.}
@@
 \begin{theorem}
-\ref{thm:supervised}
+\label{thm:supervised} 
+Suppose $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta^{(\ell)}\})$ where $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R < d$ and $p(0) = 0$. 
+Let each $y^{(\ell)} \indep \mathcal{N}\big( (\theta^{(\ell)})^\top \beta, \sigma^2 \big)$, and let $\tilde{\beta}$ be the ordinary least squares estimate of $\beta$ using $\tilde{\theta}^{(1)}, ..., \tilde{\theta}^{(L)}$, the outputs of \ref{alg:bezier-param-fit} for each graph.
+Then as each $n_\ell \to \infty$, $\tilde{\beta} \stackrel{p}{\to} \beta$. 
 \end{theorem}
@@
 Two settings are examined. In the first, the data generating model reduces to a simpler model and can be adequately described by the COSIE model or multilayer DCBM. 
-The second setting involves a nonlinear latent structure that can be described as a polynomial curve and thus does not reduce to a simpler model or be analyzed by their corresponding methods.
-
-The scenarios are as follows:
-
-1. In this setting, there are two latent structures, described by two rays emanating from the origin.
-
-2.
-
-## Regression
-
-In this scenario, the responses $y_\ell$ are generated as linear responses of $\theta_\ell$ (in scenario 1) or $a_\ell$ and $b_\ell$ (in scenario 2). 
-More specifically, in scenario 1, each $y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 \theta_\ell, \sigma^2)$, and in scenario 2, each $y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 a_\ell + \beta_2 b_\ell, \sigma^2)$.
+The second setting involves graphs of variable sizes drawn from a nonlinear latent structure that can be described as a polynomial curve and thus does not reduce to a simpler model or be analyzed by their corresponding methods.
+
+In these simulations, the responses $y^{(\ell)}$ are generated as linear responses of $\theta^{(\ell)}$ (in scenario 1) or $a^{(\ell)}$ and $b^{(\ell)}$ (in scenario 2). 
+More specifically, in scenario 1, each $y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 \theta^{(\ell)}, \sigma^2)$, where $\theta^{(\ell)}$ is the angle between two line segments in the latent space, and in scenario 2, each $y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 a^{(\ell)} + \beta_2 b^{(\ell)}, \sigma^2)$, where $a^{(\ell)}$ and $b^{(\ell)}$ describe the Beta parameters of the distribution of latent vectors along a curve.
+We set $\sigma^2 = 0.25$ for both scenarios.
+
+The simulation setup for the first scenario is as follows: For each $N = 128, 256, 512, 1024$, sample $L = 128$ graphs as:
+
+  i. Draw $n_1, ..., n_L$ as $n_\ell \iid \Poisson(N)$.
+  ii. Draw $\theta^{(1)}, ..., \theta^{(L)} \iid \Uniform(\pi / 6, \pi / 3)$.
+  iii.
+For each $\ell = 1, ..., L$, draw latent vectors either from the line segment described by $p_1(t) = \begin{bmatrix} t & 0 \end{bmatrix}^\top$ or $p_2(t) = \begin{bmatrix} \cos(\theta^{(\ell)}) & \sin(\theta^{(\ell)}) \end{bmatrix}^\top$ with probability $1/2$ for each.
+  iv. Use the $L$ sets of latent vectors to draw $L$ RDPGs with adjacency matrices $A^{(1)}, ..., A^{(L)}$.
+  v. Draw each $y_\ell \indep \mathcal{N}(1 + 2 \theta^{(\ell)}, \sigma^2)$.
+  vi.
+Apply algorithm \ref{alg:mlsm-fit} to construct a linear model connecting $y_1, ..., y_\ell$ and $A^{(1)}, ..., A^{(L)}$, using the first half as training data and reporting MSE on the last half not used for training.
+
+For the second the simulation setup is similar. For each $N = 128, 256, 512, 1024$, sample $L = 128$ graphs as:
+
+  i. Draw $n_1, ..., n_L$ as $n_\ell \iid \Poisson(N)$.
+  ii. Draw each $(a^{(\ell)}, b^{(\ell)}) \iid \Uniform([0.5, 2] \times [0.5, 2])$.
+  iii.
+For each $\ell = 1, ..., L$, draw latent vectors by first taking $t_1, ..., t_{n_\ell} \sim \Betadist(a^{(\ell)}, b^{(\ell)})$ then taking each latent vector as $x_i = \begin{bmatrix} t^2 & 2 t (1-t) \end{bmatrix}^\top$.
+  iv. Use the $L$ sets of latent vectors to draw $L$ RDPGs with adjacency matrices $A^{(1)}, ..., A^{(L)}$.
+  v. Draw each $y_\ell \indep \mathcal{N}(1 + 2 \alpha^{(\ell)} + b^{(\ell)}, \sigma^2)$.
+  vi.
+Apply algorithm \ref{alg:mlsm-fit} to construct a linear model connecting $y_1, ..., y_\ell$ and $A^{(1)}, ..., A^{(L)}$, using the first half as training data and reporting MSE on the last half not used for training.
+
+In figure \ref{fig:angle-regression-sim}, we compare algorithm \ref{alg:mlsm-fit} to fitting a regression model using graph transitivity and average degree of each vertex. 
+Since this scenario is similar to a multilayer DCBM, we expect performance to be very similar, which is what we observe in the results. 
+However, we cannot model this simulation as a multilayer DCBM since although each graph can be described as a DCBM, the graphs do not share the same vertex set. 
+In figure \ref{fig:beta-regression-sim}, we see that linear regression using graph summary statistics performs better for smaller graphs, but as we draw more vertices, the ASE improves and so algorithm \ref{mlsm-fit} performs better, whereas graph summary statistics cannot improve with the number of vertices.
+Figures \ref{fig:theta-mse} and \ref{fig:beta-mse} show that the quasi-MLE estimators computed via algorithm \ref{alg:bezier-param-fit} are consistent.
 
@@
   scale_colour_brewer(palette = 'Set1') +
-  labs(x = 'vertices per graph', y = 'MSE', colour = 'predictor') +
+  labs(x = 'Expected vertices per graph', y = 'MSE', colour = 'Predictor') +
   geom_line(aes(x = n, y = mse.theta, colour = 'estimated angle')) + 
@@
                 width = .1) + 
-  labs(x = 'number of vertices', y = expression('MSE('~hat(theta)~')')) + 
+  labs(x = 'Expected vertices per graph', y = expression('MSE('~hat(theta)~')')) + 
   theme_bw() + 
@@
   scale_x_log10(breaks = sort(n.vec)) + 
-  labs(x = 'number of vertices', y = 'MSE', colour = 'predictor')
+  labs(x = 'Expected vertices per graph', y = 'MSE', colour = 'Predictor')
 ```
@@
 
-## Classification
-
-In this simulation experiment, the latent vectors were sampled along a Bezier curve defined by $g(t) = \bigl[ t^2 \; 2 t (1-t) \bigr]^\top$. 
-The timepoints $t_i$ were sampled as iid Beta random variables with two sets of parameters, $(\alpha_1, \beta_1)$ and $(\alpha_2, \beta_2)$. 
-The setup is as follows:
-
-1. Draw response variables $y_1, ..., y_n \iid \Multinomial(1/2, 1/2)$.
-2. For each $i = 1, ..., n$, draw $t_i \mid y_i \indep \Betadist(\alpha_{y_i}, \beta_{y_i})$, such that $\alpha_1 = 1$, $\beta_1 = 2$, $\alpha_2 = 2$, and $\beta_2 = 1$.
-3. Construct each latent vector as $x_i = g(t_i)$ and compile them in data matrix $X = \bigl[ x_1 \; \cdots \; x_n \bigr]^\top$.
-4. Sample graph and its adjacency matrix as $A \sim \RDPG(X)$.
-
-For each graph, we constructed the ASE, which was used to estimate the parameters $(\hat{\alpha}, \hat{\beta})$ for the graph, using the maximum likelihood method. 
-The estimated parameters were then used as predictors $y_1, ..., y_n$, setting aside half for training and half for testing. 
-We investigated graphs of size $|V| = 32, 64, 128, 256$. 
-The number of graphs for each experiment was set to $L = 64$.
-For each (number of graphs, size of graph) pair, we performed 32 replicates. 
-Figure \ref{fig:class-sim-ase} shows the ASE of one graph.  
-Figure \ref{fig:classification-sim} shows the boxplots of the classification error rates.
-
-```{r class-sim-ase, cache = TRUE, fig.cap = 'One simulated graph (left) and its ASE (right). The red curve is the fitted quadratic Bezier curve on the ASE.'}
-K <- 2
-n <- 256
-
-z <- 2
-
-t. <- rbeta(n, 1, 1)
-x1 <- t. ^ 2
-x2 <- 2 * t. * (1 - t.)
-X <- cbind(x1, x2)
-P <- X %*% t(X)
-A <- draw.graph(P)
-Xhat <- embedding(A, 2, 0)
-
-A.graph <- tidygraph::as_tbl_graph(unname(A), directed = FALSE)
-graph.plot <- ggraph(A.graph, layout = 'auto') + 
-  geom_edge_link(alpha = n ^ -.6) + 
-  geom_node_point(show.legend = FALSE) + 
-  labs(colour = NULL) + 
-  theme_void() + 
-  scale_colour_brewer(palette = 'Set1')
-
-curve.est <- estimate.bezier.curve.2(Xhat, 
-                                     degree = 2, 
-                                     intercept = FALSE, 
-                                     initialization = 'isomap', 
-                                     normalize = TRUE)
-ase.plot <- ggplot() + 
-  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2])) + 
-  geom_line(aes(x = curve.est$X[, 1], y = curve.est$X[, 2]), 
-             colour = 'red',
-             size = .2) + 
-  coord_fixed() + 
-  labs(x = expression(x[1]), y = expression(x[2])) + 
-  theme_bw()
-
-gridExtra::grid.arrange(graph.plot, ase.plot, nrow = 1)
-```
-
-```{r classification-sim, cache = TRUE, fig.cap = 'Median classification error rate and its IQR vs. number of vertices in each graph.', fig.height = 2, fig.width = 4}
-sim.df <- readr::read_csv('~/Documents/postdoc-first-project/classification-sim.csv')
-
-n.vec <- unique(sim.df$n)
-
-sim.df %>% 
-  dplyr::group_by(n) %>% 
-  dplyr::summarise(acc.med = median(acc),
-                   acc.high = quantile(acc, .75),
-                   acc.low = quantile(acc, .25)) %>% 
-  ggplot() + 
-  geom_line(aes(x = n, y = 1 - acc.med)) + 
-  geom_point(aes(x = n, y = 1 - acc.med)) + 
-  geom_errorbar(aes(x = n, ymin = 1 - acc.high, ymax = 1 - acc.low),
-                width = .1) + 
-  scale_x_log10(breaks = sort(n.vec)) +
-  # scale_y_log10() +
-  theme_bw() + 
-  labs(x = 'number of vertices', y = 'error rate')
-```
-
-## Regression
-
-In the first regression simulation,
-
-In the next regression simulation, we applied
-
-1. Draw angles $\theta_1, ..., \theta_N \iid \Uniform(\pi / 6, \pi / 3)$.
-2. For each $k = 1, ..., N$, 
-    i. Draw $t_1, ..., t_n \iid \Uniform(0, 1)$;
-    ii. Draw $z_1, ..., z_n \iid \Multinomial(1/2, 1/2)$;
-    iii. For each $i = 1, ..., n$, set $x_i = \begin{cases} \bigl[ t_i \; 0 \bigr]^\top \; z_i = 1 \\ \bigl[ t_i \cos \theta_\ell \; t_i \sin \theta_\ell \bigr]^\top \; z_i = 2 \end{cases}$;
-    iv. Collect $X = \bigl[ x_1 \; \cdots \; x_n \bigr]^\top$ and draw $A \sim \RDPG(X)$;
-    v. Set the response $y_\ell = \beta_0 + \beta_1 \theta_\ell$
-
-In this simulation, we set $\beta_0 = \beta_1 = 1$. 
-The number of graphs was set to $L = 64$, and the number of vertices per graph was set to $n = 128, 256, 512, 1024$. 
-For each $n$, we simulated 32 replicates.
-
-```{r angle-reg-example, cache = TRUE}
-n <- 512
-a.vec <- c(1, 1)
-b.vec <- c(1, 1)
-theta.min <- pi / 6
-theta.max <- pi / 3
-beta1 <- 1
-beta0 <- 1
-
-z <- sample(seq(2), n, replace = TRUE)
-t. <- rbeta(n, a.vec[z], b.vec[z])
-theta <- runif(1, theta.min, theta.max)
-y <- beta0 + beta1 * theta
-x1 <- ifelse(z == 1, t., cos(theta) * t.)
-x2 <- ifelse(z == 1, 0, sin(theta) * t.)
-X <- cbind(x1, x2)
-P <- X %*% t(X)
-A <- draw.graph(P)
-Xhat <- embedding(A, 2, 0)
-clustering <- manifold.clustering(Xhat, 
-                                  degree = 1,
-                                  initialization = z,
-                                  intercept = FALSE)
-ase.plot <- plot.estimated.curves(Xhat, clustering) +
-  scale_colour_brewer(palette = 'Set1')
-
-A.graph <- tidygraph::as_tbl_graph(unname(A), directed = FALSE)
-graph.plot <- ggraph(A.graph, layout = 'auto') + 
-  geom_edge_link(alpha = n ^ -.6) + 
-  geom_node_point(colour = 'black',
-                  fill = 'white',
-                  show.legend = FALSE) + 
-  labs(colour = NULL) + 
-  theme_void() + 
-  scale_colour_brewer(palette = 'Set1')
-
-gridExtra::grid.arrange(graph.plot, ase.plot, nrow = 1)
-```
-
-```{r angle-reg-results, cache = TRUE, fig.height = 2, fig.width = 4}
-sim.df <- readr::read_csv('~/dev/multilayer-rdpg/simulations/angle-regression-sim.csv')
-n.vec <- c(128, 256, 512, 1024)
-N <- 64
-
-sim.df %>% 
-  dplyr::group_by(n) %>% 
-  dplyr::summarise(mse.med = median(mse),
-                   mse.high = quantile(mse, .75),
-                   mse.low = quantile(mse, .25)) %>% 
-  ggplot() + 
-  geom_line(aes(x = n, y = mse.med)) + 
-  geom_point(aes(x = n, y = mse.med)) + 
-  geom_errorbar(aes(x = n, ymin = mse.high, ymax = mse.low),
-                width = .1) + 
-  scale_x_log10(breaks = sort(n.vec)) +
-  # scale_y_log10() +
-  theme_bw() + 
-  labs(x = 'number of vertices', y = 'MSE')
-```
-
 # Applications
@@
 The left and right are of the left and right hemispheres, respectively.
-Each vertex represents a neuron, which are labeled by neuron type.', fig.height = 2, figh.width = 12, cache = TRUE}
+Each vertex represents a neuron, which are labeled by neuron type.
+The red vertices are Kenyon Cells, the blue vertices are Input and Output Neurons, and the green vertices are Projection Neurons.', fig.height = 2, figh.width = 12, cache = TRUE}
 library(mbstructure)
@@
 z.r <- as.numeric(graph.r$vdf$type)
+z.r[z.r == 3] <- 2
+z.r[z.r == 4] <- 3
 n.r <- length(z.r)
@@
 z.l <- as.numeric(graph.l$vdf$type)
+z.l[z.l == 3] <- 2
+z.l[z.l == 4] <- 3
 n.l <- length(z.l)
@@
 
-```{r mbconnectome-ase, fig.cap = 'ASEs of the Drosophila connectome graphs. The embedding vectors are labeled by neuron type. ', fig.height = 2, fig.width = 8}
+```{r mbconnectome-ase, fig.cap = 'ASEs of the Drosophila connectome graphs.
+The embedding vectors are labeled by neuron type, with red being Kenyon Cells, blue being Input and Output Neurons, and green being Projection Neurons.', fig.height = 2, fig.width = 8}
 Xhat.l <- embedding(A.l, 2, 1)
@@
 If these Beta parameters are informative, we would expect the parameters for each hemisphere to match by neuron type, which is what we observe in these data (fig \ref{fig:mbconnectome-beta}).
-Using the three subgraphs of the left hemisphere as a training set to classify which subgraph of the right hemisphere belongs to which neuron type using the extracted Beta distribution parameter estimates as predictors results in 100\% accuracy.
 
 ```{r mbconnectome-beta, fig.cap = 'Beta parameter estimates for each curve and hemisphere.', fig.height = 3, fig.width = 6}
-param.est.df <- readr::read_csv('~/dev/multilayer-rdpg/drosophila.csv')
+param.est.df <- readr::read_csv('~/dev/multilayer-rdpg/drosophila.csv') %>% 
+  dplyr::mutate(region = factor(region, 
+                                levels = c('kenyon', 'in/out', 'projection')))
 
@@
 
-Next, we considered each hemisphere separately and fit a Beta distribution to the embedding of each half-network. 
-The scatterplot of the fitted $\alpha$ and $\beta$ parameters suggests that there are two clusters of brain networks (figure \ref{fig:hcp-clustering}). 
-Greater $\alpha$ values correspond to more embedding vectors farther away from the origin, corresponding to more nodes of higher degree, which in turn correspond to more brain regions with a greater number of connections.
-Conversely, greater $\beta$ values correspond to more embedding vectors closer to the origin, which in turn correspond to more brain regions with fewer connections. 
-This may suggest that there are underlying groups, one of which tends to have higher connectivity in the left hemisphere, and the other having higher connectivity in the right hemisphere.
-
-```{r hcp-clustering, cache = TRUE, fig.cap = 'Scatterplot of fitted parameters of the Beta distribution.
-The left image are the parameters fitted on the embedding of the left hemisphere, and the right image are the parameters fitted on the embedding of the right hemisphere.', fig.width = 8, fig.height = 4}
-beta.df <- ase.stats.df[c('a.l', 'a.r', 'b.l', 'b.r')]
-
-zhat.l <- mclust::Mclust(beta.df[, c('a.l', 'b.l')], G = 2)$classification
-plot.l <- ggplot(beta.df) + 
-  geom_point(aes(x = a.l, y = b.l, colour = factor(zhat.l))) + 
-  coord_fixed() + 
-  theme_bw() + 
-  labs(x = expression(hat(alpha)[l]), y = expression(hat(beta)[l]), colour = 'cluster') + 
-  scale_colour_brewer(palette = 'Set1')
-
-zhat.r <- mclust::Mclust(beta.df[, c('a.r', 'b.r')], G = 2)$classification
-plot.r <- ggplot(beta.df) + 
-  geom_point(aes(x = a.r, y = b.r, colour = factor(zhat.r))) + 
-  coord_fixed() +
-  theme_bw() + 
-  labs(x = expression(hat(alpha)[r]), y = expression(hat(beta)[r]), colour = 'cluster') + 
-  scale_colour_brewer(palette = 'Set1')
-
-gridExtra::grid.arrange(plot.l, plot.r, nrow = 1)
-```
-
 # Discussion
 
+Our main contribution is applying the latent structure model, a specific form of the random dot product graph, to multiple graphs for regression and classification problems. 
+We show that if the response variable is connected to the distribution parameter of a LSM, we can construct a consistent estimator for the parameter and then use that estimator to construct consistent supervised learning algorithms to estimate the responses.
+Since LSMs are not necessarily linear in the latent space, we show that our approach can accommodate nonlinear structures in the latent space, which cannot be adequately captured by other multiple graph models such as the multilayer DCBM or COSIE graph model.
+Furthermore, we remove the restriction of the graphs sharing the same vertex set, although the MLSM can still apply to such scenarios.
+
 \section*{Appendix A: Proofs of Theorems}
 
+\begin{lemma}
+\label{lem:bezier-consistency}
+Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R$ with coefficients $b$, and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
+Let $\hat{T}$ and $\hat{b}$ be the minimizers of the loss function in equation \ref{eq:bezier-ls} from $A$, and let $\tilde{X} = \hat{T} \hat{b}$, the estimated vectors along the fitted Bezier curve. 
+Then as $n \to \infty$, $\|X - \tilde{X} W\|_{2,\infty} \stackrel{p}{\to} 0$ for some $W \in \mathbb{O}(d)$.
+\end{lemma}
+
+\begin{proof}
+Let $X$ be the true latent positions of the LSM and $\hat{X}$ be the ASE of $A$ drawn from $X$. 
+Then $X = T b$, and $\hat{X} = \hat{T} \hat{b} + \hat{E}$, where $\hat{E}$ is the matrix of distances from the estimated positions along the fitted Bezier curves and the embedding vectors. 
+Then
+$$
+\begin{aligned}
+\|X - \tilde{X} W\| & = \|T b - \hat{T} \hat{b} W\|_{2,\infty} \\
+& = \|T b - (\hat{T} \hat{b} + \hat{E} - \hat{E}) W \|_{2,\infty} \\
+& = \|T b - (\hat{T} \hat{b} + \hat{E}) W + \hat{E} W\|_{2,\infty} \\
+& \leq \|X - \hat{X} W\|_{2,\infty} + \|\hat{E} W\|_{2,\infty}.
+\end{aligned}
+$$
+The first term $\|X - \hat{X} W\|_{2,\infty}$ converges to $0$ in probability \citep{lyzinski2014}. 
+Since the ASE is consistent, the embedding vectors will converge in probability to a Bezier curve, so $\|\hat{E} W\|_{2,\infty} \stackrel{p}{\to} 0$. 
+\end{proof}
+
 \newpage

### 2024-11-15 draft.tex
first draft of paper

@@
 graphs have been introduced, such as the multilayer RDPG, multiple RPDG,
-and multilayer stochastic block model. In this work, we use these joint
-random graph models in the context of statistical learning, such as
-classification and regression, by introducing the multiple latent
-structure model, in which the graphs share a common latent structure
-with different parameters that correspond to different response
-variables. Then we propose various estimation techniques involving
-manifold learning to estimate these parameters and in turn predict the
-responses, with theorems guaranteeing convergence of the predictions.
-Simulations, as well as applications on brain connectivity networks,
-verify the performance of our methods.
+and common subspace independent edge model. In this work, we use these
+joint random graph models in the context of graph classification and
+regression by introducing the multiple latent structure model, in which
+the graphs share a common latent structure with different parameters
+that correspond to different response variables. Then we propose various
+estimation techniques involving manifold learning to estimate these
+parameters and in turn predict the responses, with theorems guaranteeing
+convergence of the predictions. Simulations, as well as applications on
+brain connectivity networks, verify the performance of our methods.
 \end{abstract}
@@
 treat the network as a random graph that comes from some probability
-model. In particular, if we sum up the network in an adjacency matrix
-\(A \in \mathbb{R}^{n \times n}\), then one probability model might be
-to draw each element \(A_{ij}\), which represents the existence of an
-edge or the edge weight from vertex \(i\) to \(j\), independently from
-some distribution, perhaps with a unique parameter for the pair
-\((i,j)\), e.g., \(A_{ij} \indep F_{\theta_{ij}}\). The classical
-example of this is the \(\ER\) model \citep{Gilbert:1959}, in which
-every edge is drawn from the same distribution, typically a Bernoulli
-distribution, using the same parameter, i.e.,
-\(A_{ij} \iid \Bernoulli (p)\). The inhomogeneous Bernoulli graph
-extends this by allowing each edge, represented by \(A_{ij}\), to have
-its own parameter, \(P_{ij}\), e.g., in the Bernoulli case,
-\(A_{ij} \indep \Bernoulli(P_{ij})\). Typically, the parameters are
-collected into an edge probability matrix (in the case of unweighted
-graphs) or edge parameter matrix (in the case of weighted graphs),
-denoted as \(P \in \mathbb{R}^{n \times n}\). The network analysis
-problem in this setting is to estimate \(P\) given \(A\).
+model. One such approach in particular is the independent edge approach.
+If we sum up the network in an adjacency matrix
+\(A \in \mathbb{R}^{n \times n}\), then this model draws each element
+\(A_{ij}\), which represents the existence of an edge or the edge weight
+from vertex \(i\) to \(j\), independently from some distribution,
+perhaps with a unique parameter for the pair \((i,j)\), e.g.,
+\(A_{ij} \indep F_{\theta_{ij}}\). The classical example of this is the
+\(\ER\) model \citep{Gilbert:1959}, in which every edge is drawn from
+the same distribution, typically a Bernoulli distribution, using the
+same parameter, i.e., \(A_{ij} \iid \Bernoulli (p)\). The inhomogeneous
+Bernoulli graph extends this by allowing each edge, represented by
+\(A_{ij}\), to have its own parameter, \(P_{ij}\), e.g., in the
+Bernoulli case, \(A_{ij} \indep \Bernoulli(P_{ij})\). Typically, the
+parameters are collected into an edge probability matrix (in the case of
+unweighted graphs) or edge parameter matrix (in the case of weighted
+graphs), denoted as \(P \in \mathbb{R}^{n \times n}\). The network
+analysis problem in this setting is to estimate \(P\) given \(A\).
 
@@
 overparameterized. On the other hand, the classical \(\ER\) model is
-often too restrictive to describe real, observed networks. Much work has
-been done to develop models that are constrained enough for robust
-statistical inference while being generalizable to describe a wide range
-of networks. One such family of random graph models is the random dot
-product graph (RDPG), first proposed by
+often too restrictive to describe real, observed networks. Various
+approaches have been taken to restrict \(P\), such as by setting the
+matrix rank of \(P\) to a small value. One such family of random graph
+models is the random dot product graph (RDPG), first proposed by
 \citet{10.1007/978-3-540-77004-6_11}, which is a type of latent space
@@
 
-In this paper, we apply these highly structured GRDPG models to
-supervised learning problems in which there are multiple graphs,
-\(G_1, ..., G_L\), each with a response \(y^{(1)}, ..., y^{(L)}\). Each
-graph, represented by adjacency matrix \(A^{(l)}\), is presumed to be
-drawn from a GRDPG with parameter set \(\theta_l\), and in turn, the
-response \(y_l\) is drawn from a univariate distribution with parameter
-\(\theta_l\). Since the latent vectors of each \(A^{(l)}\) are presumed
-to be distributed on a structured subset of the latent space, e.g., a
-manifold, we call this the multiple latent structure model (MLSM). Much
-previous work has been done on multilayer networks, such as the
-multilayer degree corrected block model of \citet{agterberg2022joint},
-multiple random eigen graphs of \citet{8889404}, and the multiple common
-subspace independent edge graphs of \citet{arroyo2020inference}. In each
-of these models, the main inference problem is to identify parameters
-that are shared across multiple graphs. Our contribution is in applying
-similar methods to identify parameters specific to each graph, which are
-in turn used as features for supervised learning.
-
-The following sections are organized as follows: Section 2 introduces
-the structured and parameterized
-
-\section{Methods}\label{methods}
+In this paper, we apply these highly structured random dot product graph
+models to supervised learning problems in which there are multiple
+graphs, \(G_1, ..., G_L\), each with a response
+\(y^{(1)}, ..., y^{(L)}\). Each graph, represented by adjacency matrix
+\(A^{(\ell)}\), is presumed to be drawn from a RDPG with parameter set
+\(\theta_l\), and in turn, the response \(y_l\) is drawn from a
+univariate distribution with parameter \(\theta_l\). Since the latent
+vectors of each \(A^{(\ell)}\) are presumed to be distributed on a
+structured subset of the latent space, e.g., a manifold, we call this
+the multiple latent structure model (MLSM). Much previous work has been
+done on multilayer networks, such as the multilayer degree corrected
+block model of \citet{agterberg2022joint}, multiple random eigen graphs
+of \citet{8889404}, and the common subspace independent edge graphs of
+\citet{arroyo2020inference}. In each of these models, the main inference
+problem is to identify parameters that are shared across multiple
+graphs. Our contribution is in applying similar methods to identify
+parameters specific to each graph, which are in turn used as features
+for supervised learning. This further allows us to model graphs with
+differing vertex sets.
+
+The remainder of this paper is organized as follows: Section 2 builds on
+the latent structure model to introduce the multiple latent structure
+model and proposes manifold learning algorithms for estimation. Sections
+3 and 4 provide simulation studies and real data analyses. Section 5
+concludes. Proofs of theorems are provided in the Appendix.
+
+\section{Setup and Methodology}\label{setup-and-methodology}
 
@@
 
-Let \(G_1, ..., G_L\) be a collection of undirected graphs where each
-\(G_l = (V_l, E_l)\). In this setup, each graph can have its own vertex
-and edge sets. Each graph is represented by adjacency matrices
-\(A^{(1)}, ..., A^{(L)}\), where each
-\(A^{(l)} \in \mathbb{R}^{n_l \times n_l}\) is a symmetric and hollow
-matrix such that \(A^{(l)}_{ij}\) represents the edge weight between
+Let \(G^{(1)}, ..., G^{(L)}\) be a collection of undirected graphs where
+each \(G^{(\ell)} = (V^{(\ell)}, E^{(\ell)})\). In this setup, each
+graph can have its own vertex and edge sets. These graphs are
+represented by adjacency matrices \(A^{(1)}, ..., A^{(L)}\), where each
+\(A^{(\ell)} \in \mathbb{R}^{n_l \times n_l}\) is a symmetric and hollow
+matrix such that \(A^{(\ell)}_{ij}\) represents the edge weight between
 vertices \(i\) and \(j\), and \(n_l\) represents the size of the vertex
@@
 parameter matrices such that each
-\(A^{(l)}_{ij} \indep H(P^{(l)}_{ij})\) for each
+\(A^{(\ell)}_{ij} \indep H(P^{(\ell)}_{ij})\) for each
 \(1 \leq i < j \leq n_l\) and \(H\) is a specified probability
-distribution such that \(E[A^{(l)}_{ij}] = P^{(l)}_{ij}\). For ease of
-notation, we use \(A^{(l)} \sim H(P^{(l)})\) for an adjacency matrix
-\(A^{(l)}\) drawn from edge parameter matrix \(P^{(l)}\) in this way.
-Denote \(X^{(1)}, ..., X^{(L)}\) as the sets of latent vectors
-corresponding to each graph under the GRDPG framework. Each
-\(X^{(l)} \in \mathbb{R}^{n_l \times d}\), i.e., the graphs share the
-same latent dimension, and \(x_i^{(l)} \in \mathbb{R}^d\) is the latent
-vector corresponding to the \(i\)\textsuperscript{th} vertex in the
-\(l\)\textsuperscript{th} graph.
+distribution such that \(E[A^{(\ell)}_{ij}] = P^{(\ell)}_{ij}\). For
+ease of notation, we use \(A^{(\ell)} \sim H(P^{(\ell)})\) for an
+adjacency matrix \(A^{(\ell)}\) drawn from edge parameter matrix
+\(P^{(\ell)}\) in this way. Denote \(X^{(1)}, ..., X^{(L)}\) as the sets
+of latent vectors corresponding to each graph under the GRDPG framework.
+Each \(X^{(\ell)} \in \mathbb{R}^{n_\ell \times d}\), i.e., the graphs
+share the same latent dimension, and \(x_i^{(\ell)} \in \mathbb{R}^d\)
+is the latent vector corresponding to the \(i\)\textsuperscript{th}
+vertex in the \(\ell\)\textsuperscript{th} graph. \(\hat{X}^{(\ell)}\)
+is the adjacency spectral embedding of \(A^{(\ell)}\) and is an
+estimator for \(X^{(\ell)}\).
 
@@
 
-We begin by defining the RDPG and the LSM.
+We begin by defining the random dot product graph and the latent
+structure model, then building the Multiple Latent Structure Model from
+there.
 
@@
 that we assume from graph to graph is that they are all LSMs with the
-same underlying distribution family. If we impose further restrictions,
-namely that the latent structures are linear, we obtain the multilayer
-random dot product graph \citep{jones2021multilayer}.
+same underlying distribution family. Unlike multilayer models, such as
+the multilayer random dot product graph \citep{jones2021multilayer} or
+COmmon Subspace Independent Edge (COSIE) graph
+\citep{arroyo2020inference}, there is no assumption that the graphs are
+connected in some way, such as via a shared vertex set. In the MLSM,
+each graph is drawn from the same process with different parameters,
+whereas in multilayer models the graphs share the same vertex set but
+the edges are drawn from different distributions. The MLSM can describe
+multilayer graphs if further restrictions are imposed on it.
 
@@
 We propose our methodology for classification and regression in three
-steps:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Curve-fitting algorithm to recover the LSM curves from the ASEs.
-\item
-  Quasi-maximum likelihood estimation to recover the parameters of the
-  underlying distributions.
-\item
-  Classification or regression to determine the relationship between the
-  parameters and the responses.
-\end{enumerate}
+steps: First, we propose a curve-fitting algorithm to recover the LSM
+curves from the ASEs. Second, we the parameters of the underlying
+distributions from the curve fit via maximum likelihood estimation.
+Third, we apply classification or regression methods to determine the
+relationship between the estimated parameters and the responses.
+Finally, we provide theorems that show the consistency of the curve fits
+and maximum likelihood estimators.
 
@@
 non-self-intersecting Bezier polynomials \citep{10.5555/320367}. A
-polynomial curve of degree \(R\) in \(\mathbb{R}^d\) is uniquely defined
-by a Bezier polynomial within the space of Bezier polynomials of degree
-at most \(R\), up to reverse order, although there are infinitely many
-Bezier polynomials of degree greater than \(R\) that can describe the
-curve \citep{SANCHEZREYES2022102118}. We further restrict the curves to
-the case \(p(0) = 0\), i.e., the curve begins at the origin. This is
+polynomial curve of degree \(R\) defined by
+\(p : [0, 1] \to \mathbb{R}^d\) is uniquely defined by a Bezier
+polynomial within the space of Bezier polynomials of degree at most
+\(R\), up to reverse order, although there are infinitely many Bezier
+polynomials of degree greater than \(R\) that can describe the curve
+\citep{SANCHEZREYES2022102118}. We further restrict the curves to the
+case \(p(0) = 0\), i.e., the curve begins at the origin. This is
 consistent with well-known models such as the DCBM and PABM when viewed
@@
 
-For initialization, the ordering of \(t_1, ..., t_n\) is determined via
-a one-dimensional Isomap embedding \citep{Tenenbaum2000-ff}.
-
 Combining equation \ref{eq:bezier-ols} with the solutions to equation
@@
 \end{algorithm}
-
-\begin{lemma}
-\label{lem:bezier-consistency}
-Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R$ with coefficients $b$, and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
-Let $\hat{T}$ and $\hat{b}$ be the minimizers of the loss function in equation \ref{eq:bezier-ls} from $A$, and let $\tilde{X} = \hat{T} \hat{b}$, the estimated vectors along the fitted Bezier curve. 
-Then as $n \to \infty$, $\|X - \tilde{X} W\|_{2,\infty} \stackrel{p}{\to} 0$ for some $W \in \mathbb{O}(d)$.
-\end{lemma}
-
-\begin{proof}
-Let $X$ be the true latent positions of the LSM and $\hat{X}$ be the ASE of $A$ drawn from $X$. 
-Then $X = T b$, and $\hat{X} = \hat{T} \hat{b} + \hat{E}$, where $\hat{E}$ is the matrix of distances from the estimated positions along the fitted Bezier curves and the embedding vectors. 
-Then
-$$
-\begin{aligned}
-\|X - \tilde{X} W\| & = \|T b - \hat{T} \hat{b} W\|_{2,\infty} \\
-& = \|T b - (\hat{T} \hat{b} + \hat{E} - \hat{E}) W \|_{2,\infty} \\
-& = \|T b - (\hat{T} \hat{b} + \hat{E}) W + \hat{E} W\|_{2,\infty} \\
-& \leq \|X - \hat{X} W\|_{2,\infty} + \|\hat{E} W\|_{2,\infty}.
-\end{aligned}
-$$
-The first term $\|X - \hat{X} W\|_{2,\infty}$ converges to $0$ in probability \citep{lyzinski2014}. 
-Since the ASE is consistent, the embedding vectors will converge in probability to a Bezier curve, so $\|\hat{E} W\|_{2,\infty} \stackrel{p}{\to} 0$. 
-\end{proof}
-
-\begin{lemma}
-Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R$ with coefficients $b$, and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
-Let $\hat{T}$ and $\hat{b}$ be the minimizers of the loss function in equation \ref{eq:bezier-ls} from $A$. 
-Then $\|T - \hat{T}\|_{2, \infty} \stackrel{p}{\to} 0$ as if and only if $\|b - \hat{b} W\|_F \stackrel{p}{\to} 0$. 
-\end{lemma}
 
@@
 
-\begin{proof}
-
-\end{proof}
-
-\subsubsection{Recovery of the Underlying
-Distributions}\label{recovery-of-the-underlying-distributions}
+\subsubsection{\texorpdfstring{Quasi-Maximum Likelihood Estimation for
+\(\theta\)}{Quasi-Maximum Likelihood Estimation for \textbackslash theta}}\label{quasi-maximum-likelihood-estimation-for-theta}
+
+The estimation of \(T\) and the restrictions imposed on the family of
+Bezier curves allows for the estimation of \(\theta\) via maximum
+likelihood estimation (algorithrm \ref{alg:bezier-param-fit}). Since
+\(\hat{T}\) is a consistent estimator for \(T\) by theorem
+\ref{thm:bezier-consistency}, if \(t_1, ..., t_n \sim F_\theta\), a
+maximum likelihood estimator using \(\hat{t}_1, ..., \hat{t}_n\) is a
+consistent estimator for \(\theta\). Algorithm
+\ref{alg:bezier-param-fit} details the procedure.
 
@@
 
-\begin{proof}
-By theorem \ref{thm:bezier-consistency}, 
-\end{proof}
-
 \subsubsection{Classification and
 Regression}\label{classification-and-regression}
+
+Applying algorithm \ref{alg:bezier-param-fit} to supervised learning
+problems is straightforward. First, we extract each
+\(\hat{\theta}^{(\ell)}\) from each graph, then apply a supervised
+learning algorithm (e.g., ordinary least squares) for predicting each
+\(y^{(\ell)}\) from each \(\hat{\theta}^{(\ell)}\). Since
+\ref{alg:bezier-param-fit} produces consistent estimators, for large
+graphs the supervised learning algorithm in this regime will perform
+just as well as if we already knew each \(\theta^{(\ell)}\). Therefore,
+\(\tilde{\beta}\), the estimator based on
+\(\hat{\theta}^{(1)}, ..., \hat{\theta}^{(L)}\), is consistent for
+\(\beta\) (theorem \ref{thm:supervised}).
 
@@
   }
-  Estimate $\hat{\theta}_\ell$ from $\hat{t}_1, ..., \hat{t}_{n_\ell}$ via maximum likelihood estimation.\;
-}
-Fit a model for predicting $y_1, ..., y_L$ using the estimated parameters $\theta_1, ..., \theta_L$ (e.g., via linear regression).\;
+  Estimate $\hat{\theta}^{(\ell)}$ from $\hat{t}_1, ..., \hat{t}_{n_\ell}$ via maximum likelihood estimation.\;
+}
+Fit a model for predicting $y^{(\ell)}, ..., y^{(\ell)}$ using the estimated parameters $\theta^{(\ell)}, ..., \theta^{(\ell)}$ (e.g., via linear regression).\;
 \caption{Procedure for fitting a classification or regression model for an MLSM.}
@@
 model reduces to a simpler model and can be adequately described by the
-COSIE model or multilayer DCBM. The second setting involves a nonlinear
-latent structure that can be described as a polynomial curve and thus
-does not reduce to a simpler model or be analyzed by their corresponding
-methods.
-
-The scenarios are as follows:
+COSIE model or multilayer DCBM. The second setting involves graphs of
+variable sizes drawn from a nonlinear latent structure that can be
+described as a polynomial curve and thus does not reduce to a simpler
+model or be analyzed by their corresponding methods.
+
+In these simulations, the responses \(y^{(\ell)}\) are generated as
+linear responses of \(\theta^{(\ell)}\) (in scenario 1) or
+\(a^{(\ell)}\) and \(b^{(\ell)}\) (in scenario 2). More specifically, in
+scenario 1, each
+\(y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 \theta^{(\ell)}, \sigma^2)\),
+where \(\theta^{(\ell)}\) is the angle between two line segments in the
+latent space, and in scenario 2, each
+\(y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 a^{(\ell)} + \beta_2 b^{(\ell)}, \sigma^2)\),
+where \(a^{(\ell)}\) and \(b^{(\ell)}\) describe the Beta parameters of
+the distribution of latent vectors along a curve. We set
+\(\sigma^2 = 0.25\) for both scenarios.
+
+The simulation setup for the first scenario is as follows: For each
+\(N = 128, 256, 512, 1024\), sample \(L = 128\) graphs as:
 
 \begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  In this setting, there are two latent structures, described by two
-  rays emanating from the origin.
-\item
+\def\labelenumi{\roman{enumi}.}
+\tightlist
+\item
+  Draw \(n_1, ..., n_L\) as \(n_\ell \iid \Poisson(N)\).
+\item
+  Draw
+  \(\theta^{(1)}, ..., \theta^{(L)} \iid \Uniform(\pi / 6, \pi / 3)\).
+\item
+  For each \(\ell = 1, ..., L\), draw latent vectors either from the
+  line segment described by
+  \(p_1(t) = \begin{bmatrix} t & 0 \end{bmatrix}^\top\) or
+  \(p_2(t) = \begin{bmatrix} \cos(\theta^{(\ell)}) & \sin(\theta^{(\ell)}) \end{bmatrix}^\top\)
+  with probability \(1/2\) for each.
+\item
+  Use the \(L\) sets of latent vectors to draw \(L\) RDPGs with
+  adjacency matrices \(A^{(1)}, ..., A^{(L)}\).
+\item
+  Draw each
+  \(y_\ell \indep \mathcal{N}(1 + 2 \theta^{(\ell)}, \sigma^2)\).
+\item
+  Apply algorithm \ref{alg:mlsm-fit} to construct a linear model
+  connecting \(y_1, ..., y_\ell\) and \(A^{(1)}, ..., A^{(L)}\), using
+  the first half as training data and reporting MSE on the last half not
+  used for training.
 \end{enumerate}
 
-\subsection{Regression}\label{regression}
-
-In this scenario, the responses \(y_\ell\) are generated as linear
-responses of \(\theta_\ell\) (in scenario 1) or \(a_\ell\) and
-\(b_\ell\) (in scenario 2). More specifically, in scenario 1, each
-\(y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 \theta_\ell, \sigma^2)\),
-and in scenario 2, each
-\(y_\ell \indep \mathcal{N}(\beta_0 + \beta_1 a_\ell + \beta_2 b_\ell, \sigma^2)\).
+For the second the simulation setup is similar. For each
+\(N = 128, 256, 512, 1024\), sample \(L = 128\) graphs as:
+
+\begin{enumerate}
+\def\labelenumi{\roman{enumi}.}
+\tightlist
+\item
+  Draw \(n_1, ..., n_L\) as \(n_\ell \iid \Poisson(N)\).
+\item
+  Draw each
+  \((a^{(\ell)}, b^{(\ell)}) \iid \Uniform([0.5, 2] \times [0.5, 2])\).
+\item
+  For each \(\ell = 1, ..., L\), draw latent vectors by first taking
+  \(t_1, ..., t_{n_\ell} \sim \Betadist(a^{(\ell)}, b^{(\ell)})\) then
+  taking each latent vector as
+  \(x_i = \begin{bmatrix} t^2 & 2 t (1-t) \end{bmatrix}^\top\).
+\item
+  Use the \(L\) sets of latent vectors to draw \(L\) RDPGs with
+  adjacency matrices \(A^{(1)}, ..., A^{(L)}\).
+\item
+  Draw each
+  \(y_\ell \indep \mathcal{N}(1 + 2 \alpha^{(\ell)} + b^{(\ell)}, \sigma^2)\).
+\item
+  Apply algorithm \ref{alg:mlsm-fit} to construct a linear model
+  connecting \(y_1, ..., y_\ell\) and \(A^{(1)}, ..., A^{(L)}\), using
+  the first half as training data and reporting MSE on the last half not
+  used for training.
+\end{enumerate}
+
+In figure \ref{fig:angle-regression-sim}, we compare algorithm
+\ref{alg:mlsm-fit} to fitting a regression model using graph
+transitivity and average degree of each vertex. Since this scenario is
+similar to a multilayer DCBM, we expect performance to be very similar,
+which is what we observe in the results. However, we cannot model this
+simulation as a multilayer DCBM since although each graph can be
+described as a DCBM, the graphs do not share the same vertex set. In
+figure \ref{fig:beta-regression-sim}, we see that linear regression
+using graph summary statistics performs better for smaller graphs, but
+as we draw more vertices, the ASE improves and so algorithm
+\ref{mlsm-fit} performs better, whereas graph summary statistics cannot
+improve with the number of vertices. Figures \ref{fig:theta-mse} and
+\ref{fig:beta-mse} show that the quasi-MLE estimators computed via
+algorithm \ref{alg:bezier-param-fit} are consistent.
 
@@
 \end{figure}
-
-\subsection{Classification}\label{classification}
-
-In this simulation experiment, the latent vectors were sampled along a
-Bezier curve defined by \(g(t) = \bigl[ t^2 \; 2 t (1-t) \bigr]^\top\).
-The timepoints \(t_i\) were sampled as iid Beta random variables with
-two sets of parameters, \((\alpha_1, \beta_1)\) and
-\((\alpha_2, \beta_2)\). The setup is as follows:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Draw response variables \(y_1, ..., y_n \iid \Multinomial(1/2, 1/2)\).
-\item
-  For each \(i = 1, ..., n\), draw
-  \(t_i \mid y_i \indep \Betadist(\alpha_{y_i}, \beta_{y_i})\), such
-  that \(\alpha_1 = 1\), \(\beta_1 = 2\), \(\alpha_2 = 2\), and
-  \(\beta_2 = 1\).
-\item
-  Construct each latent vector as \(x_i = g(t_i)\) and compile them in
-  data matrix \(X = \bigl[ x_1 \; \cdots \; x_n \bigr]^\top\).
-\item
-  Sample graph and its adjacency matrix as \(A \sim \RDPG(X)\).
-\end{enumerate}
-
-For each graph, we constructed the ASE, which was used to estimate the
-parameters \((\hat{\alpha}, \hat{\beta})\) for the graph, using the
-maximum likelihood method. The estimated parameters were then used as
-predictors \(y_1, ..., y_n\), setting aside half for training and half
-for testing. We investigated graphs of size \(|V| = 32, 64, 128, 256\).
-The number of graphs for each experiment was set to \(L = 64\). For each
-(number of graphs, size of graph) pair, we performed 32 replicates.
-Figure \ref{fig:class-sim-ase} shows the ASE of one graph.\\
-Figure \ref{fig:classification-sim} shows the boxplots of the
-classification error rates.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{draft_files/figure-latex/class-sim-ase-1}
-
-}
-
-\caption{One simulated graph (left) and its ASE (right). The red curve is the fitted quadratic Bezier curve on the ASE.}\label{fig:class-sim-ase}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{draft_files/figure-latex/classification-sim-1}
-
-}
-
-\caption{Median classification error rate and its IQR vs. number of vertices in each graph.}\label{fig:classification-sim}
-\end{figure}
-
-\subsection{Regression}\label{regression-1}
-
-In the first regression simulation,
-
-In the next regression simulation, we applied
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Draw angles
-  \(\theta_1, ..., \theta_N \iid \Uniform(\pi / 6, \pi / 3)\).
-\item
-  For each \(k = 1, ..., N\),
-
-  \begin{enumerate}
-  \def\labelenumii{\roman{enumii}.}
-  \tightlist
-  \item
-    Draw \(t_1, ..., t_n \iid \Uniform(0, 1)\);
-  \item
-    Draw \(z_1, ..., z_n \iid \Multinomial(1/2, 1/2)\);
-  \item
-    For each \(i = 1, ..., n\), set
-    \(x_i = \begin{cases} \bigl[ t_i \; 0 \bigr]^\top \; z_i = 1 \\ \bigl[ t_i \cos \theta_\ell \; t_i \sin \theta_\ell \bigr]^\top \; z_i = 2 \end{cases}\);
-  \item
-    Collect \(X = \bigl[ x_1 \; \cdots \; x_n \bigr]^\top\) and draw
-    \(A \sim \RDPG(X)\);
-  \item
-    Set the response \(y_\ell = \beta_0 + \beta_1 \theta_\ell\)
-  \end{enumerate}
-\end{enumerate}
-
-In this simulation, we set \(\beta_0 = \beta_1 = 1\). The number of
-graphs was set to \(L = 64\), and the number of vertices per graph was
-set to \(n = 128, 256, 512, 1024\). For each \(n\), we simulated 32
-replicates.
-
-\begin{center}\includegraphics{draft_files/figure-latex/angle-reg-example-1} \end{center}
-
-\begin{center}\includegraphics{draft_files/figure-latex/angle-reg-results-1} \end{center}
 
@@
 The left and right are of the left and right hemispheres, respectively.
-Each vertex represents a neuron, which are labeled by neuron type.}\label{fig:mbconnectome-graph}
+Each vertex represents a neuron, which are labeled by neuron type.
+The red vertices are Kenyon Cells, the blue vertices are Input and Output Neurons, and the green vertices are Projection Neurons.}\label{fig:mbconnectome-graph}
 \end{figure}
@@
 
-\caption{ASEs of the Drosophila connectome graphs. The embedding vectors are labeled by neuron type. }\label{fig:mbconnectome-ase}
+\caption{ASEs of the Drosophila connectome graphs.
+The embedding vectors are labeled by neuron type, with red being Kenyon Cells, blue being Input and Output Neurons, and green being Projection Neurons.}\label{fig:mbconnectome-ase}
 \end{figure}
@@
 is what we observe in these data (fig \ref{fig:mbconnectome-beta}).
-Using the three subgraphs of the left hemisphere as a training set to
-classify which subgraph of the right hemisphere belongs to which neuron
-type using the extracted Beta distribution parameter estimates as
-predictors results in 100\% accuracy.
 
@@
 
-Next, we considered each hemisphere separately and fit a Beta
-distribution to the embedding of each half-network. The scatterplot of
-the fitted \(\alpha\) and \(\beta\) parameters suggests that there are
-two clusters of brain networks (figure \ref{fig:hcp-clustering}).
-Greater \(\alpha\) values correspond to more embedding vectors farther
-away from the origin, corresponding to more nodes of higher degree,
-which in turn correspond to more brain regions with a greater number of
-connections. Conversely, greater \(\beta\) values correspond to more
-embedding vectors closer to the origin, which in turn correspond to more
-brain regions with fewer connections. This may suggest that there are
-underlying groups, one of which tends to have higher connectivity in the
-left hemisphere, and the other having higher connectivity in the right
-hemisphere.
-
-\begin{figure}[H]
-
-{\centering \includegraphics{draft_files/figure-latex/hcp-clustering-1}
-
-}
-
-\caption{Scatterplot of fitted parameters of the Beta distribution.
-The left image are the parameters fitted on the embedding of the left hemisphere, and the right image are the parameters fitted on the embedding of the right hemisphere.}\label{fig:hcp-clustering}
-\end{figure}
-
 \section{Discussion}\label{discussion}
 
+Our main contribution is applying the latent structure model, a specific
+form of the random dot product graph, to multiple graphs for regression
+and classification problems. We show that if the response variable is
+connected to the distribution parameter of a LSM, we can construct a
+consistent estimator for the parameter and then use that estimator to
+construct consistent supervised learning algorithms to estimate the
+responses. Since LSMs are not necessarily linear in the latent space, we
+show that our approach can accommodate nonlinear structures in the
+latent space, which cannot be adequately captured by other multiple
+graph models such as the multilayer DCBM or COSIE graph model.
+Furthermore, we remove the restriction of the graphs sharing the same
+vertex set, although the MLSM can still apply to such scenarios.
+
 \section*{Appendix A: Proofs of Theorems}
+
+\begin{lemma}
+\label{lem:bezier-consistency}
+Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R$ with coefficients $b$, and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
+Let $\hat{T}$ and $\hat{b}$ be the minimizers of the loss function in equation \ref{eq:bezier-ls} from $A$, and let $\tilde{X} = \hat{T} \hat{b}$, the estimated vectors along the fitted Bezier curve. 
+Then as $n \to \infty$, $\|X - \tilde{X} W\|_{2,\infty} \stackrel{p}{\to} 0$ for some $W \in \mathbb{O}(d)$.
+\end{lemma}
+
+\begin{proof}
+Let $X$ be the true latent positions of the LSM and $\hat{X}$ be the ASE of $A$ drawn from $X$. 
+Then $X = T b$, and $\hat{X} = \hat{T} \hat{b} + \hat{E}$, where $\hat{E}$ is the matrix of distances from the estimated positions along the fitted Bezier curves and the embedding vectors. 
+Then
+$$
+\begin{aligned}
+\|X - \tilde{X} W\| & = \|T b - \hat{T} \hat{b} W\|_{2,\infty} \\
+& = \|T b - (\hat{T} \hat{b} + \hat{E} - \hat{E}) W \|_{2,\infty} \\
+& = \|T b - (\hat{T} \hat{b} + \hat{E}) W + \hat{E} W\|_{2,\infty} \\
+& \leq \|X - \hat{X} W\|_{2,\infty} + \|\hat{E} W\|_{2,\infty}.
+\end{aligned}
+$$
+The first term $\|X - \hat{X} W\|_{2,\infty}$ converges to $0$ in probability \citep{lyzinski2014}. 
+Since the ASE is consistent, the embedding vectors will converge in probability to a Bezier curve, so $\|\hat{E} W\|_{2,\infty} \stackrel{p}{\to} 0$. 
+\end{proof}
 

### 2024-11-15 hcp-a-eda.R

@@ -23,10 +23,17 @@ z <- dti.info$hemisphere %>%
 
 rois <- dti.info$aparc_aseg_region.names.short
 
-A <- dat.NumFibers[[2]] %>% 
+subject <- 2
+
+A <- dat.NumFibers[[subject]] %>% 
   as.matrix()
 A <- (A + t(A)) / 2
 diag(A) <- 0
+
+Sigma <- fmri.mc[[subject]] %>% 
+  as.data.frame() %>% 
+  subset(., select = which(!duplicated(names(.)))) %>%
+  cor()
 
 # A <- log(A)
 # A[is.infinite(A)] <- 0

### 2024-11-15 simulations/beta-classification/beta-classification-sim.R

@@ -83,7 +83,7 @@ out.df <- foreach::foreach(n = n.vec, .combine = dplyr::bind_rows) %do% {
       apply(1, which.max)
     acc.mlsm <- mean(z.test == z.hat.mlsm)
     acc.stats <- mean(z.test == z.hat.stats)
-    dplyr::tibble(n = n, acc.mslm = acc.mlsm, acc.stats = acc.stats)
+    dplyr::tibble(n = n, acc.mlsm = acc.mlsm, acc.stats = acc.stats)
   }
 }
 
@@ -94,18 +94,18 @@ ggplot(out.df) +
 
 out.df %>% 
   dplyr::group_by(n) %>% 
-  dplyr::summarise(acc.mslm.mean = mean(acc.mslm),
+  dplyr::summarise(acc.mlsm.mean = mean(acc.mlsm),
                    acc.stats.mean = mean(acc.stats),
-                   acc.mslm.se = sd(acc.mslm) / dplyr::n(),
+                   acc.mlsm.se = sd(acc.mlsm) / dplyr::n(),
                    acc.stats.se = sd(acc.stats) / dplyr::n()) %>% 
   ggplot() + 
   scale_x_log10(breaks = sort(n.vec)) + 
-  geom_line(aes(x = n, y = acc.mslm.mean, colour = 'mslm')) + 
+  geom_line(aes(x = n, y = acc.mlsm.mean, colour = 'mlsm')) + 
   geom_line(aes(x = n, y = acc.stats.mean, colour = 'stats')) + 
   geom_errorbar(aes(x = n, 
-                    ymin = acc.mslm.mean - 2 * acc.mslm.se, 
-                    ymax = acc.mslm.mean + 2 * acc.mslm.se,
-                    colour = 'mslm')) + 
+                    ymin = acc.mlsm.mean - 2 * acc.mlsm.se, 
+                    ymax = acc.mlsm.mean + 2 * acc.mlsm.se,
+                    colour = 'mlsm')) + 
   geom_errorbar(aes(x = n, 
                     ymin = acc.stats.mean - 2 * acc.stats.se, 
                     ymax = acc.stats.mean + 2 * acc.stats.se,

### 2025-01-10 draft.Rmd
typed up proofs

@@
 \begin{theorem}[Consistency of algorithm \ref{alg:bezier-param-fit}]
+\label{thm:theta-consistency}
 Suppose $A \sim \LSM(p, F_\theta)$ such that $p : [0, 1] \to \mathbb{R}^d$ is a Bezier polynomial of degree $R$ with coefficients $b$, and $F_\theta$ is a probability distribution with probability density function $f_\theta(t)$ that is absolutely continuous on support $[0, 1]$ and $\min_t f_\theta(t) > 0$ $\forall t \in [0, 1]$.
@@
 \begin{aligned}
-\|X - \tilde{X} W\| & = \|T b - \hat{T} \hat{b} W\|_{2,\infty} \\
+\|X - \tilde{X} W\|_{2, \infty} & = \|T b - \hat{T} \hat{b} W\|_{2,\infty} \\
 & = \|T b - (\hat{T} \hat{b} + \hat{E} - \hat{E}) W \|_{2,\infty} \\
@@
 
+\begin{proof}[Proof of Theorem \ref{thm:bezier-consistency}]
+Recall that for the family of Bezier curves such that $p(t) = 0$ and $p$ cannot be expressed by another Bezier polynomial of lower order, the image of $p$ cannot be expressed by another Bezier curve \citep{SANCHEZREYES2022102118}.
+Thus, $X = T b$ is unique. 
+Since the ASE is consistent, $\hat{X}$ converges to a Bezier curve, which again is uniquely expressed by $\hat{X} = \hat{T} \hat{b} = \tilde{X}$. 
+By lemma \ref{lem:bezier-consistency}, $\|X - \tilde{X} W\|_{2, \infty} = \|T b - \hat{T} \hat{b} W\|_{2, \infty} \stackrel{p}{\to} 0$.
+Then
+$$
+\begin{aligned}
+\|T b - \hat{T} \hat{b} W\|_{2, \infty} & = \|T b - \hat{T} \hat{b} W + T \hat{b} W - T \hat{b} W\|_{2,\infty} \\
+& \leq \|T (b - \hat{b} W) \|_{2,\infty} + \|(\hat{T} - T) \hat{b} W\|_{2, \infty}\\
+& \leq \|T\|_{2, \infty} \|b - \hat{b} W\|_F + \|\hat{T} - T\|_{2, \infty} \|\hat{b}\|_F.
+\end{aligned}
+$$
+Since this converges to $0$ in probability and all parts are nonnegative, both $\|b - \hat{b} W\|_F$ and $\|\hat{T} - T\|_{2, \infty}$ must converge to 0 as well. 
+\end{proof}
+
+\begin{proof}[Proof of Theorem \ref{thm:theta-consistency}]
+By theorem \ref{thm:bezier-consistency}, $\|T - \hat{T}\|_{2, \infty} \stackrel{p}{\to} 0$. 
+Each element of $T$ is $T_{ir} = \binom{R}{r} t_i^r (1 - t_i)^{R-r}$, and similarly $\hat{T}_{ir} = \binom{R}{r} \hat{t}_i^r (1 - \hat{t}_i)^{R-r}$. 
+This is unique for the class of Bezier polynomials restricted by algorithm 3. 
+Then $\sum_i^n \|t_i - \hat{t}_i\| \stackrel{p}{\to} 0$. 
+\end{proof}
+
 \newpage

### 2025-01-24 draft.Rmd
minor changes

@@
 
-It has been shown \citep{Koo_2022, rubindelanchy2017statistical} that the RDPG (as well as the generalized random dot product graph, or GRDPG, \citep{rubindelanchy2017statistical}) can describe a wide range of popular random graph models, such as the $\ER$ model, the stochastic block model (SBM) \citep{doi:10.1080/0022250X.1971.9989788}, degree corrected block model (DCBM) \citep{Karrer_2011}, and popularity adjusted block model (PABM) \citep{307cbeb9b1be48299388437423d94bf1}.
-In these examples, the (G)RPDG representation of these networks allows for inference and analysis in more familiar Euclidean space rather than on graphs, and the geometry of the configuration of latent vectors correspond to the parameters of these graph models.
-More specifically, when these network models are described as (G)RDPGs, each latent vector lies on a subspace. 
-\citet{athreya2020estimation} and \citet{Passino_2020} extended this to the latent structure model (LSM) and latent structure block model (LSBM) to include latent configurations that consist of nonlinear manifolds.
-
+It has been shown that the RDPG (as well as the generalized random dot product graph, or GRDPG, \citep{rubindelanchy2017statistical}) can describe a wide range of popular random graph models, such as the $\ER$ model, the stochastic block model (SBM) \citep{doi:10.1080/0022250X.1971.9989788}, degree corrected block model (DCBM) \citep{Karrer_2011}, and popularity adjusted block model (PABM) \citep{307cbeb9b1be48299388437423d94bf1}.
 In this paper, we apply these highly structured random dot product graph models to supervised learning problems in which there are multiple graphs, $G_1, ..., G_L$, each with a response $y^{(1)}, ..., y^{(L)}$.
@@
 Since the latent vectors of each $A^{(\ell)}$ are presumed to be distributed on a structured subset of the latent space, e.g., a manifold, we call this the multiple latent structure model (MLSM). 
-Much previous work has been done on multilayer networks, such as the multilayer degree corrected block model of \citet{agterberg2022joint}, multiple random eigen graphs of \citet{8889404}, and the common subspace independent edge graphs of \citet{arroyo2020inference}.
+There exists much existing literature on the estimation for multilayer networks, such as the multilayer degree corrected block model of \citet{agterberg2022joint}, multiple random eigen graphs of \citet{8889404}, and the common subspace independent edge graphs of \citet{arroyo2020inference}.
 In each of these models, the main inference problem is to identify parameters that are shared across multiple graphs. 
@@
 
-### Connections to Related Models
-
 In the MLSM defined in definition \ref{def:mlsm}, the only commonality that we assume from graph to graph is that they are all LSMs with the same underlying distribution family. 
@@
 
-### Classification and Regression on the MLSM
-
-To use the MLSM for classification or regression problems, we assign response variables $y_1, ..., y_L$ to each graph. 
+To use the MLSM for regression problems, we assign response variables $y_1, ..., y_L$ to each graph. 
 Then if each response $y_\ell$ depends on $p^{(\ell)}(t)$ (the parameterization of the $\ell^{th}$ curve) or $\theta_\ell$ (the parameter of the $\ell^{th}$ distribution), or some combination of the two, there is a plausible setup for a predictive modeling task for predicting $y_\ell$ after observing $A^{(\ell)}$.
-Four such scenarios are given:
-
-\begin{definition}[MLSM for classification 1]
-\label{def:mlsm-c-1}
-Let $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$, $y_1, ..., y_L \in \{1, ..., M\}$ be labels for each of the $L$ graphs of the MLSM, and each $\theta_\ell$ take on one of $M$ values $\{\phi_1, ..., \phi_M\}$ corresponding to its response, $y_\ell$, i.e., $\theta_\ell = \phi_{y_\ell}$.
-
-We observe the adjacency matrices $A^{(1)}, ..., A^{(L)}$ and the first $r$ labels $y_1, ..., y_r$, and the task is to predict the unknown labels $y_{r+1}, ..., y_L$.
-
-This can be described by the following generative model:
-
-\begin{enumerate}
-  \item Draw labels $y_1, ..., y_L \iid \Categorical(\{1, ..., M\}, \{\pi_1, ..., \pi_M\})$.
-  \item Set each $\theta_\ell = \phi_{y_\ell}$.
-  \item Draw adjacency matrices as $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$.
-\end{enumerate}
-\end{definition}
-
-\begin{definition}[MLSM for classification 2]
-\label{def:mlsm-c-2}
-Let $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{\mathcal{C}^{(\ell)}\}, F, \{\theta_\ell\}_L)$, $y_1, ..., y_L \in \{1, ..., M\}$ be labels for each of the $L$ graphs of the MLSM, and each $\mathcal{C}^{(\ell)}$ take on one of $M$ functional forms $\{p^{(1)}(t), ..., p^{(M)}(t)\}$ corresponding to its response variable, $y_\ell$, i.e., $\mathcal{C}^{(\ell)}$ is parameterized by $p^{(y_\ell)}(t)$.
-
-We observe the adjacency matrices $A^{(1)}, ..., A^{(L)}$ and the first $r$ labels $y_1, ..., y_r$, and the task is to predict the unknown labels $y_{r+1}, ..., y_L$.
-
-This can be described by the following generative model:
-
-\begin{enumerate}
-  \item Draw labels $y_1, ..., y_L \iid \Categorical(\{1, ..., M\}, \{\pi_1, ..., \pi_M\})$.
-  \item Set each $p^{(\ell)}(t) = p^{(y_\ell)}(t)$.
-  \item Draw adjacency matrices as $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$.
-\end{enumerate}
-
-As in remark \ref{remark:mlsm-mixture}, we also allow for each adjacency matrix $A^{(\ell)}$ to be drawn from a latent structure consisting of multiple curves $\{p^{(\ell)}_k\}_K$. 
-\end{definition}
 
@@
 Finally, we provide theorems that show the consistency of the curve fits and maximum likelihood estimators.
-
-### Estimation of the LSM Curves
 
@@
 
-\begin{algorithm}[H]
-\label{alg:bezier-fit-mixture}
-\DontPrintSemicolon
-\SetAlgoLined
-\setstretch{1.1}
-\KwData{Adjacency matrix $A$, embedding dimension $d$, polynomial degree $R$, number of latent structures $K$, stopping criterion $\epsilon$.}
-\KwResult{Bezier coefficients $\hat{b}$, timepoint values $\hat{t}_1, ..., \hat{t}_n$, curve memberships $\hat{z}_1, ..., \hat{z}_n$.}
-Compute $\hat{X} \in \mathbb{R}^{n \times d}$, the ASE of $A$.\;
-Initialize $\hat{z}_1, ..., \hat{z}_n \in \{1, ..., K\}$, the membership of each vertex to each latent curve.\;
-\For {$k = 1, ..., K$} {
-  Subset the rows of $\hat{X}$ by $z_i = k$ to obtain $\hat{x}_{k_1}, ..., \hat{x}_{k_{n_k}}$, where $n_k$ is the number of vertices with label $k$.\;
-  Initialize timepoints $\hat{t}_{k_1}, ..., \hat{t}_{k_{n_k}}$, (e.g., via a one-dimensional Isomap embedding \citep{Tenenbaum2000-ff}).\;
-}
-\Repeat {the change in equation \ref{eq:bezier-ls} is less than $\epsilon$.} {
-  \For {$k = 1, ..., K$} {
-    Fit $\hat{b}^{(k)}$ by equation \ref{eq:bezier-ols}, using the row subsets of $T$ and $X$ such that $z_i = k$.\;
-    \For {$i = 1, ..., n_k$} {
-      Fit $\hat{t}_{k_i}$ by equation by finding the roots of equation \ref{eq:dbezier} and choosing the roots that minimizes equation \ref{eq:bezier-ls}.\;
-    }
-  }
-  \For {$k = 1, ..., n$} {
-    Reassign $z_i$ based on the latent curve closest to $\hat{x}_i$. 
-  }
-}
-\caption{Procedure for estimating multiple LSM curves as a Bezier polynomial from an adjacency matrix.}
-\end{algorithm}
-
 \begin{theorem}[Consistency of algorithm \ref{alg:bezier-fit}]
@@
 \end{theorem}
-
-### Quasi-Maximum Likelihood Estimation for $\theta$
 
@@
 \end{theorem}
-
-### Classification and Regression
 
@@
 Fit a model for predicting $y^{(\ell)}, ..., y^{(\ell)}$ using the estimated parameters $\theta^{(\ell)}, ..., \theta^{(\ell)}$ (e.g., via linear regression).\;
-\caption{Procedure for fitting a classification or regression model for an MLSM.}
+\caption{Procedure for fitting a regression model for an MLSM.}
 \end{algorithm}
@@
 We set $\sigma^2 = 0.25$ for both scenarios. 
-
-The simulation setup for the first scenario is as follows: For each $N = 128, 256, 512, 1024$, sample $L = 128$ graphs as:
-
-  i. Draw $n_1, ..., n_L$ as $n_\ell \iid \Poisson(N)$.
-  ii. Draw $\theta^{(1)}, ..., \theta^{(L)} \iid \Uniform(\pi / 6, \pi / 3)$.
-  iii.
-For each $\ell = 1, ..., L$, draw latent vectors either from the line segment described by $p_1(t) = \begin{bmatrix} t & 0 \end{bmatrix}^\top$ or $p_2(t) = \begin{bmatrix} \cos(\theta^{(\ell)}) & \sin(\theta^{(\ell)}) \end{bmatrix}^\top$ with probability $1/2$ for each.
-  iv. Use the $L$ sets of latent vectors to draw $L$ RDPGs with adjacency matrices $A^{(1)}, ..., A^{(L)}$.
-  v. Draw each $y_\ell \indep \mathcal{N}(1 + 2 \theta^{(\ell)}, \sigma^2)$.
-  vi.
-Apply algorithm \ref{alg:mlsm-fit} to construct a linear model connecting $y_1, ..., y_\ell$ and $A^{(1)}, ..., A^{(L)}$, using the first half as training data and reporting MSE on the last half not used for training.
-
-For the second the simulation setup is similar. For each $N = 128, 256, 512, 1024$, sample $L = 128$ graphs as:
-
-  i. Draw $n_1, ..., n_L$ as $n_\ell \iid \Poisson(N)$.
-  ii. Draw each $(a^{(\ell)}, b^{(\ell)}) \iid \Uniform([0.5, 2] \times [0.5, 2])$.
-  iii.
-For each $\ell = 1, ..., L$, draw latent vectors by first taking $t_1, ..., t_{n_\ell} \sim \Betadist(a^{(\ell)}, b^{(\ell)})$ then taking each latent vector as $x_i = \begin{bmatrix} t^2 & 2 t (1-t) \end{bmatrix}^\top$.
-  iv. Use the $L$ sets of latent vectors to draw $L$ RDPGs with adjacency matrices $A^{(1)}, ..., A^{(L)}$.
-  v. Draw each $y_\ell \indep \mathcal{N}(1 + 2 \alpha^{(\ell)} + b^{(\ell)}, \sigma^2)$.
-  vi.
-Apply algorithm \ref{alg:mlsm-fit} to construct a linear model connecting $y_1, ..., y_\ell$ and $A^{(1)}, ..., A^{(L)}$, using the first half as training data and reporting MSE on the last half not used for training.
-
-In figure \ref{fig:angle-regression-sim}, we compare algorithm \ref{alg:mlsm-fit} to fitting a regression model using graph transitivity and average degree of each vertex. 
+We also vary the expected number of vertices for each graph: For each $N = 128, 256, 512$, and $1024$, we sample $L = 128$ graphs as $n_1, ..., n_L \iid \Poisson(N)$. 
+Algorithm \ref{alg:mlsm-fit} is applied to each set of $128$ graphs, then half of the estimates are used for fitting the linear regression model and the mean squared error is computed on the remainder.
+To compare our algorithm against existing methods, we extracted graph statistics as covariates for a linear regression model. 
+For each $N$, we ran $50$ replicates to obtain the average and standard error of each method.
+
+In figure \ref{fig:angle-regression-sim}, we compare algorithm \ref{alg:mlsm-fit} to fitting a regression model using graph transitivity and average degree of each vertex (scenario 1). 
 Since this scenario is similar to a multilayer DCBM, we expect performance to be very similar, which is what we observe in the results. 

### 2025-01-24 draft.tex
minor changes

@@
 
-It has been shown \citep{Koo_2022, rubindelanchy2017statistical} that
-the RDPG (as well as the generalized random dot product graph, or GRDPG,
-\citep{rubindelanchy2017statistical}) can describe a wide range of
-popular random graph models, such as the \(\ER\) model, the stochastic
-block model (SBM) \citep{doi:10.1080/0022250X.1971.9989788}, degree
-corrected block model (DCBM) \citep{Karrer_2011}, and popularity
-adjusted block model (PABM) \citep{307cbeb9b1be48299388437423d94bf1}. In
-these examples, the (G)RPDG representation of these networks allows for
-inference and analysis in more familiar Euclidean space rather than on
-graphs, and the geometry of the configuration of latent vectors
-correspond to the parameters of these graph models. More specifically,
-when these network models are described as (G)RDPGs, each latent vector
-lies on a subspace. \citet{athreya2020estimation} and
-\citet{Passino_2020} extended this to the latent structure model (LSM)
-and latent structure block model (LSBM) to include latent configurations
-that consist of nonlinear manifolds.
-
-In this paper, we apply these highly structured random dot product graph
-models to supervised learning problems in which there are multiple
-graphs, \(G_1, ..., G_L\), each with a response
-\(y^{(1)}, ..., y^{(L)}\). Each graph, represented by adjacency matrix
-\(A^{(\ell)}\), is presumed to be drawn from a RDPG with parameter set
-\(\theta_l\), and in turn, the response \(y_l\) is drawn from a
-univariate distribution with parameter \(\theta_l\). Since the latent
-vectors of each \(A^{(\ell)}\) are presumed to be distributed on a
-structured subset of the latent space, e.g., a manifold, we call this
-the multiple latent structure model (MLSM). Much previous work has been
-done on multilayer networks, such as the multilayer degree corrected
-block model of \citet{agterberg2022joint}, multiple random eigen graphs
-of \citet{8889404}, and the common subspace independent edge graphs of
+It has been shown that the RDPG (as well as the generalized random dot
+product graph, or GRDPG, \citep{rubindelanchy2017statistical}) can
+describe a wide range of popular random graph models, such as the
+\(\ER\) model, the stochastic block model (SBM)
+\citep{doi:10.1080/0022250X.1971.9989788}, degree corrected block model
+(DCBM) \citep{Karrer_2011}, and popularity adjusted block model (PABM)
+\citep{307cbeb9b1be48299388437423d94bf1}. In this paper, we apply these
+highly structured random dot product graph models to supervised learning
+problems in which there are multiple graphs, \(G_1, ..., G_L\), each
+with a response \(y^{(1)}, ..., y^{(L)}\). Each graph, represented by
+adjacency matrix \(A^{(\ell)}\), is presumed to be drawn from a RDPG
+with parameter set \(\theta_l\), and in turn, the response \(y_l\) is
+drawn from a univariate distribution with parameter \(\theta_l\). Since
+the latent vectors of each \(A^{(\ell)}\) are presumed to be distributed
+on a structured subset of the latent space, e.g., a manifold, we call
+this the multiple latent structure model (MLSM). There exists much
+existing literature on the estimation for multilayer networks, such as
+the multilayer degree corrected block model of
+\citet{agterberg2022joint}, multiple random eigen graphs of
+\citet{8889404}, and the common subspace independent edge graphs of
 \citet{arroyo2020inference}. In each of these models, the main inference
@@
 
-\subsubsection{Connections to Related
-Models}\label{connections-to-related-models}
-
 In the MLSM defined in definition \ref{def:mlsm}, the only commonality
@@
 
-\subsubsection{Classification and Regression on the
-MLSM}\label{classification-and-regression-on-the-mlsm}
-
-To use the MLSM for classification or regression problems, we assign
-response variables \(y_1, ..., y_L\) to each graph. Then if each
-response \(y_\ell\) depends on \(p^{(\ell)}(t)\) (the parameterization
-of the \(\ell^{th}\) curve) or \(\theta_\ell\) (the parameter of the
-\(\ell^{th}\) distribution), or some combination of the two, there is a
-plausible setup for a predictive modeling task for predicting \(y_\ell\)
-after observing \(A^{(\ell)}\). Four such scenarios are given:
-
-\begin{definition}[MLSM for classification 1]
-\label{def:mlsm-c-1}
-Let $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$, $y_1, ..., y_L \in \{1, ..., M\}$ be labels for each of the $L$ graphs of the MLSM, and each $\theta_\ell$ take on one of $M$ values $\{\phi_1, ..., \phi_M\}$ corresponding to its response, $y_\ell$, i.e., $\theta_\ell = \phi_{y_\ell}$.
-
-We observe the adjacency matrices $A^{(1)}, ..., A^{(L)}$ and the first $r$ labels $y_1, ..., y_r$, and the task is to predict the unknown labels $y_{r+1}, ..., y_L$.
-
-This can be described by the following generative model:
-
-\begin{enumerate}
-  \item Draw labels $y_1, ..., y_L \iid \Categorical(\{1, ..., M\}, \{\pi_1, ..., \pi_M\})$.
-  \item Set each $\theta_\ell = \phi_{y_\ell}$.
-  \item Draw adjacency matrices as $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$.
-\end{enumerate}
-\end{definition}
-
-\begin{definition}[MLSM for classification 2]
-\label{def:mlsm-c-2}
-Let $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{\mathcal{C}^{(\ell)}\}, F, \{\theta_\ell\}_L)$, $y_1, ..., y_L \in \{1, ..., M\}$ be labels for each of the $L$ graphs of the MLSM, and each $\mathcal{C}^{(\ell)}$ take on one of $M$ functional forms $\{p^{(1)}(t), ..., p^{(M)}(t)\}$ corresponding to its response variable, $y_\ell$, i.e., $\mathcal{C}^{(\ell)}$ is parameterized by $p^{(y_\ell)}(t)$.
-
-We observe the adjacency matrices $A^{(1)}, ..., A^{(L)}$ and the first $r$ labels $y_1, ..., y_r$, and the task is to predict the unknown labels $y_{r+1}, ..., y_L$.
-
-This can be described by the following generative model:
-
-\begin{enumerate}
-  \item Draw labels $y_1, ..., y_L \iid \Categorical(\{1, ..., M\}, \{\pi_1, ..., \pi_M\})$.
-  \item Set each $p^{(\ell)}(t) = p^{(y_\ell)}(t)$.
-  \item Draw adjacency matrices as $A^{(1)}, ..., A^{(L)} \sim \MLSM(\{p^{(\ell)}\}, F, \{\theta_\ell\}_L)$.
-\end{enumerate}
-
-As in remark \ref{remark:mlsm-mixture}, we also allow for each adjacency matrix $A^{(\ell)}$ to be drawn from a latent structure consisting of multiple curves $\{p^{(\ell)}_k\}_K$. 
-\end{definition}
+To use the MLSM for regression problems, we assign response variables
+\(y_1, ..., y_L\) to each graph. Then if each response \(y_\ell\)
+depends on \(p^{(\ell)}(t)\) (the parameterization of the \(\ell^{th}\)
+curve) or \(\theta_\ell\) (the parameter of the \(\ell^{th}\)
+distribution), or some combination of the two, there is a plausible
+setup for a predictive modeling task for predicting \(y_\ell\) after
+observing \(A^{(\ell)}\).
 
@@
 and maximum likelihood estimators.
-
-\subsubsection{Estimation of the LSM
-Curves}\label{estimation-of-the-lsm-curves}
 
@@
 \end{theorem}
-
-\subsubsection{\texorpdfstring{Quasi-Maximum Likelihood Estimation for
-\(\theta\)}{Quasi-Maximum Likelihood Estimation for \textbackslash theta}}\label{quasi-maximum-likelihood-estimation-for-theta}
 
@@
 \end{theorem}
-
-\subsubsection{Classification and
-Regression}\label{classification-and-regression}
 
@@
 the distribution of latent vectors along a curve. We set
-\(\sigma^2 = 0.25\) for both scenarios.
-
-The simulation setup for the first scenario is as follows: For each
-\(N = 128, 256, 512, 1024\), sample \(L = 128\) graphs as:
-
-\begin{enumerate}
-\def\labelenumi{\roman{enumi}.}
-\tightlist
-\item
-  Draw \(n_1, ..., n_L\) as \(n_\ell \iid \Poisson(N)\).
-\item
-  Draw
-  \(\theta^{(1)}, ..., \theta^{(L)} \iid \Uniform(\pi / 6, \pi / 3)\).
-\item
-  For each \(\ell = 1, ..., L\), draw latent vectors either from the
-  line segment described by
-  \(p_1(t) = \begin{bmatrix} t & 0 \end{bmatrix}^\top\) or
-  \(p_2(t) = \begin{bmatrix} \cos(\theta^{(\ell)}) & \sin(\theta^{(\ell)}) \end{bmatrix}^\top\)
-  with probability \(1/2\) for each.
-\item
-  Use the \(L\) sets of latent vectors to draw \(L\) RDPGs with
-  adjacency matrices \(A^{(1)}, ..., A^{(L)}\).
-\item
-  Draw each
-  \(y_\ell \indep \mathcal{N}(1 + 2 \theta^{(\ell)}, \sigma^2)\).
-\item
-  Apply algorithm \ref{alg:mlsm-fit} to construct a linear model
-  connecting \(y_1, ..., y_\ell\) and \(A^{(1)}, ..., A^{(L)}\), using
-  the first half as training data and reporting MSE on the last half not
-  used for training.
-\end{enumerate}
-
-For the second the simulation setup is similar. For each
-\(N = 128, 256, 512, 1024\), sample \(L = 128\) graphs as:
-
-\begin{enumerate}
-\def\labelenumi{\roman{enumi}.}
-\tightlist
-\item
-  Draw \(n_1, ..., n_L\) as \(n_\ell \iid \Poisson(N)\).
-\item
-  Draw each
-  \((a^{(\ell)}, b^{(\ell)}) \iid \Uniform([0.5, 2] \times [0.5, 2])\).
-\item
-  For each \(\ell = 1, ..., L\), draw latent vectors by first taking
-  \(t_1, ..., t_{n_\ell} \sim \Betadist(a^{(\ell)}, b^{(\ell)})\) then
-  taking each latent vector as
-  \(x_i = \begin{bmatrix} t^2 & 2 t (1-t) \end{bmatrix}^\top\).
-\item
-  Use the \(L\) sets of latent vectors to draw \(L\) RDPGs with
-  adjacency matrices \(A^{(1)}, ..., A^{(L)}\).
-\item
-  Draw each
-  \(y_\ell \indep \mathcal{N}(1 + 2 \alpha^{(\ell)} + b^{(\ell)}, \sigma^2)\).
-\item
-  Apply algorithm \ref{alg:mlsm-fit} to construct a linear model
-  connecting \(y_1, ..., y_\ell\) and \(A^{(1)}, ..., A^{(L)}\), using
-  the first half as training data and reporting MSE on the last half not
-  used for training.
-\end{enumerate}
+\(\sigma^2 = 0.25\) for both scenarios. We also vary the expected number
+of vertices for each graph: For each \(N = 128, 256, 512\), and
+\(1024\), we sample \(L = 128\) graphs as
+\(n_1, ..., n_L \iid \Poisson(N)\). Algorithm \ref{alg:mlsm-fit} is
+applied to each set of \(128\) graphs, then half of the estimates are
+used for fitting the linear regression model and the mean squared error
+is computed on the remainder. To compare our algorithm against existing
+methods, we extracted graph statistics as covariates for a linear
+regression model. For each \(N\), we ran \(50\) replicates to obtain the
+average and standard error of each method.
 
@@
 \ref{alg:mlsm-fit} to fitting a regression model using graph
-transitivity and average degree of each vertex. Since this scenario is
-similar to a multilayer DCBM, we expect performance to be very similar,
-which is what we observe in the results. However, we cannot model this
-simulation as a multilayer DCBM since although each graph can be
-described as a DCBM, the graphs do not share the same vertex set. In
-figure \ref{fig:beta-regression-sim}, we see that linear regression
-using graph summary statistics performs better for smaller graphs, but
-as we draw more vertices, the ASE improves and so algorithm
+transitivity and average degree of each vertex (scenario 1). Since this
+scenario is similar to a multilayer DCBM, we expect performance to be
+very similar, which is what we observe in the results. However, we
+cannot model this simulation as a multilayer DCBM since although each
+graph can be described as a DCBM, the graphs do not share the same
+vertex set. In figure \ref{fig:beta-regression-sim}, we see that linear
+regression using graph summary statistics performs better for smaller
+graphs, but as we draw more vertices, the ASE improves and so algorithm
 \ref{mlsm-fit} performs better, whereas graph summary statistics cannot

### 2025-01-24 draft.Rmd
minor changes

@@
 
-In figure \ref{fig:angle-regression-sim}, we compare algorithm \ref{alg:mlsm-fit} to fitting a regression model using graph transitivity and average degree of each vertex (scenario 1). 
-Since this scenario is similar to a multilayer DCBM, we expect performance to be very similar, which is what we observe in the results. 
-However, we cannot model this simulation as a multilayer DCBM since although each graph can be described as a DCBM, the graphs do not share the same vertex set. 
-In figure \ref{fig:beta-regression-sim}, we see that linear regression using graph summary statistics performs better for smaller graphs, but as we draw more vertices, the ASE improves and so algorithm \ref{mlsm-fit} performs better, whereas graph summary statistics cannot improve with the number of vertices.
-Figures \ref{fig:theta-mse} and \ref{fig:beta-mse} show that the quasi-MLE estimators computed via algorithm \ref{alg:bezier-param-fit} are consistent.
-
-```{r angle-regression-sim, fig.width = 6, cache = FALSE, fig.cap = 'The average MSE over 50 simulations in scenario 1. The errorbars are the standard error.'}
+```{r angle-regression-sim, fig.width = 6, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE over 50 simulations in scenario 1. The errorbars are the standard error.'}
 n.sigma <- 1
@@
   labs(x = 'Expected vertices per graph', y = 'MSE', colour = 'Predictor') +
-  geom_line(aes(x = n, y = mse.theta, colour = 'estimated angle')) + 
+  geom_line(aes(x = n, y = mse.theta, colour = 'estimated \nangle \n')) + 
   geom_errorbar(aes(x = n, 
@@
                     ymax = mse.theta + se.theta,
-                    colour = 'estimated angle'), 
+                    colour = 'estimated \nangle \n'), 
                 width = .1) + 
@@
   #               width = .1) + 
-  geom_line(aes(x = n, y = mse.a, colour = 'graph statistics')) + 
+  geom_line(aes(x = n, y = mse.a, colour = 'graph \nstatistics')) + 
   geom_errorbar(aes(x = n,
@@
                     ymax = mse.a + se.a,
-                    colour = 'graph statistics'), 
+                    colour = 'graph \nstatistics'), 
                 width = .1) + 
@@
 
-```{r theta-mse, fig.width = 6, cache = FALSE, fig.cap = 'The average MSE of the estimated angles in scenario 1 over 50 simulations. The errorbars are the standard error.'}
+```{r theta-mse, fig.width = 6, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE of the estimated angles in scenario 1 over 50 simulations. The errorbars are the standard error.'}
 angle.reg.df %>% 
@@
 
-```{r beta-regression-sim, fig.width = 6, cache = FALSE, fig.cap = 'The average MSE over 50 simulations in scenario 2. The errorbars are the standard error.'}
+```{r beta-regression-sim, fig.width = 6, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE over 50 simulations in scenario 2. The errorbars are the standard error.'}
 
@@
   ggplot() + 
-  geom_line(aes(x = n, y = mse.beta, colour = 'estimated beta parameters')) + 
+  geom_line(aes(x = n, y = mse.beta, colour = 'estimated \nbeta \nparameters\n')) + 
   geom_errorbar(aes(x = n, 
@@
                     ymax = mse.beta + se.beta,
-                    colour = 'estimated beta parameters'),
+                    colour = 'estimated \nbeta \nparameters\n'),
                 width = .1) + 
-  geom_line(aes(x = n, y = mse.d, colour = 'graph statistics')) + 
+  geom_line(aes(x = n, y = mse.d, colour = 'graph \nstatistics')) + 
   geom_errorbar(aes(x = n,
@@
                     ymax = mse.d + se.d,
-                    colour = 'graph statistics'),
+                    colour = 'graph \nstatistics'),
                 width = .1) + 
@@
 
-```{r beta-mse, fig.width = 6, cache = FALSE, fig.cap = 'The average MSE of the estimated beta parameters, $(a, b)$, in scenario 2 over 50 simulations. The errorbars are the standard error.'}
+```{r beta-mse, fig.width = 6, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE of the estimated beta parameters, $(a, b)$, in scenario 2 over 50 simulations.
+The errorbars are the standard error.'}
 beta.reg.df %>% 
@@
   theme_bw() + 
-  labs(x = 'number of vertices', y = 'MSE of Beta parameter estimates', 
+  labs(x = 'number of vertices', y = 'MSE of Beta \nparameter estimates', 
        colour = 'Beta \ndistribution \nparameter') + 
@@
 ```
+
+In figure \ref{fig:angle-regression-sim}, we compare algorithm \ref{alg:mlsm-fit} to fitting a regression model using graph transitivity and average degree of each vertex (scenario 1). 
+Since this scenario is similar to a multilayer DCBM, we expect performance to be very similar, which is what we observe in the results. 
+However, we cannot model this simulation as a multilayer DCBM since although each graph can be described as a DCBM, the graphs do not share the same vertex set. 
+In figure \ref{fig:beta-regression-sim}, we see that linear regression using graph summary statistics performs better for smaller graphs, but as we draw more vertices, the ASE improves and so algorithm \ref{mlsm-fit} performs better, whereas graph summary statistics cannot improve with the number of vertices.
+Figures \ref{fig:theta-mse} and \ref{fig:beta-mse} show that the quasi-MLE estimators computed via algorithm \ref{alg:bezier-param-fit} are consistent.
 

### 2025-01-24 draft.Rmd
minor updates

@@
 However, we cannot model this simulation as a multilayer DCBM since although each graph can be described as a DCBM, the graphs do not share the same vertex set. 
-In figure \ref{fig:beta-regression-sim}, we see that linear regression using graph summary statistics performs better for smaller graphs, but as we draw more vertices, the ASE improves and so algorithm \ref{mlsm-fit} performs better, whereas graph summary statistics cannot improve with the number of vertices.
+In figure \ref{fig:beta-regression-sim}, we see that linear regression using graph summary statistics performs better for smaller graphs, but as we draw more vertices, the ASE improves and so algorithm \ref{alg:mlsm-fit} performs better, whereas graph summary statistics cannot improve with the number of vertices.
 Figures \ref{fig:theta-mse} and \ref{fig:beta-mse} show that the quasi-MLE estimators computed via algorithm \ref{alg:bezier-param-fit} are consistent.
@@
 
-In the second example, we analyzed the larval *Drosophila* mushroom body connectome \citep{Eichler141762}, which has been studied as a GRDPG by \citet{athreya2020estimation}. 
+In this section, we apply \ref{alg:mslm-fit} to real data to see if we can extract estimates from the latent space that correspond to response variables. 
+In the first example, we analyzed the larval *Drosophila* mushroom body connectome \citep{Eichler141762}, which has been studied as a GRDPG by \citet{athreya2020estimation}. 
 This dataset consists of two graphs representing two networks of neurons, one for each hemisphere of the *Drosophila* brain. 

### 2025-01-24 draft.tex
minor changes
minor updates

@@
 
-In figure \ref{fig:angle-regression-sim}, we compare algorithm
-\ref{alg:mlsm-fit} to fitting a regression model using graph
-transitivity and average degree of each vertex (scenario 1). Since this
-scenario is similar to a multilayer DCBM, we expect performance to be
-very similar, which is what we observe in the results. However, we
-cannot model this simulation as a multilayer DCBM since although each
-graph can be described as a DCBM, the graphs do not share the same
-vertex set. In figure \ref{fig:beta-regression-sim}, we see that linear
-regression using graph summary statistics performs better for smaller
-graphs, but as we draw more vertices, the ASE improves and so algorithm
-\ref{mlsm-fit} performs better, whereas graph summary statistics cannot
-improve with the number of vertices. Figures \ref{fig:theta-mse} and
-\ref{fig:beta-mse} show that the quasi-MLE estimators computed via
-algorithm \ref{alg:bezier-param-fit} are consistent.
-
 \begin{figure}[H]
@@
 
+In figure \ref{fig:angle-regression-sim}, we compare algorithm
+\ref{alg:mlsm-fit} to fitting a regression model using graph
+transitivity and average degree of each vertex (scenario 1). Since this
+scenario is similar to a multilayer DCBM, we expect performance to be
+very similar, which is what we observe in the results. However, we
+cannot model this simulation as a multilayer DCBM since although each
+graph can be described as a DCBM, the graphs do not share the same
+vertex set. In figure \ref{fig:beta-regression-sim}, we see that linear
+regression using graph summary statistics performs better for smaller
+graphs, but as we draw more vertices, the ASE improves and so algorithm
+\ref{alg:mlsm-fit} performs better, whereas graph summary statistics
+cannot improve with the number of vertices. Figures \ref{fig:theta-mse}
+and \ref{fig:beta-mse} show that the quasi-MLE estimators computed via
+algorithm \ref{alg:bezier-param-fit} are consistent.
+
 \section{Applications}\label{applications}
@@
 
-In the second example, we analyzed the larval \emph{Drosophila} mushroom
-body connectome \citep{Eichler141762}, which has been studied as a GRDPG
-by \citet{athreya2020estimation}. This dataset consists of two graphs
-representing two networks of neurons, one for each hemisphere of the
-\emph{Drosophila} brain. In these graphs, each vertex is a neuron, and
-the labels correspond to one of four neuron types (Kenyon Cells, Input
-Neurons, Output Neurons, and Projection Neurons). The number of neurons
-in each hemisphere is not equal (209 in the left hemisphere and 213 in
-the right hemisphere). The resulting graphs are illustrated in figure
-\ref{fig:mbconnectome-graph}.
+In this section, we apply \ref{alg:mslm-fit} to real data to see if we
+can extract estimates from the latent space that correspond to response
+variables. In the first example, we analyzed the larval
+\emph{Drosophila} mushroom body connectome \citep{Eichler141762}, which
+has been studied as a GRDPG by \citet{athreya2020estimation}. This
+dataset consists of two graphs representing two networks of neurons, one
+for each hemisphere of the \emph{Drosophila} brain. In these graphs,
+each vertex is a neuron, and the labels correspond to one of four neuron
+types (Kenyon Cells, Input Neurons, Output Neurons, and Projection
+Neurons). The number of neurons in each hemisphere is not equal (209 in
+the left hemisphere and 213 in the right hemisphere). The resulting
+graphs are illustrated in figure \ref{fig:mbconnectome-graph}.
 

### 2025-10-17 draft.Rmd
finished cleaning up proof of theorem 3, still working on proof of theorem 4

@@
 
-doMC::registerDoMC(parallel::detectCores())
+doMC::registerDoMC(parallel::detectCores() - 2)
 ```
@@
 
-```{r angle-regression-sim, fig.width = 6, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE over 50 simulations in scenario 1. The errorbars are the standard error.'}
+```{r angle-regression-sim, fig.width = 5, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE over 50 simulations in scenario 1. The errorbars are the standard error.'}
 n.sigma <- 1
@@
   labs(x = 'Expected vertices per graph', y = 'MSE', colour = 'Predictor') +
-  geom_line(aes(x = n, y = mse.theta, colour = 'estimated \nangle \n')) + 
+  geom_line(aes(x = n, y = mse.theta, colour = 'Estimated \nangle \n')) + 
   geom_errorbar(aes(x = n, 
@@
                     ymax = mse.theta + se.theta,
-                    colour = 'estimated \nangle \n'), 
+                    colour = 'Estimated \nangle \n'), 
                 width = .1) + 
@@
   #               width = .1) + 
-  geom_line(aes(x = n, y = mse.a, colour = 'graph \nstatistics')) + 
+  geom_line(aes(x = n, y = mse.a, colour = 'Graph \nstatistics')) + 
   geom_errorbar(aes(x = n,
@@
                     ymax = mse.a + se.a,
-                    colour = 'graph \nstatistics'), 
+                    colour = 'Graph \nstatistics'), 
                 width = .1) + 
@@
 
-```{r theta-mse, fig.width = 6, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE of the estimated angles in scenario 1 over 50 simulations. The errorbars are the standard error.'}
+```{r theta-mse, fig.width = 4, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE of the estimated angles in scenario 1 over 50 simulations. The errorbars are the standard error.'}
 angle.reg.df %>% 
@@
 
-```{r beta-regression-sim, fig.width = 6, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE over 50 simulations in scenario 2. The errorbars are the standard error.'}
+```{r beta-regression-sim, fig.width = 5, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE over 50 simulations in scenario 2. The errorbars are the standard error.'}
 
@@
   ggplot() + 
-  geom_line(aes(x = n, y = mse.beta, colour = 'estimated \nbeta \nparameters\n')) + 
+  geom_line(aes(x = n, y = mse.beta, colour = 'Estimated \nbeta \nparameters\n')) + 
   geom_errorbar(aes(x = n, 
@@
                     ymax = mse.beta + se.beta,
-                    colour = 'estimated \nbeta \nparameters\n'),
+                    colour = 'Estimated \nbeta \nparameters\n'),
                 width = .1) + 
-  geom_line(aes(x = n, y = mse.d, colour = 'graph \nstatistics')) + 
+  geom_line(aes(x = n, y = mse.d, colour = 'Graph \nstatistics')) + 
   geom_errorbar(aes(x = n,
@@
                     ymax = mse.d + se.d,
-                    colour = 'graph \nstatistics'),
+                    colour = 'Graph \nstatistics'),
                 width = .1) + 
@@
 
-```{r beta-mse, fig.width = 6, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE of the estimated beta parameters, $(a, b)$, in scenario 2 over 50 simulations.
+```{r beta-mse, fig.width = 5, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE of the estimated beta parameters, $(a, b)$, in scenario 2 over 50 simulations.
 The errorbars are the standard error.'}
@@
 
-In this section, we apply \ref{alg:mslm-fit} to real data to see if we can extract estimates from the latent space that correspond to response variables. 
+In this section, we apply algorithm \ref{alg:mslm-fit} to real data to see if we can extract estimates from the latent space that correspond to response variables. 
 In the first example, we analyzed the larval *Drosophila* mushroom body connectome \citep{Eichler141762}, which has been studied as a GRDPG by \citet{athreya2020estimation}. 
@@
 ```{r load-hcp}
-load('~/Downloads/Data/HCP/HCP-A/DTI/Data.RData')
+load('data/Data.RData')
 z <- dti.info$hemisphere %>% 
@@
 ```{r}
-pairwise.angles.df <- readr::read_csv('~/dev/multilayer-rdpg/pairwise-angles.csv')
+pairwise.angles.df <- readr::read_csv('data/pairwise-angles.csv')
 
@@
 This is unique for the class of Bezier polynomials restricted by algorithm 3. 
-Then $\sum_i^n \|t_i - \hat{t}_i\| \stackrel{p}{\to} 0$. 
+Then $\sum_i^n |t_i - \hat{t}_i| \stackrel{p}{\to} 0$, so each $|t_i - \hat{t}_i| \stackrel{p}{\to} 0$. 
+By construction, the MLE $\hat{\theta}$ and the quasi-MLE $\tilde{\theta}$ are computed via the same function, just on different samples, i.e., $\hat{\theta} = h(t_1, ..., t_n)$ and $\tilde{\theta} = h(\hat{t}_1, ..., \hat{t}_n)$.
+Then since the probability distribution function of each $t_i$ is smooth, $h$ is continuous on the support. 
+Therefore, if each $\hat{t}_i \stackrel{p}{\to} t$, then $h(\hat{t}_1, ..., \hat{t}_n) \stackrel{p}{\to} h(t_1, ..., t_n)$, resulting in $\tilde{\theta} \stackrel{p}{\to} \hat{\theta} \stackrel{p}{\to} \theta$.
+
 \end{proof}
 
+\begin{proof}[Proof of Theorem \ref{thm:supervised}]
+\end{proof}
+
 \newpage

### 2025-10-17 draft.tex
finished cleaning up proof of theorem 3, still working on proof of theorem 4

@@
 
-In this section, we apply \ref{alg:mslm-fit} to real data to see if we
-can extract estimates from the latent space that correspond to response
-variables. In the first example, we analyzed the larval
+In this section, we apply algorithm \ref{alg:mslm-fit} to real data to
+see if we can extract estimates from the latent space that correspond to
+response variables. In the first example, we analyzed the larval
 \emph{Drosophila} mushroom body connectome \citep{Eichler141762}, which
@@
 This is unique for the class of Bezier polynomials restricted by algorithm 3. 
-Then $\sum_i^n \|t_i - \hat{t}_i\| \stackrel{p}{\to} 0$. 
+Then $\sum_i^n |t_i - \hat{t}_i| \stackrel{p}{\to} 0$, so each $|t_i - \hat{t}_i| \stackrel{p}{\to} 0$. 
+By construction, the MLE $\hat{\theta}$ and the quasi-MLE $\tilde{\theta}$ are computed via the same function, just on different samples, i.e., $\hat{\theta} = h(t_1, ..., t_n)$ and $\tilde{\theta} = h(\hat{t}_1, ..., \hat{t}_n)$.
+Then since the probability distribution function of each $t_i$ is smooth, $h$ is continuous on the support. 
+Therefore, if each $\hat{t}_i \stackrel{p}{\to} t$, then $h(\hat{t}_1, ..., \hat{t}_n) \stackrel{p}{\to} h(t_1, ..., t_n)$, resulting in $\tilde{\theta} \stackrel{p}{\to} \hat{\theta} \stackrel{p}{\to} \theta$.
+
+\end{proof}
+
+\begin{proof}[Proof of Theorem \ref{thm:supervised}]
 \end{proof}
